### 2015-10-27 Documentation/fokkerplanck.tex
added documentation

\documentclass[]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[utf8]{inputenc} 
\usepackage{subcaption}
\usepackage{caption}
\usepackage{hyperref} 
%opening
\title{}
\author{}

\newcommand{\R}{\ensuremath{\mathbb{R}}} % commando zonder argumenten
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\norm}[1]{\left\|#1\right\|} % commando's met argumenten
\newcommand{\pa}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\ppa}[2]{\frac{\partial^2 #1}{\partial #2^2}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

We compare solutions of partial differential equations of advection-diffusion-type

\begin{equation}
\label{fokkerplanck}
\pa{\rho(x,t)}{t} + \pa{(a(x) \rho(x,t))}{x} = D  \ppa{\rho(x,t)}{x} 
\end{equation}

generated in two ways:
\begin{itemize}
\item By explicitly solving eq. \eqref{fokkerplanck} using the discretization scheme\footnote{First-order upwind scheme for the advection part combined with the Forward-Time Central-Space-method for the difussion part}
\begin{equation}
\rho_i^{n+1} = \rho_i^n + \Delta t \left( \frac{D}{{\Delta x}^2} \left( \rho_{i+1}^{n} - 2 \rho_i^n + \rho_{i-1}^n \right)  - \frac{a(x)}{\Delta x} (\rho_i^n - \rho_{i-1}^n) \right)
\end{equation}
for the value of the probabilty density $\rho$ at position $i \Delta x$ and time $(n+1) \Delta t$.
\item By simulating an ensemble of $N$ particles evolving according to the corresponding stochastic differential equation  
\begin{equation}
    \dd X = a(X) \dd t + \sqrt{2D} \cdot \dd{W}.
\end{equation}
 The position $X^{n+1}$ at time $ (n+1) \Delta t$ of each  particle is simulated using the  Euler-Maruyama scheme
\begin{equation}
   X^{n + 1} = X^{n} + a(X^n) \Delta t +  \sqrt{2D \Delta t}\cdot \xi^n 
\end{equation}
with $\xi^n  \sim \mathcal{N} (0,1)$
\end{itemize}

\section{}

\end{document}

### 2015-10-27 Problems/Particles/checkSystem/checkSystem.py

@@ -12,12 +12,16 @@ import scipy
 import scipy.linalg
 
 import inv_transform
-import kde
+#import kde
+import histogram
 import particles
 import pde
 import precond
 
-D = 0.01
+# Fokker Planck for SDE
+a = 1  #advection coefficient
+D = 0.1  #difussion coefficient
+lambd = scipy.array([a,D])
 Dt = 0.01 
 N = 100000
 eps = 1e-7
@@ -26,28 +30,25 @@ eps = 1e-7
 h=2e-2 # kde mesh width 
 dx = 0.05
 dt = 1e-3 
-print ("nu : ", dx**2/2./D/dt)
+print("check stability condition")
+print ("nu=", dx**2/(2.*D*dt))  #stability condition: nu>1  (see https://en.wikipedia.org/wiki/FTCS_scheme)
 xL = -2.
 xR = 2.
 grid = scipy.arange(xL+dx/2.,xR,dx)
 rho = scipy.ones_like(grid)/(xR-xL)
 print ("rho: ", rho)
 
-# Fokker Planck for SDE
-a = 1
-D = 0.1
-lambd = scipy.array([a,D])
-
-Dt = 0.01
 
 # Fokker Planck for SDE
 sampler_param = inv_transform.Sampler.getDefaultParameters()
 sampler_param['seed']=0
 lifting = inv_transform.Sampler(sampler_param)
 
-param_kde = kde.KDE.getDefaultParameters()
-param_kde['h']=h
-restriction = kde.KDE(param_kde)
+# param_kde = kde.KDE.getDefaultParameters()
+# param_kde['h']=h
+# restriction = kde.KDE(param_kde)
+
+restriction = histogram.Histogram()
 
 param=particles.Particles.getDefaultParameters()
 param['eps']=eps
@@ -58,10 +59,6 @@ precond_param['Dstar']=0.
 precond_param['sigma']=scipy.zeros_like(grid)
 precond_param['kappa']=particles.doublewell
 param['precond']=precond.Precond(precond_param)
-
-param_kde = kde.KDE.getDefaultParameters()
-param_kde['h']=2e-2
-restriction = kde.KDE(param_kde)
 
 fp_sde = particles.Particles(lifting,restriction,rho,grid,lambd,param)
 
@@ -92,15 +89,14 @@ param['D']=D
 param['Dt'] = Dt
 param['eps_jac']=eps
 
-
[…]

### 2015-10-27 Problems/Particles/residual/residual.py

@@ -11,7 +11,8 @@ import scipy
 import scipy.linalg
 
 import inv_transform
-import kde 
+#import kde 
+import histogram 
 
 import particles
 import pde
@@ -47,9 +48,10 @@ rho_Dt = fp_pde.u_Dt
 
 # particle-based system
 
-param_kde = kde.KDE.getDefaultParameters()
-param_kde['h']=2e-2
-restriction = kde.KDE(param_kde)
+#param_kde = kde.KDE.getDefaultParameters()
+#param_kde['h']=2e-2
+#restriction = kde.KDE(param_kde)
+restriction = histogram.Histogram()
 
 param=particles.Particles.getDefaultParameters()
 param['Dt'] = Dt

### 2015-10-27 Problems/Particles/system/particles.py

@@ -60,7 +60,7 @@ class Particles(TSystem.TimestepperSystem):
     def integrate(self,rho,lambd):
         # lambd = [a,D]
         self.lifting.seed(self.lifting.param['seed'])
-        self.seed(self.param['seed'])
+        #self.seed(self.param['seed'])
         x = self.lift(rho)
         print "nb particles : ", len(x)
         x_Dt = self.simulate(x,lambd)
@@ -68,8 +68,8 @@ class Particles(TSystem.TimestepperSystem):
     
     def lift(self,rho):
         grid = self.grid
-        x = self.lifting.lift(rho,grid,self.param['N'])
-        return x 
+        x, seed = self.lifting.lift(rho,grid,self.param['N'])
+        return x, seed 
     
     def simulate(self,x0,lambd):
         Dt = self.param['Dt']
@@ -101,5 +101,5 @@ class Particles(TSystem.TimestepperSystem):
     def computeJacobian(self):
         dx = self.grid[1]-self.grid[0]
         N = len(self.u)
-        return self.precond.computeJacobian(N,dx) 
+        return self.precond.computeJacobian(N,dx)
     

### 2015-10-27 Problems/Particles/system/precond.py

@@ -41,3 +41,32 @@ class Precond(object):
         J[N-1,N-2] = Dstar/dx**2 + a[N-2]/(2*dx)  
         J[N-1,N-1]   = -2*Dstar/dx**2 + sigma[N-1] + a[N-2]/(2*dx)
         return J
+
+
+
+
+    def computeJacobian(self,N,dx):
+        system = self.system
+        Dstar = self.Dstar
+        kappa = self.kappa
+        sigma = self.sigma 
+
+        edges = (self.system.grid[:-1]+self.system.grid[1:])/2.
+                
+        J=scipy.zeros((N,N),scipy.float64)
+
+        for i in range(1,N-1):
+            J[i,i-1] =Dstar/dx**2 + kappa[i-1]/(2*dx)      
+            J[i,i]   =-2*Dstar/dx**2 + sigma[i]-(kappa[i]-kappa[i-1])/(2*dx)
+            J[i,i+1] =Dstar/dx**2 - kappa[i]/(2*dx)     
+                                  
+        # Homogeneous Dirichlet boundary conditions with boundary points eliminated
+        
+        # left boundary
+        J[0,0]   =-2*Dstar/dx**2 + sigma[0] -kappa[0]/(2*dx)
+        J[0,1] =Dstar/dx**2 - kappa[0]/(2*dx)
+        
+        # right boundary
+        J[N-1,N-2] = Dstar/dx**2 + kappa[N-2]/(2*dx)  
+        J[N-1,N-1]   = -2*Dstar/dx**2 + sigma[N-1] + kappa[N-2]/(2*dx)
+        return J

### 2015-10-27 Problems/WeightedParticles/checkSystem/checkSystem.py

@@ -2,6 +2,10 @@ import tables
 import scipy
 import sys
 
+from scipy.linalg import norm
+
+import matplotlib.pylab as plt
+
 sys.path.append("../system/")
 sys.path.append("../lifting/")
 sys.path.append("../restriction/")
@@ -15,19 +19,20 @@ import Solver.GMRESLinearSolver as GMRES
 import Solver.ImplicitEulerDirectLinearSolver as ImDirect
 import Utils.conditions.probability as probability
 
-#import inv_transform_sampling as inv_transform
-#import inv_transform_determ as inv_transform
-import inv_transform_determ as inv_transform
+import inv_transform_sampling as inv_transform
+ #import inv_transform_determ as inv_transform  (same results for different seeds)
 import histogram
 import particles
 import pde
 import precond
+#import TimestepperSystem
 
 if __name__=="__main__":
      
     D = None
     Dt = None
     N = None
+    M=300  #number of monte carlo steps
     # deze manier van input parameters verwerken staat in Langtangen
     while len(sys.argv) > 1 :
         option = sys.argv[1];               del sys.argv[1]
@@ -43,103 +48,193 @@ if __name__=="__main__":
     if Dt == None:
         Dt = 1e-2
     if N == None:
-        Nlarge = 1000000
-        Nsmall = 10000
+        Nlarge = 1000
+        Nsmall = 1000
     seed = 16
     # discretization parameters
     h=2e-2 # kde mesh width 
-    dx = 0.05
-    dt = 1e-2
-    print "nu : ", dx**2/2./D/dt
+    dx = 0.01
+    dt = 1e-4 # ti
+    print ("nu=", dx**2/(2.*D*dt))  #stability condition: nu>1  (see https://en.wikipedia.org/wiki/FTCS_scheme)
     xL = -1.7
     xR = 1.7
     grid = scipy.arange(xL+dx/2.,xR,dx)
-    rho = scipy.ones_like(grid)
+   # rho = scipy.ones_like(grid)
+    rho = scipy.zeros_like(grid)
+    rho[150]=1
         # 0.1*scipy.random.uniform(low=-0.5,high=0.5,size=scipy.shape(grid)[0])
     rho = rho/(sum(rho)*dx)
     print "rho: ", rho
 
     # Fokker Planck for SDE
-    a = 1
+    a = 0
[…]

### 2015-10-27 Problems/WeightedParticles/lifting/inv_transform_sampling.py

@@ -14,10 +14,10 @@ class Sampler(object):
         return param
     getDefaultParameters=staticmethod(getDefaultParameters)
     
-    def lift(self,rho,grid,N):
+    def lift(self,rho,grid,N,skip=1):
         cum,edges = self.rho2cum(rho,grid)
         y = self.rand.uniform(size=N)
-        print "sampling density with seed : ", self.cur_seed, "first sample: ", y[0]
+   #     print "sampling density with seed : ", self.cur_seed, "first sample: ", y[0]
         return self.inv_cum(y,cum,edges)
     
     def rho2cum(self,rho,grid):
@@ -54,7 +54,7 @@ class Sampler(object):
         return xn
     
     def seed(self,s):
-        print "seeding the density sampler with seed : ", s
+      #  print "seeding the density sampler with seed : ", s
         self.cur_seed = s
         self.rand.seed(self.cur_seed)
     

### 2015-10-27 Problems/WeightedParticles/residual/residual.py

@@ -11,7 +11,8 @@ import scipy
 import scipy.linalg
 
 import inv_transform
-import kde 
+#import kde 
+import histogram
 
 import particles
 import pde
@@ -20,7 +21,10 @@ import precond
 # system parameters
 a = 1
 D = 0.1
-lambd = scipy.array([a,D])
+alpha = 1   
+beta = -1
+lambd = scipy.array([a,D,alpha,beta])
+#lambd = scipy.array([a,D])
 xL = -2.
 xR = 2.
 
@@ -41,21 +45,27 @@ param['D']=D
 param['a']=a
 param['Dt'] = Dt
 param['eps_jac']=eps_jac
-fp_pde = pde.FokkerPlanck(rho0,grid,pde.doublewell,lambd,param)
-rho_Dt = fp_pde.u_Dt
+
 
 # particle-based system
 
-param_kde = kde.KDE.getDefaultParameters()
-param_kde['h']=2e-2
-restriction = kde.KDE(param_kde)
-
+#param_kde = kde.KDE.getDefaultParameters()
+#param_kde['h']=2e-2
+#restriction = kde.KDE(param_kde)
+param_histogram = histogram.Histogram.getDefaultParameters()
+restriction = histogram.Histogram(param_histogram)
+        
 param=particles.Particles.getDefaultParameters()
 param['Dt'] = Dt
 precond_param=precond.Precond.getDefaultParameters()
 precond_param['Dstar']=0.
 precond_param['sigma']=1.
 param['precond']=precond.Precond(precond_param)
+
+
+fp_pde = pde.FokkerPlanck(rho0,grid,pde.doublewell,lambd,param)
+rho_Dt = fp_pde.u_Dt
+
 
 Nlist = scipy.array([10000,20000,50000,100000,200000,500000,700000])
 nN = len(Nlist)

### 2015-10-27 Problems/WeightedParticles/system/particles.py

@@ -11,7 +11,8 @@ import System.TimestepperSystem as TSystem
 
 def create_doublewell(alpha,beta):
     def doublewell(x):
-        return -4*x**3+3*(alpha+beta)*x**2-2*alpha*beta*x
+       # return -4*x**3+3*(alpha+beta)*x**2-2*alpha*beta*x
+        return 2*x-4*x**3  
     return doublewell
 
 class Particles(TSystem.TimestepperSystem):
@@ -45,16 +46,17 @@ class Particles(TSystem.TimestepperSystem):
         self.restriction = restriction
         self.neq = 1
         self.rand=scipy.random.mtrand.RandomState()
-        self.precond = param['precond']
+        self.precond = param['precond']    #  (  keyError: 'precond' )
         self.precond.system = self
         self.Dt = param['Dt']
+        self.dt = param['dt']
         TSystem.TimestepperSystem.__init__(self,rho,lambd,param)
     
     def getDefaultParameters():
         param = {}
         param['Nlarge']=1000000
         param['Nsmall']=10000
-        param['dt']=1e-3
+        param['dt']=1e-4
         param['Dt']=1e-2
         param['seed']=0
         param['eps']=1e-5
@@ -72,9 +74,14 @@ class Particles(TSystem.TimestepperSystem):
     
     def integrate(self,rho,lambd):
         # lambd = [a,D]
-        x,seed = self.lift(rho)
-        print "nb particles : ", len(x)
-        x_Dt = self.simulate(x,seed,lambd)
+        x= self.lift(rho)
+ #       print "x= ", x
+       # print "nb particles : ", len(x)
+        #self.seed(self.param['seed'])
+        seed =self.lifting.seed(self.lifting.param['seed'])
+        seedlist = [seed]*len(x)  #fill list with  same elements (need this datatype for inv_transform_sampling.py)
+        #print "seed= ", seed
+        x_Dt = self.simulate(x,seedlist,lambd)
         return x,x_Dt,self.restrict(x_Dt)
     
     def lift(self,rho):
@@ -82,11 +89,11 @@ class Particles(TSystem.TimestepperSystem):
         x = self.lifting.lift(rho,grid,self.param['Nlarge'],self.skip)
         return x
     
-    def simulate(self,x0,seed,lambd):
-        print "in simulate : ", lambd
+    def simulate(self,x0,seedlist,lambd):
+        #print "in simulate : ", lambd
         x_Dt = scipy.zeros_like(x0)
[…]

### 2015-10-27 Problems/WeightedParticles/system/pde.py

@@ -38,6 +38,7 @@ class FokkerPlanck(TSystem.TimestepperSystem):
         self.precond = param['precond']
         self.precond.system = self
         self.Dt = param['Dt']
+        self.dt = param['dt'] 
         TSystem.TimestepperSystem.__init__(self,rho,lambd,param)    
     
     def integrate(self,rho0,lambd):

### 2015-10-27 System/TimestepperSystem.py

@@ -25,9 +25,9 @@ class TimestepperSystem(System.System):
                 the time-reporting horizon for a matrix-vector product 
         """
         param = {}
-        param['eps'] = 1e-6
+        param['eps'] = 1e-5
         param['dt'] = 1e-3
-        param['Dt'] = 5e-3
+        param['Dt'] = 1e-2
         return param
     getDefaultParameters = staticmethod(getDefaultParameters)
     

### 2015-10-27 plotvar.py

@@ -1,0 +1,29 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Fri Oct 23 15:50:43 2015
+
+@author: pieter
+"""
+
+import numpy as np
+import matplotlib.pylab as plt
+import sys, glob
+
+np.savetxt('results/data/Jv.out', Jv_norm)
+Jv_norm = np.loadtxt('results/data/Jv.out')
+
+M=  len( Jv_norm)
+Jv_var = scipy.zeros(M)
+average= Jv_norm[-1]
+for m in range(1,M):
+    a= np.power((Jv_norm[m] - average),2) 
+    Jv_var[m]= np.sqrt(a)
+
+ax=plt.subplot(111)
+ax.set_xlim(1, 5000)
+ax.set_ylim(0, np.max(Jv_var))
+
+plot_var = plt.plot(Jv_var)
+plt.savefig("results/plots/variance_Jv.pdf")
+
+plt.show()

### 2016-01-15 Documentation/biblio.tex
newton_krylov

@article{avitabile14,
  author = "Avitabile, Daniele and Hoyle, Rebecca and Samaey, Giovanni",
  title = "Noise reduction in coarse bifurcation analysis of stochastic agent-based models: an example of consumer lock-in",
  journal = "SIAM Journal on Applied Dynamical Systems",
  volume = "13",
  number = "4",
  pages = "1583--1619",
  year = "2014",
  url = "https://lirias.kuleuven.be/handle/123456789/492023",
}

@article{watts98,
  author="Watts, Duncan J and Strogatz, Steven H",
  title= {\href{http://www.nature.com/nature/journal/v393/n6684/pdf/393440a0.pdf} {Collective dynamics of 'small-world' networks}},
  journal="Nature",
  volume="393",
  pages="440--442",
  year="1998"
}

### 2016-01-15 Documentation/draft.tex
newton_krylov

\documentclass[]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[utf8]{inputenc} 
\usepackage{subcaption}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsbsy}
\usepackage[left=4cm]{geometry}
 %http://tex.stackexchange.com/questions/595/how-can-i-get-bold-math-symbols
%opening
\title{}
\author{Pieter Van Nuffel}

\newcommand{\R}{\ensuremath{\mathbb{R}}} % commando zonder argumenten
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\E}{\ensuremath{\mathbb{E}}}
\newcommand{\norm}[1]{\left\|#1\right\|} % commando's met argumenten
\newcommand{\pa}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\ppa}[2]{\frac{\partial^2 #1}{\partial #2^2}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\U}{\ensuremath{\boldsymbol{\rho}}}
\newcommand{\cts}{\ensuremath{\boldsymbol{\Phi}_T}} %Coarse time step
\newcommand{\V}{\ensuremath{\mathbf{v}}}
\newcommand{\jv}{\ensuremath{\mathbf{\hat{Jv}}}}
\newcommand{\jvpde}{\ensuremath{\mathbf{Jv}_{FP}}}

\begin{document}

\maketitle

\begin{abstract}

We investigate the emergence of coarse equilibrium states in a stochastic system of p

 system of diffusion processes have a stabilizing force acting on each of them, corresponding to a bistable potential.

We develop a Newton-Krylov method that is able to compute accurately and efficiently coarse fixed points when the underlying fine-scale dynamics is stochastic. The main novelty of the algorithm is in the elimination of the noise that is generated when estimating Jacobian-vector products using time-integration of perturbed initial conditions. We present numerical results that demonstrate 
%the convergence properties of the numerical method, and use the method to show that macroscopic
%fronts in this model destabilise at a coarse symmetry-breaking bifurcation.
%
%We consider a system of diffusion processes that interact through their empirical mean and have a stabilizing force acting on each of them, corresponding to a bistable potential. There are three parameters that characterize the system: the strength of the intrinsic stabilization, the strength of the external random perturbations, and the degree of cooperation or interaction between them. The latter is the rate of mean reversion of each component to the empirical mean of the system. We interpret this model in the context of systemic risk and analyze in detail the effect of cooperation between the components, that is, the rate of mean reversion. We show that in a certain regime of parameters increasing cooperation tends to increase the stability of the individual agents but it also increases the overall or systemic risk. We use the theory of large deviations of diffusions interacting through their mean field.
%
%
%The present paper thus contains two main contributions. First, for the specific
%system under study, we explain the birth of the above-described macroscopic states in
%terms of coarse symmetry-breaking bifurcations. To the best of our knowledge, steps
%in this direction were taken only very recently [55, 7] and were confined to globally
%locked-in states. In the homogeneous case, we follow [5] and interpret metastable
%locked-in states as fixed points of a coarse evolution map. In the limit of infinitely
%many globally-coupled agents with homogeneous product preferences, we derive the
%coarse evolution map analytically. In the case of heterogeneous agents we employ
%stochastic continuation and show for the first time how fronts destabilise to partially
%locked-in states.
%The second main contribution of the paper is the development of a novel procedure
%to obtain coarse Jacobian-vector products with reduced variance, allowing
%the accurate evaluation of Jacobian-vector products in the presence of microscopic
%stochasticity, thus gaining full control over the linear and the nonlinear iterations
%of the Newton-Krylov solver. Even though our implementation of variance-reduced
%Jacobian-vector products is specific to the lock-in model, we believe that analogous
%strategies can be applied to other ABMs. Therefore, we provide a detailed account of
%the algorithmic steps involved in defining an accurate equation-free Newton-Krylov
%method and testing its convergence properties

\end{abstract}

\section{Problem}

\subsection{Model problem}

\input{fokkerplanck_update.tex}

\subsection{Discretization}

\section{Method}

\subsection{Newton-Krylov solver} \label{sec:Newton-Krylov}

If we want to compute steady states for the density $\U_*$ without direct simulation, we can find them by solving the non-linear system

\begin{equation}
  {F}(\U_*) = \U_* - \cts(\U_*) =0. \label{non-linear_system}
\end{equation}
%where $\mathbf{F}$ is nonlinear function from $\R^N$ to  $\R^N$. 
To find the steady state $\U_*$, we apply Newton's method to eq. \ref{non-linear_system} . Starting from an initial state $\U^0$, we iterate

\begin{eqnarray}
\begin{cases}
& \texttt{Solve         }  J(\U^n) \boldsymbol{\delta_n}  =  - {F}(\U^n) \label{linear_system}        \\          
& \texttt{Set     } \U^{n+1} = \U^n+ \boldsymbol{\delta_n} 
\end{cases}
\end{eqnarray}
until convergence. 
$J(\U^n) =   F'(\U^n)  $ denotes the system Jacobian. Each Newton iteration $n$ thus involves evaluating the Jacobian of the timestepper $J(\cts(\U))$.
Since we do not have an explicit formula for $J(\cts)$, we are forced to use an iterative method, such as GMRES, that only requires Jacobian-vector products. %the action of the jacobian times a  vector
%$J(\cts)$, we are forced to use an iterative method, such as GMRES, that only requires Jacobian-vector products.
 \cite{Brown_Krylov}.
The Jacobian $J(\cts)$ applied to a vector $\mathbf{v}$ (with unit norm) will be estimated by a finite difference approximation
\begin{eqnarray}
\label{Jv_approx}
J(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \\
&\approx & \frac{\cts (\U, \boldsymbol{\omega_1} )  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cts (\U, \boldsymbol{\omega_2}) }{\varepsilon} \nonumber
.
\end{eqnarray}

\subsection{Variance reduced Jacobian-vector product}

\section{Analysis}
\subsection{Variance}
\subsection{bias}

\section{Application: Systemic Risk}

%\subsection{Including an interaction term}

Until now, we considered a system of non-interacting particles. We will now extend our model to a mean field model by introducing a third parameter $\alpha$, which is the degree of interaction or cooporation in the system. A simple form of cooperative behaviour is the case where each agent tends to follow the state of the majority (or, each particle feels an attractive force towards the mean state of the system). To include this cooperation effect in our stochastic simulation, we add this mean reversion term to the SDE \eqref{SDE}:

\begin{equation} 
\label{SDE_meanfield}
    \dd x = \mu V(x) \dd t + \sigma  \dd{W_t} + \alpha(\bar{x} -x) \dd t ,
\end{equation}
with  $\bar{x}(t) = \frac{1}{N} \sum_{i=1}^{N} x_i(t)$ denoting the empirical mean.

An interesting application of this model to banks and insurance is the emergence of systemic risk.  Banks will try to minimize their own individual risk by spreading the risk between each other. However, this may increase the risk that they may all fail: reducing individual risk on a micro-scale can increase systemic risk on a macro-scale. 
Garnier, Papanicolaou and Yang already made use of the dynamics in eq. \eqref{SDE_meanfield}  to show that interconnectedness between agents indeed affects the stability of the whole system, causing systemic risk  \cite{Garnier}. They defined $x_i$  as the state of risk of agent $i$. The bi-stable-state structure of the potential $V(x)$ ensures that each risk variable stays around $-1$ (defined as the normal state) or $+1$ (the failed state).  A natural measure of systemic risk is then the transition probability of the empirical mean $\bar{x}$ from the normal state to the failed state.

To establish the idea, let us repeat the numerical simulations with eq. \ref{SDE_meanfield}. The evolution of the system is now characterized by  by the initial conditions, the three parameters ($\mu$, $\sigma$, $\alpha$) and by the system size $N$. Figure \ref{fig:sysrisk}  illustrates the behavior of the empiral mean $\bar{x}$. The simulations were performed with all agents initially in the normal state. Nevertheless, if randomness dominates the interaction, the agents can move immediately to the other potential well. The system then behaves like $N$ independent difussions, and hence, by the symmetry of the potential, the mean state will be attracted to a single mixed state $\bar{x}=0$. Upon increasing the interaction parameter $\alpha$, however, we find two new macroscopic states, suggesting the precence of a pitchfork bifurcation at the macoscopic level. These solutions are no stable steady states, but rather coarse matastable states. Their lifetime is linked to the finite system size.

\begin{center}
\begin{figure}
\includegraphics[width=0.5\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha1}
\includegraphics[width=0.5\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha5}

\caption{The empirical mean, simulated for different $\alpha$. \textit{Left}: the system has one single state $\bar{x}=0$. \textit{Right}: for $\alpha> \alpha_c$ two metastable equilibria emerge. 
 The red dashed lines are approximated  analytical solutions for the steady states. \label{fig:sysrisk}}
\end{figure}
\end{center}

 \begin{center}
\begin{table}
\caption{Parameter values}
  \begin{tabular} { | c  c | c |}    \hline   
     %FTCS-scheme  &   Euler-Maruyama-scheme  \\  
   \textit{ {Discretization parameters}}&    &  SDE    \\ \hline
    Discretization step  & $\Delta x $  & $10^{-2}$ \\ 
        Number of discretization steps  & $n_x$ &  340 \\ 
    Time step  &  $\Delta t$ &   $10^{-2}$ \\ 
       Number of timesteps  & $n$  &  $10^{6}$ \\ \hline
  \end{tabular}
%\quad
%     \begin{tabular}  { | c | c | c |  }  \hline
%     \multicolumn{3}{|c|} {\textit{ System parameters}   }    \\
%    \hline  
%    Diffusion coefficient  & $D $ & $0.5$ \\ 
%   Drift coefficient &  $\mu$ & $ 1$ \\ \hline
%  \end{tabular}
\end{table}
\end{center}

\subsection{Bifurcation study}

\subsubsection{Calculating fixed points}
The steady  states are computed with the Newton-Krylov solver described in section \ref{sec:Newton-Krylov} and compared with the approximated analytical solutions, calculated by Garnier et. al for small $h$ \cite{Garnier}.

\begin{center}
\begin{figure}
\includegraphics[width=1\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/steady_states.pdf}
\caption{\label{fig:bif_anal}}
\end{figure}
\end{center}

\subsubsection{Continuation}
We use a pseudo-arclength contunaution method with secant prediction steps.
In fig. \ref{todo} we show a bifurcation diagram of the densities.

\begin{center}
\begin{figure}
\includegraphics[width=0.8\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/bif_alpha}
\caption{\label{fig:bif_anal}}
\end{figure}
\end{center}

%\begin{equation}
%\xi = sqrt(1-3*C)*(1+ 6*mu/(sigma**2)*C**2*(1-2*C)/(1-3*C))
%\end{equation}  
%
%

By adding the mean reversion term, the Fokker-Plack equation \eqref{fokkerplanck}  describing the evolution of the density, now becomes

\begin{equation}
\label{fp_mean_field}
\pa{\rho(x,t)}{t} = - \mu \pa{( V(x) \rho(x,t))}{x} - \alpha \pa{}{x} \left[ \left(\int x \rho(x,t) \mathrm{d}x  -x \right) \rho(x,t) \right] + \frac{\sigma^2}{2}  \ppa{\rho(x,t)}{x} .
\end{equation}

Explicit solutions of eq. \label{fp_mean_field} are not available in general, but we can find equilibrium solutions. Assuming

\bibliographystyle{plain}
\bibliography{biblio.bib}

\end{document}

### 2016-01-15 Documentation/fokkerplanck.tex
newton_krylov

@@
 \usepackage{hyperref}
+\usepackage{amsbsy}
+\usepackage[left=4.5cm]{geometry}
+%\usepackage{cancel}
+ %http://tex.stackexchange.com/questions/595/how-can-i-get-bold-math-symbols
 %opening
@@
 
-\newcommand{\R}{\ensuremath{\mathbb{R}}} % commando zonder argumenten
+%\newcommand{\R}{\ensuremath{\mathbb{R}}} % commando zonder argumenten
 \newcommand{\C}{\ensuremath{\mathbb{C}}}
 \newcommand{\N}{\ensuremath{\mathbb{N}}}
+\newcommand{\E}{\ensuremath{\mathbb{E}}}
 \newcommand{\norm}[1]{\left\|#1\right\|} % commando's met argumenten
@@
 \newcommand{\dd}{\ensuremath{\mathrm{d}}}
+\newcommand{\U}{\ensuremath{\boldsymbol{\rho}}}
+\newcommand{\cts}{\ensuremath{\boldsymbol{\Phi}_T}} %Coarse time step
+\newcommand{\V}{\ensuremath{\mathbf{v}}}
+\newcommand{\jv}{\ensuremath{\mathbf{\hat{Jv}}}}
+\newcommand{\jvpde}{\ensuremath{\mathbf{Jv}_{FP}}}
+%\newcommand{\expect}[1]{\ensuremath { \mathbb{E} \left[ #1  \right] }
+
 \begin{document}
@@
 
-We compare solutions of partial differential equations of advection-diffusion-type
+\section{Introduction}
+
+We look for solutions of partial differential equations of advection-diffusion-type
 
@@
 
-generated in two ways:
+in two ways:
 \begin{itemize}
 \item By explicitly solving eq.
-\eqref{fokkerplanck} using the discretization scheme\footnote{First-order upwind scheme for the advection part combined with the Forward-Time Central-Space-method for the difussion part}
+\eqref{fokkerplanck} using the discretization scheme\footnote{First-order upwind scheme for the advective part combined with the Forward-Time Central-Space-method for the diffusive part}
 \begin{equation} 
+\label{pde_discretization}
 \rho_i^{n+1} = \rho_i^n + \Delta t \left( \frac{D}{{\Delta x}^2} \left( \rho_{i+1}^{n} - 2 \rho_i^n + \rho_{i-1}^n \right)  - \frac{a(x)}{\Delta x} (\rho_i^n - \rho_{i-1}^n) \right)
 \end{equation}
-for the value of the probabilty density $\rho$ at position $i \Delta x$ and time $(n+1) \Delta t$.
+for the value of the probabilty density $\rho$ at position $x=i \Delta x$ and time $t=(n+1) \Delta t$.
 \item By simulating an ensemble of $N$ particles evolving according to the corresponding stochastic differential equation  
 \begin{equation} 
-    \dd X = a(X) \dd t + \sqrt{2D} \cdot \dd{W}.
+\label{SDE}
+    \dd x = a(x) \dd t + \sqrt{2D} \cdot \dd{W_t}.
 \end{equation}
- The position $X^{n+1}$ at time $ (n+1) \Delta t$ of each  particle is simulated using the  Euler-Maruyama scheme
+ The position $x^{n+1}$ of each  particle  at time $t= (n+1) \Delta t$ is simulated using the Euler-Maruyama scheme
 \begin{equation}
-   X^{n + 1} = X^{n} + a(X^n) \Delta t +  \sqrt{2D \Delta t}\cdot \xi^n 
+   x^{n + 1} = x^{n} + a(x^n) \Delta t +  \sqrt{2D \Delta t}\cdot \xi^n \label{Euler-Mar}
 \end{equation}
@@
 
-\section{}
+\section{Coarse time stepper}
+
+To simulate the time evolution of  the density $\U(t)$, we construct  a coarse time stepper $\cts$ which allows the performance of time-steps at the macroscopic level, using only the stochastic simulation of the position vectors of the $N$ particles at the microscopic level, %$\mathbf{x}(t+ n \Delta t)$,
+generated by eq. \eqref{Euler-Mar}.
+
+To achieve this, we will define two operators (lifting in section \ref{section:lifting} and
+restriction in section \ref{section:restriction}) that relate the microscopic and macroscopic levels of description.
+Once these lifting $\mathcal{L}$ and restriction  operators $\mathcal{R}$ have been constructed, a coarse time-stepper $\cts$ 
+to evolve the macroscopic state $\U$ over a time interval of length $n \Delta t$ is constructed as a three-step-procedure (lift–evolve–restrict):
+
+\begin{equation}
+\U(t + n \Delta t) = \cts(\U) = (\mathcal{R} \circ \mathcal{E}(n \Delta t) \circ  \mathcal{L}(\mathbf{\omega})  ) (\U(t))
+\end{equation}
+where $\mathcal{E}(n \Delta t)(\U(t))$ is the simulation of the SDE for $N$ particles over $n$ timesteps.
+
+\subsection{Lifting: $\U \rightarrow \mathbf{x} $}
+Given the density \U, we need to sample a position vector $x_i$ for every particle $i \leq N$. We use the inverse transformation method for this. \label{section:lifting}
+The particle does not only gets an inital position, but also a seed for generating random steps in the simulation.
+
+\subsection{Restriction:  $\mathbf{x} \rightarrow \U $}
+\label{section:restriction}
+The restriction operator $\mathcal{R}: \mathbb{Q}^N \rightarrow \mathbb{Q}^k$ maps the microscopic state $\mathbf{x}$ (determined by the position vectors of $N$ particles) to a density \U, discretisized in $k$ bins.
+This is done by counting the number of particles in every bin $\Delta_j$ for $1\leq j \leq k$: %= x_{j+\frac{1}{2}}-x_{j-\frac{1}{2}}$.
+
+\begin{equation}
+\frac{1}{N} \sum_{i=1}^{N}  w^i \cdot \chi_{\Delta_j}(x^i) = \rho_j  \label{restriction}
+\end{equation}
+
+with 
+\begin{equation}
+\chi_{\Delta_j}(x) = \begin{cases}
+  1 & \mbox{if } x \in \Delta_j, \\
+  0 & \mbox{if } x \notin \Delta_j. \\
+\end{cases}
+\end{equation}
+
+%\begin{equation}
+% \frac{1}{k}\sum_{j=1}^{k} \rho_j =  \frac{1}{N} \sum_{i=1}^{N}  w_i x_i
+%\end{equation}
+
+and setting all weights $w_i =1 $ for $1\leq i \leq N$.
+
+The reason why we explicitly introduced these weights in the restriction operator will be clarified in section \ref{Jv_approx} where we will need to evaluate  the coarse time stepper  $\cts (\U + \varepsilon \mathbf{v})$, now applied to the density shifted with a certain perturbation $\varepsilon \mathbf{v}$.
+%We explicitly introduced these weights in the restriction operator \eqref{restriction} because they can help to reduce the variance in the jacobian-vector-products.
+In the finite difference approximation we need to evaluate $\cts (\U + \varepsilon \mathbf{v})$, the coarse time stepper applied to the density shifted with a certain perturbation.
+%For the weighted restriction operator, we choose different weights $w_i$ for every particle.
+(we only use this for calculating the jacobian-vector-produ cts.
+(Why?
+Well, the weights are just one if we dont use a perturbation)f
+To evaluate the perturbated restriction-operator we will use the weights $ w^i_{\varepsilon}$, determined such that
+\begin{equation}
+\frac{1}{N} \sum_{i=1}^{N}  w^i_{\varepsilon} \cdot \chi_{\Delta_j} (x^i) = \rho_j + \varepsilon v_j . \label{restriction_eps}
+\end{equation}
+We do this by computing the weight per bin as $ w^j_{\varepsilon} = 1+ \frac{\varepsilon v_j}{\rho_j}$ and assign this value to each particle in $\Delta_j$.
+So, small perturbations on the density lead to small perturbations in the weights.
+The advantage of this weighted restriction operator lies in the possibility to use the same realizations $\mathbf{x}$ in the unperturbated \eqref{restriction} and the perturbated  \eqref{restriction_eps} restriction-operator.
+
+\section{Convergence of variance-reduced Jacobian-vector-products}
+
+\subsection{Why do we need variance reduction of the Jacobian-vector products?  \label{section:Jv}}
+
+If we want to compute steady states for the density $\U_*$ (without direct simulation), we can find them by solving the non-linear system
+
+\begin{equation}
+  \U_* - \cts(\U_*) =0.
+\end{equation}
+In each Newton iteration, one needs to solve a linear system involving the Jacobian of \cts, denoted as $D(\cts)$.  Since we do not have an explicit formula for
+$D(\cts)$ we are forced to use an iterative method (such as GMRES) that only
+requires Jacobian-vector products. 
+The Jacobian $D(\cts)$ applied to a vector $\mathbf{v}$ (with unit norm) will be estimated by a finite difference approximation
+\begin{eqnarray}
+\label{Jv_approx}
+D(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \\
+&\approx & \frac{\cts (\U, \boldsymbol{\omega_1} )  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cts (\U, \boldsymbol{\omega_2}) }{\varepsilon} \nonumber
+.
+\end{eqnarray}
+
+If we use the solution of the PDE, eq.
+\eqref{pde_discretization}, the time stepper is deterministic and the calculation of the Jacobian-vector products is straightforward.
+If we use the solutions of the SDE however, we have to deal with numerical noise in evaluating eq.
+\ref{Jv_approx}.
+Because the coarse time-stepper is stochastic, repeating $\cts$ with two sets of random numbers $\boldsymbol{\omega_1},  \boldsymbol{\omega_2}$  will give different results.
+For $ \varepsilon \ll 1$ this will result in an $\mathcal{O}(1/(\varepsilon^2 N))$ variance.
+Consequently the variance on the Jacobian-vector-products will grow unboundedly as $\varepsilon$ tends to zero and $D(\cts) \mathbf{v}$ completely loses the structure of the perturbation \V.
+
+\subsection{How do we reduce the variance of the Jacobian-vector products?}
+This numerical noise can be reduced by using the same random numbers $\boldsymbol{\omega}$ for the unperturbed and perturbed simulations.
+If we apply the weighted restriction operator \eqref{restriction_eps},  we get the same microscopic realizations in the lifting step - the only difference is in the computation of the weights.
+As such,  we impose $\boldsymbol{\omega_1} = \boldsymbol{\omega_2}$ in eq.
+\eqref{Jv_approx} and consequently
+the variance of $D(\cts) \cdot \mathbf{v}$ is bounded and of  $\mathcal{O}(1/ N)$.
+Fig.
+\ref{Var_N} shows that the variance on the stochastic solution for the Jacobian-vector-product converges to zero with $\mathcal{O}(1/ N)$ and that it does not depend on the value of $\varepsilon$.
+
+%\begin{figure}
+%\includegraphics{../results/plots/ms_Jveps-5eps-6}
+%\caption{}
+%\label{Convergence_Jv}
+%\end{figure}
+
+\subsection{Results}
+
+From now on, let us abbreviate the Jacobian-vector product $D(\cts) \cdot \mathbf{v}$  estimated from the for particle-based time-stepper as $\jv$.
+We also chose the perturbation vector $\V$ to be sinusiodal  (the convergence plots will be  independent from $\V$).
+If we calculate $\jv$ from the stochastic simulations with different number of particles $N$, we can test two things: the convergence to zero of the variance on $\jv$ and the convergence to zero of the estimated bias on $\jv$.
+We will also plot the mean squared error (MSE) which incorporates both effects (see eq.
+\eqref{MSE}.
+The bias can be estimated by comparing the expectation value $\norm{\jv}$ with the Jacobian-vector product for the PDE solved on a fine grid, denoted as $\jvpde$.
+The estimated expectation value $\mathbb{E}(\jv)$ is the Jacobian-vector averaged over $M$ stochastic simulations.
+
+%\begin{equation} \label{MSE}
+%\texttt{MSE}(\jv, \mathbf{\bar{Jv} }) = \frac{ \mathbb{E} \left[   \left(  \jv - \mathbf{\bar{Jv}}  \right)^T  \cdot  \left(  \jv - \mathbf{\bar{Jv}}  \right) \right]} {n_x}  =    \texttt{Var}(\jv) +  \left( \texttt{Bias} (\jv)\right)^T \cdot  \left( \texttt{Bias} (\jv)\right)
+%\end{equation}
+%with 
+%\begin{equation} \nonumber
+% \texttt{Var}(\hat{Jv}) = \frac{ \mathbb{E} \left[   \left(  \jv - \mathbb{E}[\jv]   \right)^T  \cdot  \left(   \jv - \mathbb{E}[\jv]    \right) \right]} {n_x} 
+%\end{equation}
+
+\begin{equation} \label{MSE}
+\texttt{MSE}(\jv, \jvpde ) =  \mathbb{E} \left[   \left(  \jv - \jvpde  \right)^2 \right]  =   \texttt{Var}(\jv) +  \left( \mathbf{Bias}(\jv, \jvpde )  \right)^2
+\end{equation}
+with 
+\begin{equation} \label{Var}
+ \texttt{Var}(\jv) =  \mathbb{E} \left[   \left(  \jv - \mathbb{E}[\jv]   \right)^2 \right] 
+\end{equation}
+and 
+\begin{equation} 
+ \mathbf{Bias}(\jv, \jvpde ) = \mathbb{E}[\jv]  - \jvpde ,
+\end{equation}
+where we defined the inproduct of a $n_x$-dimensional vector $\V$ as $\mathbf{v}^2 = \frac{\V^T \cdot \V}{n_x}$.
+Thus, in eq.
+\eqref{MSE} and \eqref{Var} we divide by the number of discretization steps $n_x$ \footnote{If the number of discretisations steps for solving the PDE happens to be  higher than the number of bins $n_x$, we rescale the dimension of $\jvpde$ to $n_x$}.
+This allows us to make a meaningful comparison with solutions on finer or coarser grids.
+
+%\begin{figure}
+%\includegraphics[width=14.5cm]{../Problems/WeightedParticles/checkSystem/plots/MSE_N_eps}
+%\caption{The mean squared error between the stochastic and the deterministic solution for the Jacobian-vector-product converges to zero with $\mathcal{O}(1/ N)$ and it does not depend on the value of $\varepsilon$.}
+%\label{MSE_N}
+%\end{figure}
+
+\begin{figure}
+\includegraphics[width=14.5cm]{../Problems/WeightedParticles/checkSystem/plots/Var_N_eps}
+\caption{The variance on the stochastic solution for the Jacobian-vector-product converges to zero with $\mathcal{O}(1/ N)$ and it does not depend on the value of $\varepsilon$.}
+\label{Var_N}
+\end{figure}
+
+%\begin{figure}
+%\includegraphics[width=14.5cm]{../Problems/WeightedParticles/checkSystem/plots/Bias_N_eps}
+%\caption{The squared bias between the stochastic and the deterministic solution for the Jacobian-vector-product does not depend on the value of $\varepsilon$.}
+%\label{Bias_N}
+%\end{figure}
+
+%
+%\begin{figure}
+%\includegraphics[width=8.5cm]{../Problems/WeightedParticles/checkSystem/plots/25-11Bias_e-_4}
+%\includegraphics[width=8.5cm]{../Problems/WeightedParticles/checkSystem/plots/25-11Bias_e-_5}
+%\includegraphics[width=8.5cm]{../Problems/WeightedParticles/checkSystem/plots/25-11Bias_e-_6}
+%\includegraphics[width=8.5cm]{../Problems/WeightedParticles/checkSystem/plots/25-11Bias_e-_7}
+%\caption{The bias on the stochastic solution for the Jacobian-vector-product converges to zero if the number of particles is increased and it does not depend on the value of $\varepsilon$.}
+%\label{bias_v}
+%\end{figure}
+%
+
+\begin{table}
+
+\caption{Parameter values}
+   %FTCS-scheme  &   Euler-Maruyama-scheme  \\
+
+     \begin{tabular} { | c  c | c | c |}    \hline   
+     %FTCS-scheme  &   Euler-Maruyama-scheme  \\  
+   \textit{ {Discretization parameters}}    &  & PDE  & SDE  \\ \hline
+    Discretization step  & $\Delta x $ & $10^{-4}$  & $10^{-2}$ \\ 
+        Number of discretization steps &  $n_x$ &  $34 000$ & 340 \\ 
+    Time step  &  $\Delta t$ & $10^{-8}$ &  $10^{-4}$ \\ 
+       Number of timesteps  & $n$  &  $10^{6}$ &  100   \\ \hline
+  \end{tabular}
+\quad
+     \begin{tabular}  { | c | c | c |  }  \hline
+     \multicolumn{3}{|c|} {\textit{ System parameters}   }    \\
+    \hline  
+    Diffusion coefficient  & $D $ & $0.5$ \\ 
+   Drift coefficient &  $a$ & $ 1$ \\ \hline
+  \multicolumn{3}{|c|} {\textit{ Simulation parameters}   }    \\ \hline
+   Number of realizations  & $M $ & $100$ \\ \hline
+    Perturbation size & $\varepsilon$  & $ 10^{-5}$ \\ \hline
+  \end{tabular}
+
+\end{table}
+
+%Total simulation time for particle-based time-stepper 
+%
+%Total Simulation time for PDE-based time stepper
+
+%\subsection{Variance}
+%\begin{equation}
+%\mathbb{E} \left[ (\rho -\mathbb{E}[\rho ])^2 \right] =?
+%\end{equation}
+%
+%
+%
+
+%\begin{equation}
+%D(\cts) \cdot \mathbf{v} \approx  \frac{\cts (\U, \boldsymbol{\omega_1} )  + \varepsilon \mathbf{v}  - \cts (\U, \boldsymbol{\omega_2}) }{\varepsilon} 
+%\end{equation}
+
+%\subsection{Convergence}
+
+%\section{Some documentation about the use of random numbers in the simulation}
+%
+%Every MC-step starts with a new seed for seeding the density sampler in the lifting step.
+This same seed is used for the time evolution of every particle.
+So, for a given MC-step and a given timestep, the brownian increment is the same for all particles in the simulation.
+This means that the sum of all Brownian steps is the same for all particles.
+This does not mean that $\Delta(x+\Delta T) - \Delta(x)$ is the same for all particles, because this depends on the drift-term, which depends on the position, which is different for every particle.
+
+\section{Convergence of the bias}
+We calculate the expectation value of $\jv$ by substituting eq. \eqref{} in the finite-difference approximation.
+
+\begin{equation}
+\E[\jv] =  \E \left[ \frac{\cts (\hat{\U} + \varepsilon \mathbf{v} )  - \cts (\hat{\U})}{\varepsilon}  \right]
+\end{equation}
+with \begin{equation}
+\cts (\hat{\U})  = \frac{1}{N}\sum_{i=1}^{N} \chi_{\Delta_j} \left[\mathcal{E} (x_i(t) \right]
+\end{equation}
+and 
+\begin{equation}
+\cts (\hat{\U} + \varepsilon \V )  = \frac{1}{N}\sum_{i=1}^{N}  (1+ \frac{\varepsilon \V}  {\U(x_i(t))}  ) \cdot \chi_{\Delta_j} \left[\mathcal{E} (x_i(t) \right] .
+\end{equation}
+
+\begin{eqnarray}
+\E[\jv] &=&  \E \left[  \frac{1}{N}\sum_{i=1}^{N}   \frac{ \chi_{\Delta_j} \left[  \mathcal{E} (x_i(t)) \right] \V}  {\U(x_i(t))}     \right] \\
+		&=&   \frac{\V}{N} \E \left[  \sum_{i=1}^{N}   \frac{ \chi_{\Delta_j} \left[  \mathcal{E} (x_i(t)) \right] }  {\U(x_i(t))}     \right] \\
+		&=&  \frac{\V}{N}  \sum_{i=1}^{N}   \frac{  \E \left[  \chi_{\Delta_j} \left[  \mathcal{E} (x_i(t)) \right]  \right] }  {\U(x_i(t))} \\
+		&=& \frac{\V}{N}  \sum_{i=1}^{N}   \frac{  \chi_{\Delta_j} \left[  \E \left[   \mathcal{E} (x_i(t)) \right]  \right] }  {\U(x_i(t))} 
+\end{eqnarray}
+
+%\begin{eqnarray}
+%\E[\jv] &=&  \E \left[ \frac{\cts (\hat{\U} + \varepsilon \mathbf{v} )  - \cts (\hat{\U})}{\varepsilon}  \right]
+%        &=&  \E \left[ \frac{1}{N} ] 
+%\end{eqnarray}
+
+\begin{figure}
+\includegraphics[width=14.5cm]{../Problems/WeightedParticles/checkSystem/plots/Bias_Jv_N_M.pdf}
+\caption{The expectation value of the stochastic solution of the Jacobian-vector-product is calculated by averaging over $M$ stochastic realizations.
+The estimated bias between this solution and the deterministic one is squared and plotted as a function of the number of particles $N$ for different values of $M$.
+For large particle numbers, the estimated bias becomes independent of $M$ and converges to the real bias.
+This real bias is the numerical error in the solution of the PDE [Need to check this last statement].}
+\label{Bias_Jv_N_M}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=14.5cm]{../Problems/WeightedParticles/checkSystem/plots/Bias_Jv_M_N.pdf}
+\caption{The expectation value of the stochastic solution of the Jacobian-vector-product is calculated by averaging over $M$ stochastic realizations.
+The estimated bias between this solution and the deterministic one is squared and plotted as a function of the number of realizations $M$ for simulations with different particle number $N$.
+For large particle numbers, the estimated bias becomes independent of $M$ and converges to the real bias.
+This real bias is the numerical error in the solution of the PDE [Need to check this last statement].}
+\label{Bias_Jv_M_N}
+\end{figure}-
+
+\begin{itemize}
+\item
+\end{itemize}
+
+%\section{Simulation}
+%\subsection{Default input values}
+%
+%
+%\begin{itemize}
+%\item $\varepsilon  =10^{-5}$
+% \item $ N =10 000$
+%\end{itemize}
 

### 2016-01-15 Documentation/systemic_risk.tex
newton_krylov

--- 2016-01-15 Documentation/draft.tex
@@
 \usepackage{amsbsy}
-\usepackage[left=4cm]{geometry}
+\usepackage[left=4.5cm]{geometry}
+%\usepackage{cancel}
  %http://tex.stackexchange.com/questions/595/how-can-i-get-bold-math-symbols
@@
 \title{}
-\author{Pieter Van Nuffel}
+\author{}
 
-\newcommand{\R}{\ensuremath{\mathbb{R}}} % commando zonder argumenten
+%\newcommand{\R}{\ensuremath{\mathbb{R}}} % commando zonder argumenten
 \newcommand{\C}{\ensuremath{\mathbb{C}}}
@@
 \newcommand{\jvpde}{\ensuremath{\mathbf{Jv}_{FP}}}
+%\newcommand{\expect}[1]{\ensuremath { \mathbb{E} \left[ #1  \right] }
 
@@
 
-We investigate the emergence of coarse equilibrium states in a stochastic system of p
-
- system of diffusion processes have a stabilizing force acting on each of them, corresponding to a bistable potential.
-
-We develop a Newton-Krylov method that is able to compute accurately and efficiently coarse fixed points when the underlying fine-scale dynamics is stochastic.
-The main novelty of the algorithm is in the elimination of the noise that is generated when estimating Jacobian-vector products using time-integration of perturbed initial conditions.
-We present numerical results that demonstrate
+%
+%We investigate coarse equilibrium states of a stochastic model 
+%
+%
+% we derive an analytic
+%approximate coarse evolution-map for the expected average purchase. We then study the emergence
+%of coarse fronts when the agents are split into two factions with opposite preferences. We
+%develop a novel Newton-Krylov method that is able to compute accurately and efficiently coarse
+%fixed points when the underlying fine-scale dynamics is stochastic. The main novelty of the algorithm
+%is in the elimination of the noise that is generated when estimating Jacobian-vector products
+%using time-integration of perturbed initial conditions. We present numerical results that demonstrate
 %the convergence properties of the numerical method, and use the method to show that macroscopic
@@
 
-\section{Problem}
-
-\subsection{Model problem}
-
-\input{fokkerplanck_update.tex}
-
-\subsection{Discretization}
-
-\section{Method}
-
-\subsection{Newton-Krylov solver} \label{sec:Newton-Krylov}
-
-If we want to compute steady states for the density $\U_*$ without direct simulation, we can find them by solving the non-linear system
+\section{Systemic risk}
 
 \begin{equation}
-  {F}(\U_*) = \U_* - \cts(\U_*) =0. \label{non-linear_system}
-\end{equation}
-%where $\mathbf{F}$ is nonlinear function from $\R^N$ to  $\R^N$. 
-To find the steady state $\U_*$, we apply Newton's method to eq. \ref{non-linear_system} . Starting from an initial state $\U^0$, we iterate
-
-\begin{eqnarray}
-\begin{cases}
-& \texttt{Solve         }  J(\U^n) \boldsymbol{\delta_n}  =  - {F}(\U^n) \label{linear_system}        \\          
-& \texttt{Set     } \U^{n+1} = \U^n+ \boldsymbol{\delta_n} 
-\end{cases}
-\end{eqnarray}
-until convergence. 
-$J(\U^n) =   F'(\U^n)  $ denotes the system Jacobian. Each Newton iteration $n$ thus involves evaluating the Jacobian of the timestepper $J(\cts(\U))$.
-Since we do not have an explicit formula for $J(\cts)$, we are forced to use an iterative method, such as GMRES, that only requires Jacobian-vector products.
-%the action of the jacobian times a  vector
-%$J(\cts)$, we are forced to use an iterative method, such as GMRES, that only requires Jacobian-vector products.
- \cite{Brown_Krylov}.
-The Jacobian $J(\cts)$ applied to a vector $\mathbf{v}$ (with unit norm) will be estimated by a finite difference approximation
-\begin{eqnarray}
-\label{Jv_approx}
-J(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \\
-&\approx & \frac{\cts (\U, \boldsymbol{\omega_1} )  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cts (\U, \boldsymbol{\omega_2}) }{\varepsilon} \nonumber
-.
-\end{eqnarray}
-
-\subsection{Variance reduced Jacobian-vector product}
-
-\section{Analysis}
-\subsection{Variance}
-\subsection{bias}
-
-\section{Application: Systemic Risk}
-
-%\subsection{Including an interaction term}
-
-Until now, we considered a system of non-interacting particles.
-We will now extend our model to a mean field model by introducing a third parameter $\alpha$, which is the degree of interaction or cooporation in the system.
-A simple form of cooperative behaviour is the case where each agent tends to follow the state of the majority (or, each particle feels an attractive force towards the mean state of the system).
-To include this cooperation effect in our stochastic simulation, we add this mean reversion term to the SDE \eqref{SDE}:
-
-\begin{equation} 
-\label{SDE_meanfield}
-    \dd x = \mu V(x) \dd t + \sigma  \dd{W_t} + \alpha(\bar{x} -x) \dd t ,
-\end{equation}
-with  $\bar{x}(t) = \frac{1}{N} \sum_{i=1}^{N} x_i(t)$ denoting the empirical mean.
-
-An interesting application of this model to banks and insurance is the emergence of systemic risk.
-Banks will try to minimize their own individual risk by spreading the risk between each other.
-However, this may increase the risk that they may all fail: reducing individual risk on a micro-scale can increase systemic risk on a macro-scale.
-Garnier, Papanicolaou and Yang already made use of the dynamics in eq.
-\eqref{SDE_meanfield}  to show that interconnectedness between agents indeed affects the stability of the whole system, causing systemic risk  \cite{Garnier}.
-They defined $x_i$  as the state of risk of agent $i$.
-The bi-stable-state structure of the potential $V(x)$ ensures that each risk variable stays around $-1$ (defined as the normal state) or $+1$ (the failed state).
-A natural measure of systemic risk is then the transition probability of the empirical mean $\bar{x}$ from the normal state to the failed state.
-
-To establish the idea, let us repeat the numerical simulations with eq.
-\ref{SDE_meanfield}.
-The evolution of the system is now characterized by  by the initial conditions, the three parameters ($\mu$, $\sigma$, $\alpha$) and by the system size $N$.
-Figure \ref{fig:sysrisk}  illustrates the behavior of the empiral mean $\bar{x}$.
-The simulations were performed with all agents initially in the normal state.
-Nevertheless, if randomness dominates the interaction, the agents can move immediately to the other potential well.
-The system then behaves like $N$ independent difussions, and hence, by the symmetry of the potential, the mean state will be attracted to a single mixed state $\bar{x}=0$.
-Upon increasing the interaction parameter $\alpha$, however, we find two new macroscopic states, suggesting the precence of a pitchfork bifurcation at the macoscopic level.
-These solutions are no stable steady states, but rather coarse matastable states.
-Their lifetime is linked to the finite system size.
-
-\begin{center}
-\begin{figure}
-\includegraphics[width=0.5\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha1}
-\includegraphics[width=0.5\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha5}
-
-\caption{The empirical mean, simulated for different $\alpha$. \textit{Left}: the system has one single state $\bar{x}=0$. \textit{Right}: for $\alpha> \alpha_c$ two metastable equilibria emerge. 
- The red dashed lines are approximated  analytical solutions for the steady states. \label{fig:sysrisk}}
-\end{figure}
-\end{center}
-
- \begin{center}
-\begin{table}
-\caption{Parameter values}
-  \begin{tabular} { | c  c | c |}    \hline   
-     %FTCS-scheme  &   Euler-Maruyama-scheme  \\  
-   \textit{ {Discretization parameters}}&    &  SDE    \\ \hline
-    Discretization step  & $\Delta x $  & $10^{-2}$ \\ 
-        Number of discretization steps  & $n_x$ &  340 \\ 
-    Time step  &  $\Delta t$ &   $10^{-2}$ \\ 
-       Number of timesteps  & $n$  &  $10^{6}$ \\ \hline
-  \end{tabular}
-%\quad
-%     \begin{tabular}  { | c | c | c |  }  \hline
-%     \multicolumn{3}{|c|} {\textit{ System parameters}   }    \\
-%    \hline  
-%    Diffusion coefficient  & $D $ & $0.5$ \\ 
-%   Drift coefficient &  $\mu$ & $ 1$ \\ \hline
-%  \end{tabular}
-\end{table}
-\end{center}
-
-\subsection{Bifurcation study}
-
-\subsubsection{Calculating fixed points}
-The steady  states are computed with the Newton-Krylov solver described in section \ref{sec:Newton-Krylov} and compared with the approximated analytical solutions, calculated by Garnier et.
-al for small $h$ \cite{Garnier}.
-
-\begin{center}
-\begin{figure}
-\includegraphics[width=1\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/steady_states.pdf}
-\caption{\label{fig:bif_anal}}
-\end{figure}
-\end{center}
-
-\subsubsection{Continuation}
-We use a pseudo-arclength contunaution method with secant prediction steps.
-In fig. \ref{todo} we show a bifurcation diagram of the densities.
-
-\begin{center}
-\begin{figure}
-\includegraphics[width=0.8\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/bif_alpha}
-\caption{\label{fig:bif_anal}}
-\end{figure}
-\end{center}
-
-%\begin{equation}
-%\xi = sqrt(1-3*C)*(1+ 6*mu/(sigma**2)*C**2*(1-2*C)/(1-3*C))
-%\end{equation}  
-%
-%
-
-By adding the mean reversion term, the Fokker-Plack equation \eqref{fokkerplanck}  describing the evolution of the density, now becomes
-
-\begin{equation}
-\label{fp_mean_field}
-\pa{\rho(x,t)}{t} = - \mu \pa{( V(x) \rho(x,t))}{x} - \alpha \pa{}{x} \left[ \left(\int x \rho(x,t) \mathrm{d}x  -x \right) \rho(x,t) \right] + \frac{\sigma^2}{2}  \ppa{\rho(x,t)}{x} .
+\label{fokkerplanck}
+\pa{\rho(x,t)}{t} = - \mu \pa{( U(x) \rho(x,t))}{x} - \alpha \pa{}{x} \left[ \left(\int x \rho(x,t) \mathrm{d}x  -x \right) \rho(x,t) \right] + D  \ppa{\rho(x,t)}{x} 
 \end{equation}
 
-Explicit solutions of eq. \label{fp_mean_field} are not available in general, but we can find equilibrium solutions. Assuming
+\begin{figure}
+\includegraphics{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha5}
 
-\bibliographystyle{plain}
-\bibliography{biblio.bib}
+\end{figure}
 

### 2016-01-15 Documentation/talk_15-12.tex
newton_krylov

\documentclass{beamer}
%\usepackage[dutch]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[makeroom]{cancel}
\usepackage{color}
\newcommand<>{\xxcancel}[1]{\alt#2{\cancel{#1}\vphantom{#1}}{#1}} %has apparently no effect
%\renewcommand{\CancelColor}{blue}

\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\E}{\ensuremath{\mathbb{E}}}
\newcommand{\norm}[1]{\left\|#1\right\|} % commando's met argumenten
\newcommand{\pa}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\ppa}[2]{\frac{\partial^2 #1}{\partial #2^2}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\U}{\ensuremath{\boldsymbol{\rho}}}
\newcommand{\cts}{\ensuremath{\boldsymbol{\Phi}_T}} %Coarse time step
\newcommand{\V}{\ensuremath{\mathbf{v}}}
\newcommand{\jv}{\ensuremath{\mathbf{\hat{Jv}}}}
\newcommand{\jvpde}{\ensuremath{\mathbf{Jv}_{FP}}}

\begin{document}

 % % %

\section[Pieter Van Nuffel]{}

\begin{frame}
\frametitle{Pieter Van Nuffel}

\begin{itemize}
\item 2006-2011: Master in Physics and Astronomy (UGent)
	\begin{itemize}
	\item {\textbf{Master thesis}}: Simulation of Information Entropy in Financial Markets with Molecular Dynamics.
	\end{itemize}
\item 2012-2015: Teaching assistent (KU Leuven KULAK)
\item 2015-current: PhD student
	\begin{itemize}
	\item {\textbf{Promotor}}: Giovanni Samaey
	\item {\textbf{Research topics and interests}}: 
	Coarse Bifurcation Analysis of Stochastic Agent-based
	Network Models, Uncertainty Quantification,  Multiscale Methods
	\end{itemize}
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Fokker-Planck equation}

\begin{equation*}
\label{fokkerplanck}
\pa{\rho(x,t)}{t} + \pa{(a(x) \rho(x,t))}{x} = D  \ppa{\rho(x,t)}{x} 
\end{equation*}
\begin{center}
\includegraphics[width=6.5cm]{../Problems/WeightedParticles/checkSystem/plots/rho_pde}
\end{center}

Simulate an ensemble of $N$ particles evolving according to the corresponding SDE
\begin{equation*}
    \dd x = a(x) \dd t + \sqrt{2D} \cdot \dd{W_t}.
\end{equation*}

\end{frame}

\begin{frame}

\frametitle{Coarse time stepper}

\begin{equation*}
\U(t + n \Delta t) = \cts(\U) = (\mathcal{R} \circ \mathcal{E}(n \Delta t, \omega) \circ  \mathcal{L}(\mathbf{\omega})  ) (\U(t))
\end{equation*}

\begin{itemize} [<+->]
\item 
\begin{block} 
{Lifting $\mathcal{L}: \U \rightarrow \mathbf{x}  $}
\end{block}
\item
 \begin{block} {Evolution  $\mathcal{E}$: \\ Simulation of the SDE for $N$ particles over $n$ timesteps. }
\begin{equation*}
   \mathbf{x}^{n + 1} = \mathbf{x}^{n} + a(\mathbf{x}^n) \Delta t +  \sqrt{2D \Delta t}\cdot \boldsymbol{\xi}^n 
\end{equation*}
with \begin{equation*}\xi^n(\omega)  \sim \mathcal{N} (0,1)
\end{equation*}
\end{block}

\item 
\begin{block} {Restriction $\mathcal{R}: \mathbf{x} \rightarrow \U $}
\begin{equation*}
\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(x_i) = \rho_j  
\end{equation*}
\end{block}

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Evaluating Jacobian-vector products}
\begin{equation*}
 \U_* - \cts(\U_*) =0
\end{equation*}

\begin{equation*}
\mathbf{Jv} = D(\cts) \cdot \mathbf{v} \approx \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} 
\end{equation*}

\pause
\begin{block}{Problem: numerical noise for $ \varepsilon \ll 1$ }
Repeating $\cts$ with two sets of random numbers $\boldsymbol{\omega_1},  \boldsymbol{\omega_2}$  will give different results
$\rightarrow  \mathcal{O}(1/(\varepsilon^2 N))$ variance. 
\end{block}

\end{frame}

\begin{frame} \frametitle{Variance reduction of Jacobian-vector products}

\begin{eqnarray}
\mathbf{Jv} = D(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \nonumber \\
&\approx & \frac{\cts (\U, \boldsymbol{\omega_1} )  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cts (\U, \boldsymbol{\omega_2}) }{\varepsilon} \nonumber
\end{eqnarray}

\pause 
\begin{block}{Solution: \\ Perturbations on the density $\rightarrow$  perturbations in the weights }
\begin{eqnarray*}
\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(x_i) &=& \rho_j  \\
\frac{1}{N} \sum_{i=1}^{N}  w^i_{\varepsilon} \cdot \chi_{\Delta_j} (x^i) &=& \rho_j + \varepsilon v_j .
\end{eqnarray*}
For the perturbated density: compute the weight per bin as $ w^j_{\varepsilon} = 1+ \frac{\varepsilon v_j}{\rho_j}$ and assign this value to each particle in $\Delta_j$.
\end{block}

\end{frame}

\begin{frame} \frametitle{Variance reduction of Jacobian-vector products}

\begin{eqnarray}
\mathbf{Jv} = D(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \nonumber \\
&\approx &  \frac{ \cancel{\cts (\U, \boldsymbol{\omega_1} ) }  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cancel{ \cts (\U, \boldsymbol{\omega_1}) } }{\varepsilon} \nonumber
\end{eqnarray}

\begin{block}{Solution: \\ Perturbations on the density $\rightarrow$  perturbations in the weights }
\begin{eqnarray*}
\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(x_i) &=& \rho_j  \\
\frac{1}{N} \sum_{i=1}^{N}  w^i_{\varepsilon} \cdot \chi_{\Delta_j} (x^i) &=& \rho_j + \varepsilon v_j .
\end{eqnarray*}
For the perturbated density: compute the weight per bin as $ w^j_{\varepsilon} = 1+ \frac{\varepsilon v_j}{\rho_j}$ and assign this value to each particle in $\Delta_j$.
\end{block}

\end{frame}

\begin{frame} \frametitle{Variance reduction of Jacobian-vector products}

\begin{eqnarray}
\mathbf{Jv} = D(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \nonumber \\
&\approx &  \frac{ \xxcancel{\cts (\U, \boldsymbol{\omega_1} ) }  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \xxcancel{ \cts (\U, \boldsymbol{\omega_1}) } }{\varepsilon} \nonumber
\end{eqnarray}

\begin{block}{Solution: \\ Perturbations on the density $\rightarrow$  perturbations in the weights }
\begin{eqnarray*}
\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(x_i) &=& \rho_j  \\
\frac{1}{N} \sum_{i=1}^{N}  w^i_{\varepsilon} \cdot \chi_{\Delta_j} (x^i) &=& \rho_j + \varepsilon v_j .
\end{eqnarray*}
For the perturbated density: compute the weight per bin as $ w^j_{\varepsilon} = 1+ \frac{\varepsilon v_j}{\rho_j}$ and assign this value to each particle in $\Delta_j$.
\end{block}
\end{frame}

\begin{frame} \frametitle{Variance reduction of Jacobian-vector products}

\begin{eqnarray}
\mathbf{Jv} = D(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \nonumber \\
&\approx &  \frac{ \xxcancel{\cts (\U, \boldsymbol{\omega_1} ) }  + \xxcancel{\varepsilon} D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \xxcancel{ \cts (\U, \boldsymbol{\omega_1}) } }{\xxcancel{\varepsilon}} \nonumber
\end{eqnarray}

\begin{block}{Solution: \\ Perturbations on the density $\rightarrow$  perturbations in the weights }
\begin{eqnarray*}
\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(x_i) &=& \rho_j  \\
\frac{1}{N} \sum_{i=1}^{N}  w^i_{\varepsilon} \cdot \chi_{\Delta_j} (x^i) &=& \rho_j + \varepsilon v_j .
\end{eqnarray*}
For the perturbated density: compute the weight per bin as $ w^j_{\varepsilon} = 1+ \frac{\varepsilon v_j}{\rho_j}$ and assign this value to each particle in $\Delta_j$.
\end{block}
\end{frame}

\begin{frame}
\frametitle{Variance reduction of Jacobian-vector products}

\begin{equation*} 
 \texttt{Var}(\jv) =  \hat{\mathbb{E}} \left[   \left(  \jv - \hat{\mathbb{E}}[\jv]   \right)^2 \right] \sim \mathcal{O}(1/ N)
\end{equation*}

\includegraphics[width=9.5cm]{../Problems/WeightedParticles/checkSystem/plots/Var_N_eps}

\end{frame}

\begin{frame}
\frametitle{Evaluating the bias on Jacobian-vector products}

\begin{equation*} 
 \mathbf{Bias}(\jv, \jvpde ) = \mathbb{\hat{E}}[\jv]  - \jvpde 
\end{equation*}

\pause

with 
\begin{equation*}
\jvpde \approx \frac{ \U^{\varepsilon}(t + n \Delta t)  - \U (t +  n\Delta t)}{\varepsilon}
\end{equation*}
 \pause

and $\U$ calculated by explicitly solving

\begin{equation*}
\label{fokkerplanck}
\pa{\rho(x,t)}{t} + \pa{(a(x) \rho(x,t))}{x} = D  \ppa{\rho(x,t)}{x} 
\end{equation*}

 using

\begin{equation*} 
\rho_i^{n+1} = \rho_i^n + \Delta t \left( \frac{D}{{\Delta x}^2} \left( \rho_{i+1}^{n} - 2 \rho_i^n + \rho_{i-1}^n \right)  - \frac{a(x)}{\Delta x} (\rho_i^n - \rho_{i-1}^n) \right)
\end{equation*}
for the value of $\rho$ at position $x=i \Delta x$ and time $t=(n+1) \Delta t$

\end{frame}

\begin{frame}
\frametitle{Evaluating the bias on Jacobian-vector products}

\begin{equation*} 
 \mathbf{Bias}(\jv, \jvpde ) = \mathbb{\hat{E}}[\jv]  - \jvpde 
\end{equation*}

 \begin{figure}
 \includegraphics[width=9.5cm]{../Problems/WeightedParticles/checkSystem/plots/Bias_Jv_N_M.pdf}

 \end{figure}

\end{frame}

\begin{frame}
\frametitle{Evaluating the bias on Jacobian-vector products}

\begin{equation*} 
 \mathbf{Bias}(\jv, \jvpde ) = \mathbb{\hat{E}}[\jv]  - \jvpde 
\end{equation*}

 \begin{figure}
 \includegraphics[width=9.5cm]{../Problems/WeightedParticles/checkSystem/plots/Bias_Jv_M_N.pdf}

 \end{figure}

\end{frame}

\end{document}

### 2016-01-15 Point/Point.py

@@ -116,7 +116,7 @@ class Point(scipy.sparse.linalg.LinearOperator):
     def setState(self,u,lambd):
         self.u=u
         self.lambd=lambd
-        self.system.setState(u,lambd)
+    #    self.system.setState(u,lambd)  #need  to use this again for continuation
         # this is meant to deal with preconditioner matrix that needs to be
         # built if the state changes.
         if not self.precond == None:
@@ -405,7 +405,7 @@ class Point(scipy.sparse.linalg.LinearOperator):
         sfree=s[n:n+l]
         sart=s[n+l:]
         sx_norm=norm(sx)
-        sfree_norm=norm(sfree)
+       # sfree_norm=norm(sfree)
         
         result=scipy.zeros(scipy.shape(s),scipy.float64)
         

### 2016-01-15 Problems/Particles/lifting/.#inv_transform.py

@@ -1,1 +1,0 @@
-pieter@hactar.8758:1443514223

### 2016-01-15 Problems/Particles/residual/.#residual.py

@@ -1,1 +1,0 @@
-pieter@hactar.3786:1443424054

### 2016-01-15 Problems/Particles/system/particles.py

@@ -37,8 +37,8 @@ class Particles(TSystem.TimestepperSystem):
         self.restriction = restriction
         self.neq = 1
         self.rand=scipy.random.mtrand.RandomState()
-        self.precond = param['precond']
-        self.precond.system = self
+        #self.precond = param['precond']
+        #self.precond.system = self
         self.Dt = param['Dt']
         TSystem.TimestepperSystem.__init__(self,rho,lambd,param)
     
@@ -68,8 +68,8 @@ class Particles(TSystem.TimestepperSystem):
     
     def lift(self,rho):
         grid = self.grid
-        x, seed = self.lifting.lift(rho,grid,self.param['N'])
-        return x, seed 
+        x = self.lifting.lift(rho,grid,self.param['N'])
+        return x
     
     def simulate(self,x0,lambd):
         Dt = self.param['Dt']
@@ -87,6 +87,9 @@ class Particles(TSystem.TimestepperSystem):
             tcur += dt
             # the random number
             dW=self.rand.normal(loc=0.,scale=scipy.sqrt(2*D*dt),size=N)
+          #  if tcur == dt:    #only print random number for first time step
+          #      print 'dW =',  dW            
+            
             # the process
             drift_term = a * drift(x)
             x=x+drift_term*dt+dW

### 2016-01-15 Problems/WeightedParticles/checkSystem/checkSystem.py

@@ -32,7 +32,7 @@ if __name__=="__main__":
     D = None
     Dt = None
     N = None
-    M=300  #number of monte carlo steps
+    M=1  #number of monte carlo steps
     # deze manier van input parameters verwerken staat in Langtangen
     while len(sys.argv) > 1 :
         option = sys.argv[1];               del sys.argv[1]
@@ -47,27 +47,31 @@ if __name__=="__main__":
         D = 1./2.
     if Dt == None:
         Dt = 1e-2
+        
     if N == None:
         Nlarge = 1000
         Nsmall = 1000
-    seed = 16
+    seed = 1
     # discretization parameters
     h=2e-2 # kde mesh width 
     dx = 0.01
-    dt = 1e-4 # ti
-    print ("nu=", dx**2/(2.*D*dt))  #stability condition: nu>1  (see https://en.wikipedia.org/wiki/FTCS_scheme)
+    dt = 1e-5 # ti
+    r= (2.*D*dt)/dx**2
+    if r>1: 
+        print 'Stability condition not fulfilled, because r=',r #(see https://en.wikipedia.org/wiki/FTCS_scheme)
+        sys.exit("Ending Script. Make sure that r<1") 
     xL = -1.7
     xR = 1.7
     grid = scipy.arange(xL+dx/2.,xR,dx)
-   # rho = scipy.ones_like(grid)
-    rho = scipy.zeros_like(grid)
-    rho[150]=1
+
+    rho = scipy.ones_like(grid)
+    #rho[150]=1
         # 0.1*scipy.random.uniform(low=-0.5,high=0.5,size=scipy.shape(grid)[0])
     rho = rho/(sum(rho)*dx)
     print "rho: ", rho
 
     # Fokker Planck for SDE
-    a = 0
+    a = 1
     zeta = 0.
   #  alphas = [1.]
     alpha=1
@@ -97,11 +101,6 @@ if __name__=="__main__":
     sampler_param = inv_transform.Sampler.getDefaultParameters()
     sampler_param['seed']=1
     lifting1 = inv_transform.Sampler(sampler_param)
-
-    #sampler_param = inv_transform.Sampler.getDefaultParameters()
-    sampler_param['seed']=2
-    lifting2 = inv_transform.Sampler(sampler_param)
-
    
     param_histogram = histogram.Histogram.getDefaultParameters()
[…]

### 2016-01-15 Problems/WeightedParticles/checkSystem/check_bif_sde.py

@@ -1,0 +1,232 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+#import matplotlib.pylab as plt
+
+sys.path.append("../system/")
+sys.path.append("../lifting/")
+sys.path.append("../restriction/")
+sys.path.append("..")
+sys.path.append("../..")
+
+
+import histogram
+import particles
+import inv_transform_sampling as inv_transform
+
+import Point.Point as Point
+import Solver.NewtonSolver as NewtonSolver
+import Solver.GMRESLinearSolver as GMRES
+import Solver.ImplicitEulerDirectLinearSolver as ImDirect
+import Utils.conditions.probability as probability
+#import Continuer.Continuer as Continuer 
+
+if __name__=="__main__":
+    print 'Start solving sde'
+    D = 1./2.
+    Dt = 100
+    seed = 16
+    # discretization parameters
+   #h=2e-2 # kde mesh width 
+   # dt=1e-6
+    xL = -1.7
+    xR = 1.7
+    
+    a = 1
+    zeta = 0.
+    alpha=1
+    beta = -1
+    lambd = scipy.array([a,D,alpha,beta,zeta])
+            
+    #SDE
+    t1 = time.time()
+    
+    print 'Start solving sde'
+    dx = 1e-2
+#    factor = int (dx/dx_fp)
+#    print "Discretisation for solving sde is ", factor, " times coarser than the discretisation for solving the pde"
+    dt = 1e-2 # ti
+    N=5000
+    Nlarge= N
+    Nsmall =N
+    
+    
+    grid = scipy.arange(xL+dx/2.,xR,dx)
+    rho = scipy.ones_like(grid)
+    rho = rho/(sum(rho)*dx)
+    
+    
+
+    v=scipy.zeros_like(grid)
+    for j in range(len(grid)): 
+        v[j]= np.sin(j*2*np.pi/len(grid))
+
+#    v=scipy.zeros_like(grid)
+#    v[20]=1
+#    v[30]=-1
+     
+    
+  #  sampler_param = inv_transform.Sampler.getDefaultParameters()
+  #  sampler_param['seed']=0
+  #  lifting =  inv_transform.Sampler(sampler_param)
+    
+    h=2e-2 # kde mesh width     
+    M=1 #number of monte carlo steps                
[…]

### 2016-01-15 Problems/WeightedParticles/checkSystem/check_pde.py

@@ -1,0 +1,81 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+
+import tables
+import scipy
+import sys
+import numpy as np
+
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+#import matplotlib.pylab as plt
+
+sys.path.append("../system/")
+sys.path.append("../lifting/")
+sys.path.append("../restriction/")
+sys.path.append("..")
+sys.path.append("../..")
+sys.path.append("../../..")
+
+import pde
+
+
+if __name__=="__main__":
+    D = 1./2.
+    Dt = 1e-2
+    seed = 16
+    # discretization parameters
+   #h=2e-2 # kde mesh width 
+    dx = 1e-3
+#    dxlist= [1e-2,  5e-3, 1e-3, 5e-4] 
+    dxlist= [1e-2, 1e-3]
+    dt = min(dxlist)**2/(2*D)
+    #r= (2.*D*dt)/(dx)**2
+    r= (2.*D*dt)/(min(dxlist))**2
+    if r>1: 
+        print 'Stability condition not fulfilled, because r=',r #(see https://en.wikipedia.org/wiki/FTCS_scheme)
+        sys.exit("Ending Script. Make sure that r<1") 
+    xL = -1.7
+    xR = 1.7
+    
+    a = 1
+    zeta = 0.
+    alpha=1
+    beta = -1
+    lambd = scipy.array([a,D,alpha,beta,zeta])
+    param=pde.FokkerPlanck.getDefaultParameters()
+    param['Dt'] = Dt
+    param['dt']=dt         
+    norm_Jv_fp= scipy.zeros(len(dxlist))
+    norm_rho_fp= scipy.zeros(len(dxlist))
+   # rho_Dt_fp[i]= np.ndarray(shape=) #[scipy.zeros((len
+    
+    for i in range (0, len(dxlist)):
+        dx =dxlist[i]
+        print dx
+        grid = scipy.arange(xL+dx/2.,xR,dx)
+        rho = scipy.ones_like(grid)
+        rho = rho/(sum(rho)*dx)
+        print "rho: ", rho
+ 
+        fp_pde = pde.FokkerPlanck(rho,grid,pde.doublewell,lambd,param)
+
+        v=scipy.zeros_like(grid)
[…]

### 2016-01-15 Problems/WeightedParticles/checkSystem/check_sde.py

@@ -1,0 +1,232 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+#import matplotlib.pylab as plt
+
+sys.path.append("../system/")
+sys.path.append("../lifting/")
+sys.path.append("../restriction/")
+sys.path.append("..")
+sys.path.append("../..")
+
+
+import histogram
+import particles
+import inv_transform_sampling as inv_transform
+
+import Point.Point as Point
+import Solver.NewtonSolver as NewtonSolver
+import Solver.GMRESLinearSolver as GMRES
+import Solver.ImplicitEulerDirectLinearSolver as ImDirect
+import Utils.conditions.probability as probability
+#import Continuer.Continuer as Continuer 
+
+if __name__=="__main__":
+    print 'Start solving sde'
+    D = 1./2.
+    Dt = 100
+    seed = 16
+    # discretization parameters
+   #h=2e-2 # kde mesh width 
+   # dt=1e-6
+    xL = -1.7
+    xR = 1.7
+    
+    a = 1
+    zeta = 0.
+    alpha=1
+    beta = -1
+    lambd = scipy.array([a,D,alpha,beta,zeta])
+            
+    #SDE
+    t1 = time.time()
+    
+    print 'Start solving sde'
+    dx = 1e-2
+#    factor = int (dx/dx_fp)
+#    print "Discretisation for solving sde is ", factor, " times coarser than the discretisation for solving the pde"
+    dt = 1e-2 # ti
+    N=1000
+    Nlarge= N
+    Nsmall =N
+    
+    
+    grid = scipy.arange(xL+dx/2.,xR,dx)
+    rho = scipy.ones_like(grid)
+    rho = rho/(sum(rho)*dx)
+    
+    
+
+    v=scipy.zeros_like(grid)
+    for j in range(len(grid)): 
+        v[j]= np.sin(j*2*np.pi/len(grid))
+
+#    v=scipy.zeros_like(grid)
+#    v[20]=1
+#    v[30]=-1
+     
+    
+  #  sampler_param = inv_transform.Sampler.getDefaultParameters()
+  #  sampler_param['seed']=0
+  #  lifting =  inv_transform.Sampler(sampler_param)
+    
+    h=2e-2 # kde mesh width     
+    M=1 #number of monte carlo steps                
[…]

### 2016-01-15 Problems/WeightedParticles/checkSystem/post_processing.py

@@ -1,0 +1,177 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+import matplotlib.pylab as plt
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+import matplotlib.pylab as plt
+
+
+if __name__=="__main__":
+
+
+
+    rho_Dt_fp= np.loadtxt('data/obs/25-11-Rho_pde_dx_e-4_eps-5.out') 
+    Jv_pde= np.loadtxt('data/obs/25-11-Jv_pde_dx_e-4_eps-5.out')
+    
+    sde_rho_sq = np.loadtxt('data/obs/25-11-rho_sq_eps-5.out')
+    E_rho = np.loadtxt('data/obs/25-11-E_rho_eps-5.out')
+    sq_E_rho  =np.loadtxt('data/obs/25-11-sq_E_rho_eps-5.out')
+    
+    sde_Jv_sq = np.loadtxt('data/obs/25-11-Jv_sq_eps-5.out')
+    E_Jv = np.loadtxt('data/obs/25-11-E_Jv_eps-5.out')   #lots of data, not needed when only interested in the variance
+    sq_E_Jv =  np.loadtxt('data/obs/25-11-sq_E_Jv_eps-5.out' )
+    
+    
+
+            
+    #SDE
+    rho_Dt_sde = E_rho[-1]
+    Jv_sde = E_Jv[-1]
+
+    
+    resize_factor = int (len(rho_Dt_fp)/len(rho_Dt_sde))
+  #  resize_factor = 10
+    print "Discretisation for solving sde is ",  resize_factor , " times coarser than the discretisation for solving the pde"
+    
+    rho_coarse = scipy.zeros(len(rho_Dt_sde))
+    for i in range (0,len(rho_coarse)):
+        bin_av = 0
+        for j in range (0,  resize_factor ):
+            bin_av = bin_av+ rho_Dt_fp[i*resize_factor+j]
+        rho_coarse[i] = bin_av/resize_factor
+        
+    bins =  len(Jv_sde)
+    Jv_coarse = scipy.zeros(bins)
+    for i in range (0,len(Jv_coarse)):
[…]

### 2016-01-15 Problems/WeightedParticles/gmres/gmres_sde.py

@@ -17,7 +17,7 @@ import Utils.conditions.probability as probability
 
 #import inv_transform_sampling as inv_transform
 #import inv_transform_determ as inv_transform
-import inv_transform_determ as inv_transform
+import inv_transform_sampling as inv_transform
 import histogram
 import particles
 import pde
@@ -43,7 +43,7 @@ if __name__=="__main__":
     if Dt == None:
         Dt = 1e-3
     if N == None:
-        Nlarge = 1000000
+        Nlarge = 100000
         Nsmall = 1000
         
     seed = 16
@@ -63,76 +63,75 @@ if __name__=="__main__":
     # Fokker Planck for SDE
     a = 1
     zeta = 0.
-    alphas = [1.,0.999]
+    alpha = 1.
     beta = -1
 
     p = None
     fp_sde = None
-    
-    for i in range(len(alphas)):
-    
-        alpha = alphas[i]
-        print "alpha ... ",  alpha
 
-        # Fokker Planck for SDE
-        lambd = scipy.array([a,D,alpha,beta,zeta])
+    # Fokker Planck for SDE
+    lambd = scipy.array([a,D,alpha,beta,zeta])
 
-        sampler_param = inv_transform.Sampler.getDefaultParameters()
-        sampler_param['seed']=seed
-        lifting = inv_transform.Sampler(sampler_param)
+    sampler_param = inv_transform.Sampler.getDefaultParameters()
+    sampler_param['seed']=seed
+    lifting = inv_transform.Sampler(sampler_param)
    
-        param_histogram = histogram.Histogram.getDefaultParameters()
-        param_histogram['h']=h
-        restriction = histogram.Histogram(param_histogram)
+    param_histogram = histogram.Histogram.getDefaultParameters()
+    param_histogram['h']=h
+    restriction = histogram.Histogram(param_histogram)
 
-        param=particles.Particles.getDefaultParameters()
-        param['Dt'] = Dt
-        param['Nlarge']=Nlarge   
-        param['Nsmall']=Nsmall   
-        param['dt']=dt
-        precond_param=precond.Precond.getDefaultParameters()
-        precond_param['Dstar']=1./2.
-        precond_param['sigma']=scipy.zeros_like(grid)
-        precond_param['kappa']=scipy.zeros_like(grid)
[…]

### 2016-01-15 Problems/WeightedParticles/lifting/inv_transform_sampling.py

@@ -7,6 +7,13 @@ class Sampler(object):
         self.param = param
         self.rand = scipy.random.mtrand.RandomState()
         self.seed(param['seed'])
+
+
+    def set_rg_state(self, rg_state):
+        self.rand.set_state(rg_state) 
+    
+    def get_rg_state(self):
+        return self.rand.get_state()
     
     def getDefaultParameters():
         param = {}
@@ -17,6 +24,7 @@ class Sampler(object):
     def lift(self,rho,grid,N,skip=1):
         cum,edges = self.rho2cum(rho,grid)
         y = self.rand.uniform(size=N)
+      #  print y
    #     print "sampling density with seed : ", self.cur_seed, "first sample: ", y[0]
         return self.inv_cum(y,cum,edges)
     
@@ -34,7 +42,7 @@ class Sampler(object):
             if rho[n]<0 :
                 print "encountered a negative density : ", rho[n]
             cum[n+1]=cum[n]+rho[n]*(edges[n+1]-edges[n])
-        print "estimating cumulative density : last value ", cum[-1], " (should be 1.)"
+        print "estimating cumulative density : last value = ", cum[-1], " (should be 1.)"
         if scipy.absolute(cum[-1]-1)>1e-7:
             print "cumulative: ", cum
         return cum,edges
@@ -54,7 +62,7 @@ class Sampler(object):
         return xn
     
     def seed(self,s):
-      #  print "seeding the density sampler with seed : ", s
+        print "seeding the density sampler with seed : ", s
         self.cur_seed = s
         self.rand.seed(self.cur_seed)
     

### 2016-01-15 Problems/WeightedParticles/restriction/histogram.py

@@ -19,6 +19,3 @@ class Histogram(object):
         edges = scipy.r_[domain[0],(grid[1:]+grid[:-1])/2.,domain[-1]]
         bins = scipy.digitize(x,edges)-1
         return bins
-    
-
-

### 2016-01-15 Problems/WeightedParticles/system/particles.py

@@ -1,7 +1,7 @@
 import scipy
 from scipy.linalg import norm
 from scipy import sqrt
-
+import numpy as np
 import sys         
 sys.path.append('../')   #added to resolve importerror of nephew-directories
 sys.path.append('../../')
@@ -25,32 +25,44 @@ class Particles(TSystem.TimestepperSystem):
     
     rho_t + [a*(x-x**3) \rho]_x = D \rho_xx 
     """
-    def __init__(self,lifting,restriction,rho,grid,lambd,control=None,param=None):
+    def __init__(self,lifting,restriction,rho,grid,lambd, x_prev=None, w_prev=None, control=None,param=None):
+   
         if param == None :
             param = Particles.getDefaultParameters()
         self.param = param
         self.drift=create_doublewell(alpha=lambd[2],beta=lambd[3])
         self.control= control
-        if control == None:
-            self.N=param['Nlarge']
-            self.skip=1
-        else:
-            self.N=param['Nsmall']
-            self.skip = param['Nlarge']/param['Nsmall']
+#        if control == None:
+#            self.N=param['Nlarge']
+#            self.skip=1
+#        else:
+#            self.N=param['Nsmall']
+#            self.skip = param['Nlarge']/param['Nsmall']
+        self.skip=1
+        self.N= param['Nsmall']
+        self.Nlarge= param['Nlarge']
         self.grid = grid
         # grid points are cell centers
         xL = 3./2.*grid[0]-1./2.*grid[1]
         xR = 3./2.*grid[-1]-1./2.*grid[-2]
         self.domain = scipy.array([xL,xR])
+       # self.discr_factor= param['discr_factor']
         self.lifting = lifting
         self.restriction = restriction
         self.neq = 1
-        self.rand=scipy.random.mtrand.RandomState()
-        self.precond = param['precond']    #  (  keyError: 'precond' )
-        self.precond.system = self
+        self.rand=scipy.random.mtrand.RandomState() #https://docs.scipy.org/doc/numpy-1.6.0/reference/generated/numpy.random.mtrand.RandomState.html
+     #   if(rg_state != None):
+   #         self.rand.set_state(rg_state) #tried this for new random numbers in simulations with different particles numbers, but not needed
+
+  #      self.precond = param['precond']    #  (  keyError: 'precond' )
+#       self.precond.system = self
         self.Dt = param['Dt']
[…]

### 2016-01-15 Problems/WeightedParticles/system/pde.py

@@ -15,8 +15,8 @@ def doublewell(x):
 
 
 def d2udx2(u,dx):
-    u_bc=scipy.r_[u[0],u,u[-1]]
-    return 1./dx**2*(u_bc[2:]-2*u_bc[1:-1]+u_bc[:-2]) 
+    u_bc=scipy.r_[u[0],u,u[-1]]        #add first and last element (so they appear two time and u_bc[1:-1] is original vector with same dimesnion then u_bc[:-2] )
+    return 1./dx**2*(u_bc[2:]-2*u_bc[1:-1]+u_bc[:-2]) #FTCS scheme
 
     
 class FokkerPlanck(TSystem.TimestepperSystem):
@@ -35,8 +35,8 @@ class FokkerPlanck(TSystem.TimestepperSystem):
             param = FokkerPlanck.getDefaultParameters()
         self.param = param
         self.neq = 1
-        self.precond = param['precond']
-        self.precond.system = self
+      #  self.precond = param['precond']
+ #       self.precond.system = self
         self.Dt = param['Dt']
         self.dt = param['dt'] 
         TSystem.TimestepperSystem.__init__(self,rho,lambd,param)    

### 2016-01-15 System/TimestepperSystem.py

@@ -4,6 +4,8 @@ from scipy.linalg import norm
 import System
 
 
+
+
 class TimestepperSystem(System.System):
 
     def __init__(self,u,lambd,parameters=None):
@@ -40,6 +42,8 @@ class TimestepperSystem(System.System):
         self.u = u
         self.lambd = lambd
         self.u_Dt = self.integrate(u,lambd)
+       # self.x_prev = x_prev
+        print self.u_Dt
         print "leaving setState"
     def getResidual(self):
         """
@@ -67,10 +71,12 @@ class TimestepperSystem(System.System):
         If the system Jacobian is J, this function returns (an
         approximation to) (1-J)*v 
         """
+        
+        print "WE ARE IN FP"
         eps = self.param['eps']
         u_eps = self.u + eps * v/norm(v)
         u_eps_Dt = self.integrate(u_eps,self.lambd)
-        result = v-(u_eps_Dt - self.u_Dt)/eps*norm(v)
+        result = (u_eps_Dt - self.u_Dt)/eps*norm(v)
         return result   
     
     def getParameterDerivative(self,i):

### 2016-01-15 plotvar.py

@@ -8,22 +8,46 @@ Created on Fri Oct 23 15:50:43 2015
 import numpy as np
 import matplotlib.pylab as plt
 import sys, glob
+import scipy
 
-np.savetxt('results/data/Jv.out', Jv_norm)
-Jv_norm = np.loadtxt('results/data/Jv.out')
 
-M=  len( Jv_norm)
-Jv_var = scipy.zeros(M)
-average= Jv_norm[-1]
-for m in range(1,M):
-    a= np.power((Jv_norm[m] - average),2) 
-    Jv_var[m]= np.sqrt(a)
+#Nlist = scipy.array([1000,2000,3000,4000,5000,6000,7000,8000,9000,10000])
+Nlist = scipy.array([1000,2000,4000,8000,16000,32000])
+#np.savetxt('results/data/Jv_ms_e-6.out' , Jv_ms_list)
+#np.savetxt('results/data/rho_ms_e-6.out' , rho_ms_list)
+Jv_ms_eps6 = np.loadtxt('results/data/cluster/Jv_ms_eps-6.out' )
+#rho_ms_e6 = np.loadtxt('results/data/rho_ms_e-6.out' )
+Jv_ms_eps5 = np.loadtxt('results/data/cluster/Jv_ms_eps-5.out' )
+#rho_ms_e6_cl = np.loadtxt('results/data/cluster/rho_ms_e-6.out' )
+#Jv_ms_e5 = np.loadtxt('results/data/Jv_ms_e-5.out' )
+#rho_ms_e5 = np.loadtxt('results/data/rho_ms.out' )
 
-ax=plt.subplot(111)
-ax.set_xlim(1, 5000)
-ax.set_ylim(0, np.max(Jv_var))
+#plot_ms_rho = plt.plot(Nlist, rho_ms_list)
+plot6 = plt.plot(Nlist, Jv_ms_eps6, label = '$\epsilon=10^{-6}$')
+#plot6cl = plt.plot(Nlist, Jv_ms_e6_cl, label = '$\epsilon=10^{-6}$')
+plot5 = plt.plot(Nlist, Jv_ms_eps5, label = '$\epsilon=10^{-5}$')
 
-plot_var = plt.plot(Jv_var)
-plt.savefig("results/plots/variance_Jv.pdf")
+plt.xscale('log')
+plt.yscale('log')
 
[…]

### 2016-04-15 Documentation/GMRES.tex
bifurcation diagramma

\subsubsection{Analysis of GMRES} \label{GMRES}

The $k$th GMRES iteration minimizes the residual $r= $ over $x_0$ + $\mathcal{K}_k$, where $x_0$ is the initial iterate and $\mathcal{K}_k$ is the the $k$th Krylov subspace
$\mathcal{K_k} = \texttt{Span} \{ r_0, J \cdot r_0, . . . , J_{k-1} \cdot r_0 \}$.
Clearly, (fast) matrix-vector products play a crucial role in generating this sequence since each subsequent vector in the sequence is obtain from the previous one by multiplication by $A$.

\subsubsection{experiments}
For $N=1e^6$ particles and $\Delta_T =0.1$: variance and bias becomes lower, for decreasing GMRES-tolerance, until $\epsilon_{GMRES}= 1e-6$, then blow-up was observed.

\begin{figure}
\includegraphics[width=0.5\linewidth]{../Problems/WeightedParticles/checkSystem/plots/GMRES/GMRES_N_Dt_1e-1}
\includegraphics[width=0.5\linewidth]{../Problems/WeightedParticles/checkSystem/plots/GMRES/Spectrum_Dt_1e-1}
\caption{
\textit{Left:} The linear iterations necessary to
obtain convergence decrease as we increase the number of particles  $N$.
\textit{Right:} The linear iterations necessary to
obtain convergence depends on the spectrum of the eigenvalues of the Jacobian. The clustering radius $\rho$ decreases as we increase the number of particles  $N$.
} 
\label{fig:Spectrum_Dt_1e-1}
\end{figure}

To get a better understanding of the convergence behaviour of the GMRES-method, we refer to the following theorem.
\begin{Theorem}[Proof in \cite{broyden2004krylov} and \cite{campbell96}]
Let $ \mathbf{A} \in \mathbb{C}^{n\times n}$ and assume it has $m$ distinct eigenvalues $\lambda_i$ clustered so that for some $\rho > 0 $ and $z \in \mathcal{C}$ , $|\lambda_i - z| \leq \rho$, together with $p$ eigenvalues $\mu_i$ (outliers), each of multiplicity $m_j$ and each lying outside the cluster $|\mu_i - z| > \rho$. Let $d= \sum_{j=1}^p m_j$  so that $n= d+m$. Then, if GMRES is applied to a set of equations with such a matrix of coefficients, it follows, for $k>0$,

\begin{equation}
\| \mathbf{r}_{d+k+1} \| \leq C  \left( \frac{\rho}{|z|} \right)^k  \| \mathbf{r}_1 \| ,
\end{equation}
where $C$ is a constant not depending on $k$.
\end{Theorem}

The theorem states that GMRES may behave as if it were a two-stage process. In the first phase the terms due to the outliers are eliminated before rapid linear convergence sets in for the second phase. The rate of convergence in this second stage is determined by $\frac{\rho}{\lvert z\lvert}$, where $z$ represents the center of the cluster and  $\rho$ its radius. The closer $z$ is to the origin, the slower the expected convergence of the algoritm. If the radius of the cluster $\rho$ is small, we expect a faster convergence of the algorithm. This is observed in figure \ref{fig:Spectrum_Dt_1e-1}.

### 2016-04-15 Documentation/coarse_time_stepper.tex
bifurcation diagramma

--- 2015-10-27 Documentation/fokkerplanck.tex
@@
-\documentclass[]{article}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[utf8]{inputenc} 
-\usepackage{subcaption}
-\usepackage{caption}
-\usepackage{hyperref} 
-%opening
-\title{}
-\author{}
+To simulate the time evolution of  the density $\U(t)$, we construct  a coarse time stepper $\cts$ which allows the performance of time-steps at the macroscopic level, using only the stochastic simulation of the position vectors of the $N$ particles at the microscopic level, %$\mathbf{X}(t+ n \Delta t)$,
+generated by eq. \eqref{Euler-Mar}.
 
-\newcommand{\R}{\ensuremath{\mathbb{R}}} % commando zonder argumenten
-\newcommand{\C}{\ensuremath{\mathbb{C}}}
-\newcommand{\N}{\ensuremath{\mathbb{N}}}
-\newcommand{\norm}[1]{\left\|#1\right\|} % commando's met argumenten
-\newcommand{\pa}[2]{\frac{\partial #1}{\partial #2}}
-\newcommand{\ppa}[2]{\frac{\partial^2 #1}{\partial #2^2}}
-\newcommand{\dd}{\ensuremath{\mathrm{d}}}
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
-\end{abstract}
-
-We compare solutions of partial differential equations of advection-diffusion-type
+To achieve this, we will define two operators (lifting in subsection \ref{section:lifting} and
+restriction in subsection \ref{section:restriction}) that relate the microscopic and macroscopic levels of description.
+Once these lifting $\mathcal{L}$ and restriction  operators $\mathcal{R}$ have been constructed, a coarse time-stepper $\cts$ 
+to evolve the macroscopic state $\U$ over a time interval of length $n \Delta t$ is constructed as a three-step-procedure (lift–evolve–restrict):
 
 \begin{equation}
-\label{fokkerplanck}
-\pa{\rho(x,t)}{t} + \pa{(a(x) \rho(x,t))}{x} = D  \ppa{\rho(x,t)}{x} 
+\U(t + n \Delta t) = \cts(\U) = (\mathcal{R} \circ \mathcal{E}(n \Delta t) \circ  \mathcal{L}(\mathbf{\omega})  ) (\U(t))
 \end{equation}
+where $\mathcal{E}(n \Delta t)(\U(t))$ is the simulation of the SDE for $N$ particles over $n$ timesteps.
 
-generated in two ways:
-\begin{itemize}
-\item By explicitly solving eq.
-\eqref{fokkerplanck} using the discretization scheme\footnote{First-order upwind scheme for the advection part combined with the Forward-Time Central-Space-method for the difussion part}
+\subsubsection{Lifting: $\U \rightarrow \mathbf{X} $} \label{section:lifting}
+Given the density \U, we need to sample a position vector $X_i$ for every particle $i \leq N$.
+We compute $X$ from a $N$-dimensional vector $\bf{U}$  with uniform random elements $U_i \in [0,1]$ such that $\U(X_i)=U_i$, using the inverse transformation method.
+The particle does not only gets an inital position, but also a seed for generating random steps in the simulation.
+
+\subsubsection{Restriction:  $\mathbf{X} \rightarrow \U $}
+\label{section:restriction}
+The restriction operator $\mathcal{R}: \mathbb{Q}^N \rightarrow \mathbb{Q}^k$ maps the microscopic state $\mathbf{X}$ (determined by the position vectors of $N$ particles) to a density \U, discretisized in $k$ bins.
+This is done by counting the number of particles in every bin $\Delta_j$ for $1\leq j \leq k$: %= x_{j+\frac{1}{2}}-x_{j-\frac{1}{2}}$.
+
 \begin{equation}
-\rho_i^{n+1} = \rho_i^n + \Delta t \left( \frac{D}{{\Delta x}^2} \left( \rho_{i+1}^{n} - 2 \rho_i^n + \rho_{i-1}^n \right)  - \frac{a(x)}{\Delta x} (\rho_i^n - \rho_{i-1}^n) \right)
+\frac{1}{N} \sum_{i=1}^{N}  w^i \cdot \chi_{\Delta_j}(X^i) = \rho_j  \label{restriction}
 \end{equation}
-for the value of the probabilty density $\rho$ at position $i \Delta x$ and time $(n+1) \Delta t$.
-\item By simulating an ensemble of $N$ particles evolving according to the corresponding stochastic differential equation  
+with 
 \begin{equation}
-    \dd X = a(X) \dd t + \sqrt{2D} \cdot \dd{W}.
+\chi_{\Delta_j}(X) = \begin{cases}
+  1 & \mbox{if } X \in \Delta_j, \\
+  0 & \mbox{if } X \notin \Delta_j. \\
+\end{cases}
 \end{equation}
- The position $X^{n+1}$ at time $ (n+1) \Delta t$ of each  particle is simulated using the  Euler-Maruyama scheme
+%\begin{equation}
+% \frac{1}{k}\sum_{j=1}^{k} \rho_j =  \frac{1}{N} \sum_{i=1}^{N}  w_i x_i
+%\end{equation}
+and setting all weights $w_i =1 $ for $1\leq i \leq N$.
+
+The reason why we explicitly introduced these weights in the restriction operator will be clarified in section \ref{Jv_approx} where we will need to evaluate  the coarse time stepper  $\cts (\U + \varepsilon \mathbf{v})$, now applied to the density shifted with a certain perturbation $\varepsilon \mathbf{v}$.
+%We explicitly introduced these weights in the restriction operator \eqref{restriction} because they can help to reduce the variance in the jacobian-vector-products.
+In the finite difference approximation we need to evaluate $\cts (\U + \varepsilon \mathbf{v})$, the coarse time stepper applied to the density shifted with a certain perturbation.
+%For the weighted restriction operator, we choose different weights $w_i$ for every particle.
+(we only use this for calculating the jacobian-vector-produ cts.
+(Why?
+Well, the weights are just one if we dont use a perturbation)f
+To evaluate the perturbated restriction-operator we will use the weights $ w^i_{\varepsilon}$, determined such that
 \begin{equation}
-   X^{n + 1} = X^{n} + a(X^n) \Delta t +  \sqrt{2D \Delta t}\cdot \xi^n 
+\frac{1}{N} \sum_{i=1}^{N}  w^i_{\varepsilon} \cdot \chi_{\Delta_j} (X^i) = \rho_j + \varepsilon v_j . \label{restriction_eps}
 \end{equation}
-with $\xi^n  \sim \mathcal{N} (0,1)$
-\end{itemize}
-
-\section{}
-
-\end{document}
+We do this by computing the weight per bin as $ w^j_{\varepsilon} = 1+ \frac{\varepsilon v_j}{\rho_j}$ and assign this value to each particle in $\Delta_j$.
+So, small perturbations on the density lead to small perturbations in the weights.
+The advantage of this weighted restriction operator lies in the possibility to use the same realizations $\mathbf{X}$ in the unperturbated \eqref{restriction} and the perturbated  \eqref{restriction_eps} restriction-operator.

### 2016-04-15 Documentation/draft.tex
bifurcation diagramma

@@
 \usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{amssymb}
@@
 %opening
-\title{}
+\title{Variance reduction in coarse bifurcation analysis of stochastic models}
 \author{Pieter Van Nuffel}
@@
 \newcommand{\U}{\ensuremath{\boldsymbol{\rho}}}
-\newcommand{\cts}{\ensuremath{\boldsymbol{\Phi}_T}} %Coarse time step
+\newcommand{\cts}{\ensuremath{\boldsymbol{\Phi}^N_T}} %Coarse time step
 \newcommand{\V}{\ensuremath{\mathbf{v}}} 
@@
 
+\theoremstyle{definition}
+\newtheorem{Theorem}{Theorem}
+
 \begin{document}
 
-\maketitle
+\maketitle{}
 
@@
 
+ As a toy problem we use a model where the closed model is exact.
+In general, we are interested in performing a bifurcation analysis for a model for which at which an exact, closed model at the macorscopic level is not available.
 \subsection{Model problem}
@@
 \section{Method}
+\subsection{Coarse time stepper}
+\input{coarse_time_stepper.tex}
 
@@
 %where $\mathbf{F}$ is nonlinear function from $\R^N$ to  $\R^N$. 
-To find the steady state $\U_*$, we apply Newton's method to eq. \ref{non-linear_system} . Starting from an initial state $\U^0$, we iterate
-
+To find the steady state $\U_*$, we apply Newton's method to eq.
+\ref{non-linear_system}.
+Starting from an initial state $\U^0$, we iterate %in two  steps, because the inverse of the the inverse of the Jacobian is computationally expensive)
 \begin{eqnarray}
@@
 J(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \\
-&\approx & \frac{\cts (\U, \boldsymbol{\omega_1} )  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cts (\U, \boldsymbol{\omega_2}) }{\varepsilon} \nonumber
-.
+&\approx & \frac{\cts (\U, \boldsymbol{\omega_1} )  + \varepsilon J(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cts (\U, \boldsymbol{\omega_2}) }{\varepsilon} \nonumber
 \end{eqnarray}
 
-\subsection{Variance reduced Jacobian-vector product}
+\input{GMRES}
+
+\subsection{Variance reduced Jacobian-vector products}
+
+If we use the solution of the PDE, eq.
+\eqref{pde_discretization}, the time stepper is deterministic and the calculation of the Jacobian-vector products is straightforward.
+If we use the solutions of the SDE however, we have to deal with numerical noise in evaluating eq.
+\ref{Jv_approx}.
+Because the coarse time-stepper is stochastic, repeating $\cts$ with two sets of random numbers ($\boldsymbol{\omega_1}$ and $ \boldsymbol{\omega_2}$)  will give different results.
+For $ \varepsilon \ll 1$ this will result in an $\mathcal{O}(1/(\varepsilon^2 N))$ variance.
+Consequently the variance on the Jacobian-vector-products will grow unboundedly as $\varepsilon$ tends to zero and $J(\cts) \cdot \mathbf{v}$ completely loses the structure of the perturbation \V.
+
+This numerical noise can be reduced by using the same random numbers $\boldsymbol{\omega}$ for the unperturbed and perturbed simulations.
+If we apply the weighted restriction operator \eqref{restriction_eps},  we get the same microscopic realizations in the lifting step - the only difference is in the computation of the weights.
+As such,  we impose $\boldsymbol{\omega_1} = \boldsymbol{\omega_2}$ in eq.
+\eqref{Jv_approx} and consequently the variance of $J(\cts) \cdot \mathbf{v}$ is bounded and of  $\mathcal{O}(1/ N)$.
+In the limit of infinitely many particles, the result will converge to the exact Jacobian-vector product.
+For finite values of $N$, there will be noise in the Jacobian-vector product as a result of the random selection of a subset of all possible realizations.
+The presented procedure only prevents noise blowup that would arise if a different selection of realizations were considered for the perturbed and unperturbed coarse time-step.
 
@@
 \subsection{Variance}
+
+Fig.
+\ref{Var_N} shows that the variance on the stochastic solution for the Jacobian-vector-product converges to zero with $\mathcal{O}(1/ N)$ and that it does not depend on the value of $\varepsilon$.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\linewidth]{../Problems/Particles/checkSystem/plots/Var_N_eps_nw}
+\caption[Effect of variance reduction]{In the case of unweighted restriction, the variance on the stochastic solution of the Jacobian-vector-product becomes unbounded as we decrease the perturbation size $\epsilon$.
+By using weights in the restriction step, the variance does not longer depend on the value of $\varepsilon$ and   converges to zero with $\mathcal{O}(1/ N)$.}
+\label{Var_N}
+\end{figure}
+
+%\input{GMRES.tex}
+
+The convergence of the density to the real solution, strongly depends on the number of particles $N$ and on the choise of $\epsilon$ in the GMRES-method.
+
+\subsection{}
+
 \subsection{bias}
@@
 
+\subsubsection{Nota's}
+Tijdens elke continueringsstap, wordt een nieuw punt gecreëerd dat initieel de dichtheid \U krijgt van het vorig punt en met een parameter (D) die horizontaal word geupdate.
+Dan wordt de norm van het residu berekend $\norm{ \U(t) - \U(t+\Delta t)} $, waarin de dichtheid op de volgende tijdsstap verkregen wordt door een simulatie van het systeem met de nieuwe parameter.
+De grootte van het residu is dus afhankelijk van de stapgrootte in de bifurcatieparamter.
+De newtontolerantie is de norm van het residu.
+Als de newton-tolerantie te groot gekozen wordt (in vergelijking met de stapgrootte van het residu), dan wordt er geen Newton-iteratie uitgevoerd.
+Als de Newton-tolerantie evenwel te klein gekozen wordt, dan bereikt de Newton-solver mogelijk geen convergentie ten gevolge van de ruis.
+De Newton-tolerantie moet dus weloverwogen gekozen worden.
+
+\subsection{Results for the pde}
+Ook hier zien we dat de keuze van de newtontolerantie belangrijk is (te groot = kans op dezelfde toestanden , te klein= failed iterations).
+Dit kan echter opgevangen worden door de oplossing van dx nauwkeuriger te maken door de tolerantie van de GMRES te verfijnen (het probleem was dat er dx=0 oplossingen werden teruggegeven)
+
+\begin{figure}
+\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
+\caption{  Bifurcation diagram of the steady states calculated with the Newton-Krylov-Solver for the PDE.
+$\Delta t = 10^{-4}, \Delta T = 10^{-2}, \Delta x = 10^{-2}, \Delta D = 0.01, \epsilon_{GMRES}=10^{-5},  \delta_{Newton} = 10^{-7}, \epsilon=10^{-5}$
+}
+\end{figure}
+
+%\begin{center}
+%\begin{table}
+%\caption{Parameter values}
+%  \begin{tabular} { | c  | c |}    \hline   
+%     %FTCS-scheme  &   Euler-Maruyama-scheme  \\  
+%   \textit{ {Discretization parameters}}&    &  SDE    \\ \hline
+%       $\Delta t $  & $10^{-4}$ \\ 
+%       $\Delta T $ &  $10^{-2}$ \\ 
+%      $ \Delta x &  $10^{-2}$ \\ 
+%       $ \Delta D$  $  &  $ 0.01 $ \\ \hline
+%  \end{tabular}
+%\end{table}
+%\end{center}
+
 \subsubsection{Calculating fixed points}
@@
 
+\section{Newton convergence}
+
 \bibliographystyle{plain}

### 2016-04-15 Continuer/Continuer.py

@@ -23,11 +23,11 @@ class Continuer(object):
             of points obtained through continuation.
 
         """
-        for i in range(len(points)):
-            points[i].param['free']+=params['free']   
-            points[i].param['extra_condition']+=\
-                    params['extra_condition']    
-            points[i].extra_condition+=params['extra_condition']    
+#        for i in range(len(points)):
+           # points[i].param['free']+=params['free']   
+        #    points[i].param['extra_condition']+=\
+             #       params['extra_condition']    
+   #         points[i].extra_condition+=params['extra_condition']    
         self.points = points
         self.bound_condition = params['bound_condition']
         if params == None:
@@ -56,7 +56,7 @@ class Continuer(object):
         
         plotx = {}
         plotx['field'] = 'lambd'
-        plotx['func'] = 0
+        plotx['func'] = 1 #second element of lambd (usually Difussion coefficient)
         param['plotx']=plotx
         ploty = {}
         ploty['field'] = 'u'
@@ -64,6 +64,9 @@ class Continuer(object):
         param['ploty']=ploty
 
         param['growth_factor']=1.2
+        
+        param['cont_step'] = 0.2
+
         param['print']='short'
         param['bound_condition']=[]
         return param
@@ -100,10 +103,8 @@ class Continuer(object):
             l = len(self.points)
             prev_point = self.points[l-2]
             last_point = self.points[l-1]
-
             secant = last_point.axpy(-1,prev_point)
-            dist = scipy.linalg.norm(secant.u)+\
-                    scipy.linalg.norm(secant.lambd)
+            dist = scipy.linalg.norm(secant.u)+ scipy.linalg.norm(secant.lambd)
             if successful and not bound:  
                 # het vorige punt was successvol berekend
                 steplength = growth*dist
@@ -112,12 +113,11 @@ class Continuer(object):
             elif bound:
                 # we hebben de rand geraakt
                 # nu moeten we de stap verkleinen en zien of het zo gaat
-                steplength = steplength/2
-                
+                steplength = steplength/2           
             new_point = secant.axpy(-steplength/dist,last_point)
             new_point.secant = {}
-            new_point.secant['u']=secant.u
-            new_point.secant['lambd']=secant.lambd
+            new_point.secant['u']=secant.u              #Waarom wordt dit bijgehouden?
+            new_point.secant['lambd']=secant.lambd    #Waarom wordt dit bijgehouden?
             status = new_point.correct()
             if status == 0:
                 new_success = True
@@ -151,7 +151,57 @@ class Continuer(object):
[…]

### 2016-04-15 Point/Point.py

@@ -8,6 +8,9 @@ import Solver.Solver as Solver
 import Solver.LinearSolver as LinearSolver
 import System.System as System
 
+import matplotlib.pylab as plt
+
+
 class Point(scipy.sparse.linalg.LinearOperator):
 
     """ 
@@ -116,13 +119,35 @@ class Point(scipy.sparse.linalg.LinearOperator):
     def setState(self,u,lambd):
         self.u=u
         self.lambd=lambd
-    #    self.system.setState(u,lambd)  #need  to use this again for continuation
+        self.system.setState(u,lambd)  #need  to use this again for continuation       
+        #om onbegrijpelijke reden stond bovenstaande lijn initieel in commentaar waardoor het residu in Newtonsolver niet werd geupdated.
+        # this is meant to deal with preconditioner matrix that needs to be
+        # built if the state changes.
+#        if not self.precond == None:
+#            self.precond.new_build()
+#        self.solver.new_build()
+        
+#    def updateLambd(self, lambd):
+#        # self.u=u
+#        # self.lambd=lambd
+#        self.system.setState(u,lambd)  #need  to use this again for continuation
+#        # this is meant to deal with preconditioner matrix that needs to be
+#        # built if the state changes.
+#        if not self.precond == None:
+#             self.precond.new_build()
+#        self.solver.new_build()
+    def updateState(self):
+        # self.u=u
+        # self.lambd=lambd
+        u, lambd = self.getState()
+        self.system.setState(u,lambd)  #need  to use this again for continuation
         # this is meant to deal with preconditioner matrix that needs to be
         # built if the state changes.
         if not self.precond == None:
-            self.precond.new_build()
+             self.precond.new_build()
         self.solver.new_build()
-    
+
+
     def getState(self):
         """
         Returns (u,lambd)
@@ -162,6 +187,7 @@ class Point(scipy.sparse.linalg.LinearOperator):
         lambd[self.param['free']]=x[n:n+l]
         lambd[self.param['artificial']]=x[n+l:]
         self.setState(u,lambd)
+        self.updateState()
     
     def getCurrentGuess(self):
         """
@@ -174,7 +200,7 @@ class Point(scipy.sparse.linalg.LinearOperator):
 
         The vector "x" has a part "u"  that describes the
         current solution, "free" for values of the free parameters 
-        and a part "art" containing the values of thhe artificial
[…]

### 2016-04-15 Problems/Particles/checkSystem/check_sde.py

@@ -27,7 +27,7 @@ a = 1  #advection coefficient
 D =  1./2. #difussion coefficient
 lambd = scipy.array([a,D])
 Dt = 0.01 
-N = 10000
+N = 1000
 eps = 1e-5
 
 # discretization parameters
@@ -41,7 +41,6 @@ grid = scipy.arange(xL+dx/2.,xR,dx)
 rho = scipy.ones_like(grid)
 rho = rho/(sum(rho)*dx)
 #print ("rho: ", rho)
-
 
 v=scipy.zeros_like(grid)
 for j in range(len(grid)): 
@@ -64,7 +63,7 @@ param['Dt'] = Dt
 param['dt']=dt
 
 param['eps']= eps
-M=500 #number of monte carlo steps  
+M=100 #number of monte carlo steps  
 
 E_rho =scipy.zeros_like(rho)
 E_Jv =scipy.zeros_like(rho)

### 2016-04-15 Problems/Particles/checkSystem/check_sde_var_eps.py

@@ -1,0 +1,164 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+#import matplotlib.pylab as plt
+
+sys.path.append("../system/")
+sys.path.append("../lifting/")
+sys.path.append("../restriction/")
+sys.path.append("..")
+sys.path.append("../..")
+
+import pde
+import histogram
+import particles
+import inv_transform
+
+if __name__=="__main__":
+    D = 1./2.
+    Dt = 1e-2
+    seed = 16
+    xL = -1.7
+    xR = 1.7
+    
+    a = 1
+    zeta = 0.
+    alpha=1
+    beta = -1
+    lambd = scipy.array([a,D,alpha,beta,zeta])
+  
+    #SDE
+    t1 = time.time()
+    
+    print 'Start solving sde'
+    dx = 1e-2
+    dt = 1e-4 # ti
+    N=1000
+    Nlarge= N
+    Nsmall =N
+    
+    
+    grid = scipy.arange(xL+dx/2.,xR,dx)
+    rho = scipy.ones_like(grid)
+    rho = rho/(sum(rho)*dx)
+    
+    
+
+    v=scipy.zeros_like(grid)
+    for j in range(len(grid)): 
+        v[j]= np.sin(j*2*np.pi/len(grid))
+
+     
+    
+  #  sampler_param = inv_transform.Sampler.getDefaultParameters()
+  #  sampler_param['seed']=0
+  #  lifting =  inv_transform.Sampler(sampler_param)
+    
+    M=100 #number of monte carlo steps                
+
+    restriction = histogram.Histogram()
+
+    param=particles.Particles.getDefaultParameters()
+    param['Nlarge']=N
+    param['Nsmall']=N    
+    param['Dt'] = Dt 
+    param['dt']=dt
+#    param['eps']=1e-5
+   # fp_sde = particles.Particles(lifting,restriction,rho,grid, lambd, param=param)       
+
+    Nlist = scipy.array([1000, 2000, 4000, 8000, 16000, 32000, 64000, 128000, 256000])
+ #   Nlist = scipy.array([1000, 2000, 4000])  #, 800, 1600, 3200])
+    eps_list_exponents=[4,5,6,7]
[…]

### 2016-04-15 Problems/Particles/checkSystem/post_processing_M.py

@@ -1,0 +1,296 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+import matplotlib.pylab as plt
+from matplotlib.path import Path #to draw trriangles indicating order of convergence
+import matplotlib.patches as patches #to draw trriangles indicating order of convergence
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+import matplotlib.pylab as plt
+
+
+if __name__=="__main__":
+   
+    Mlist= [ 20, 40, 60, 80, 100, 120, 140, 160, 180, 200]
+    
+    Nlist = scipy.array([1000, 2000,4000,8000,16000,32000,64000, 128000, 256000])
+    Nlist_inv = 1.0/Nlist
+    rho_Dt_fp= np.loadtxt('data/25-11-Rho_pde_dx_e-4_eps-5.out')   #,  np.loadtxt('25-11-Rho_pde_dx_e-4_eps-5.out') ,  
+    #np.loadtxt('25-11-Rho_pde_dx_e-4_eps-6.out') ,  np.loadtxt('25-11-Rho_pde_dx_e-4_eps-7.out') ]  #these data should be the same
+#    Jv_pde= np.loadtxt('data/25-11-Jv_pde_dx_e-4_eps-5.out') 
+   # Jv_pde_dxbig= np.loadtxt('data/Jv_pde_dx_5e-4_vsin.out') 
+    Jv_pde_dxbig= np.loadtxt('Jv_pde_coarse.out')
+  #  Jv_pde = Jv_pde_list[1]
+    
+    #SDE       
+    
+#    sde_rho_sq = scipy.zeros((len(Mlist), len(Nlist)))
+#    for M_factor in range(0,len(Mlist)):
+#        m= (M_factor+1)*20
+#        sde_rho_sq[M_factor] = np.loadtxt('data/8-12-rho_sq_m%d.out' %m)
+#    E_rho =scipy.zeros(((len(Mlist), len(Nlist), 340)))  # [m x N_i x dx] = [5 x 8 x 3400]
+#    for M_factor in range(0,5):
+#        m= (M_factor+1)*20
+#        E_rho[M_factor] = np.loadtxt('data/8-12-E_rho_m%d.out' %m)
+#    sq_E_rho = scipy.zeros((len(Mlist), len(Nlist))) 
[…]

### 2016-04-15 Problems/Particles/checkSystem/post_processing_var_eps.py

@@ -1,0 +1,116 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+import matplotlib.pylab as plt
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+import matplotlib.pylab as plt
+
+
+if __name__=="__main__":
+    D = 1./2
+    Dt = 1e-2    
+    a = 1
+    zeta = 0.
+    alpha=1
+    beta = -1
+    lambd = scipy.array([a,D,alpha,beta,zeta])
+   # param['eps'] = 1e-5
+    #param['Dt'] = Dt
+    #param['dt']=dt         
+    eps_list_exponents=[4,5, 6,7]
+    eps_list= [1e-4, 1e-5, 1e-6, 1e-7]
+
+
+    #SDE       
+   # sde_rho_sq = np.loadtxt('data/25-11-rho_sq_eps-5.out')
+   # E_rho = np.loadtxt('data/25-11-E_rho_eps-5.out')
+   # sq_E_rho  =np.loadtxt('data/25-11-sq_E_rho_eps-5.out')
+    
+    sde_Jv_sq = [ np.loadtxt('nw-Jv_sq_eps-4.out'), np.loadtxt('nw-Jv_sq_eps-5.out'),
+           np.loadtxt('nw-Jv_sq_eps-6.out'), np.loadtxt('nw-Jv_sq_eps-7.out') ] 
+#    E_Jv = [np.loadtxt('data/eps_fin-E_Jv_eps-4.out'),  np.loadtxt('data/eps_fin-E_Jv_eps-5.out'), 
+#            np.loadtxt('data/eps_fin-E_Jv_eps-6.out'), np.loadtxt('data/eps_fin-E_Jv_eps-7.out') ]    # [eps_i x N_i x dx] = [4 x 7 x 3400]
+    sq_E_Jv =  [np.loadtxt('nw-sq_E_Jv_eps-4.out') , np.loadtxt('nw-sq_E_Jv_eps-5.out' ) ,
+                 np.loadtxt('nw-sq_E_Jv_eps-6.out' ),np.loadtxt('nw-sq_E_Jv_eps-7.out' )]
+                 
+                 
[…]

### 2016-04-15 Problems/Particles/lifting/inv_transform.py

@@ -17,7 +17,7 @@ class Sampler(object):
     def lift(self,rho,grid,N):
         cum,edges = self.rho2cum(rho,grid)
         y = self.rand.uniform(size=N)
-        print "sampling density with seed : ", self.cur_seed, "first sample: ", y[0]
+     #   print "sampling density with seed : ", self.cur_seed, "first sample: ", y[0]
         return self.inv_cum(y,cum,edges)
     
     def rho2cum(self,rho,grid):

### 2016-04-15 Problems/Particles/system/particles.py

@@ -5,7 +5,7 @@ sys.path.append('../')   #added to resolve importerror of nephew-directories
 sys.path.append('../../')
 sys.path.append('../../../')
 
-
+from scipy.linalg import norm
 
 
 import System.TimestepperSystem as TSystem
@@ -62,7 +62,7 @@ class Particles(TSystem.TimestepperSystem):
         self.lifting.seed(self.lifting.param['seed'])
         #self.seed(self.param['seed'])
         x = self.lift(rho)
-        print "nb particles : ", len(x)
+       # print "nb particles : ", len(x)
         x_Dt = self.simulate(x,lambd)
         return self.restrict(x_Dt)
     
@@ -87,8 +87,8 @@ class Particles(TSystem.TimestepperSystem):
             tcur += dt
             # the random number
             dW=self.rand.normal(loc=0.,scale=scipy.sqrt(2*D*dt),size=N)
-          #  if tcur == dt:    #only print random number for first time step
-          #      print 'dW =',  dW            
+#            if tcur == dt:    #only print random number for first time step
+#                print 'dW =',  dW            
             
             # the process
             drift_term = a * drift(x)
@@ -105,4 +105,15 @@ class Particles(TSystem.TimestepperSystem):
         dx = self.grid[1]-self.grid[0]
         N = len(self.u)
         return self.precond.computeJacobian(N,dx)
-    
+        
+
+#    def applyJacobian(self,v):  #In Timestepperystem
+#
+#        eps = self.param['eps']
+#        u_eps = self.u + eps * v/norm(v)
+#        print self.seed
+#        u_eps_Dt = self.integrate(u_eps,self.lambd)
+#        print self.seed
+#        result = (u_eps_Dt - self.u_Dt)/eps*norm(v)
+#        return result   
+#    

### 2016-04-15 Problems/WeightedParticles/checkSystem/calculate_pde_coarse.py

@@ -1,0 +1,88 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+sys.path.append("../system/")
+sys.path.append("../lifting/")
+sys.path.append("../restriction/")
+sys.path.append("..")
+sys.path.append("../..")
+
+
+import pde
+import histogram
+import particles
+import inv_transform_sampling as inv_transform
+
+import matplotlib.pylab as plt
+
+if __name__=="__main__":
+    D = 1./2.
+    Dt = 1e-2
+    seed = 16
+    # discretization parameters
+   #h=2e-2 # kde mesh width 
+    dx = 1e-2
+    dt=1e-8
+#    dt = dx**2/(2*D)*r
+    print Dt/dt, " timesteps"
+#    r= (2.*D*dt)/(dx)**2
+#    print 'r= ', r
+#    if r>1: 
+#         print 'Stability condition not fulfilled, because r=',r #(see https://en.wikipedia.org/wiki/FTCS_scheme)
+#         sys.exit("Ending Script. Make sure that r<1") 
+    xL = -1.7
+    xR = 1.7
+    
+    a = 1
+    zeta = 0.
+    alpha=1
+    beta = -1
+    lambd = scipy.array([a,D,alpha,beta,zeta])
+    param=pde.FokkerPlanck.getDefaultParameters()
+    param['Dt'] = Dt
+    param['dt']=dt         
+ 
+    #PDE
+    t0 = time.time()
+    print 'Start solving pde'
+    
+    
+    #rlist= [0.0001, 0.0005, 0.001, 0.005, 0.01, 0.05, 0.1, 0.5, 1]
+   # dxlist= [0.001, 0.002, 0.005, 0.01, 0.02, 0.05, 0.1]
+
+    print 'dx= ', dx
+    r= (2.*D*dt)/(dx)**2
+    print 'r= ', r
+    if r>1: 
+        print 'Stability condition not fulfilled, because r=',r #(see https://en.wikipedia.org/wiki/FTCS_scheme)
+        #sys.exit("Ending Script. Make sure that r<1")       
+        
+    grid = scipy.arange(xL+dx/2.,xR,dx)
+    print len(grid), " discretisation steps"
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/check_bif_pde.py

@@ -1,0 +1,224 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+from pylab import *
+import os
+
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+import matplotlib.pylab as plt
+
+sys.path.append("../system/")
+sys.path.append("../lifting/")
+sys.path.append("../restriction/")
+sys.path.append("..")
+sys.path.append("../..")
+sys.path.append("../../..")
+
+import pde
+
+import Point.Point as Point
+import Solver.NewtonSolver as NewtonSolver
+import Solver.GMRESLinearSolver as GMRES
+import Solver.ImplicitEulerDirectLinearSolver as ImDirect
+import Utils.conditions.probability as probability
+import Continuer.Continuer as Continuer 
+
+
+
+
+if __name__=="__main__":
+   # D = 1./2.
+    sigma =1
+    D= 1#1/2.
+    Dt = 1e-2
+  #  Dtlist = [1e-5,2e-5, 5e-5, 1e-4, 2e-4, 5e-4, 1e-3, 2e-3, 5e-3 , 1e-2, 2e-2, 5e-2 , 1e-1, 2e-1, 5e-1, 1, 2, 5, 10]
+    Dtlist = [Dt]
+ #   Dtlist = [2e-3]
+    seed = 15
+    # discretization parameters
+   #h=2e-2 # kde mesh width 
+    dx = 1e-2
+#    dxlist= [1e-2,  5e-3, 1e-3, 5e-4] 
+  #  dxlist= [1e-2, 1e-3]
+    #dt = min(dxlist)**2/(2*D)
+    dt=5e-5
+    r= (2.*D*dt)/(dx)**2
+    print'r = ' ,  r
+
+   # r= (2.*D*dt)/(min(dxlist))**2
+    if r>1: 
+        print 'Stability condition not fulfilled, because r=',r #(see https://en.wikipedia.org/wiki/FTCS_scheme)
+        sys.exit("Ending Script. Make sure that r<1") 
+    xL = -1.7
+    xR = 1.7
+    
+    a = 0.2
+    zeta = 0.
+    alpha=1
+    beta = -1
+    lambd = scipy.array([a,D,alpha])
+    param=pde.FokkerPlanck.getDefaultParameters()
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/check_bif_sde.py

@@ -20,62 +20,57 @@ sys.path.append("../restriction/")
 sys.path.append("..")
 sys.path.append("../..")
 
-
 import histogram
 import particles
 import inv_transform_sampling as inv_transform
 
+import matplotlib.pylab as plt
+
 import Point.Point as Point
 import Solver.NewtonSolver as NewtonSolver
 import Solver.GMRESLinearSolver as GMRES
+import Solver.Richardson as Rich
 import Solver.ImplicitEulerDirectLinearSolver as ImDirect
 import Utils.conditions.probability as probability
 #import Continuer.Continuer as Continuer 
+import Continuer.Continuer as Continuer 
+
 
 if __name__=="__main__":
     print 'Start solving sde'
-    D = 1./2.
-    Dt = 100
-    seed = 16
+    D = 1
+    Dt = 1e-1  #Dt influences number of GMRES-iterations strongly
+    Dtlist = [ Dt]
+    #Dtlist = [ 5e-3, 1e-2, 2e-2, 5e-2, 1e-1, 2e-1, 5e-1, 1]
+    #Dtlist = [1e-5, 5e-5, 1e-4,  5e-4, 1e-3,  5e-3 , 1e-2, 5e-2 , 1e-1, 5e-1, 1]
+    seed = 1
     # discretization parameters
    #h=2e-2 # kde mesh width 
    # dt=1e-6
     xL = -1.7
     xR = 1.7
-    
-    a = 1
+    a = 0.2
     zeta = 0.
     alpha=1
     beta = -1
     lambd = scipy.array([a,D,alpha,beta,zeta])
-            
+       
     #SDE
     t1 = time.time()
     
     print 'Start solving sde'
     dx = 1e-2
-#    factor = int (dx/dx_fp)
-#    print "Discretisation for solving sde is ", factor, " times coarser than the discretisation for solving the pde"
-    dt = 1e-2 # ti
-    N=5000
-    Nlarge= N
-    Nsmall =N
+    dt = 1e-3 # ti
+#    N=5e5
+#    Nlarge= N 
+#    Nsmall =N
     
-    
-    grid = scipy.arange(xL+dx/2.,xR,dx)
+    grid = scipy.arange(xL+dx/2.  ,xR,dx)
     rho = scipy.ones_like(grid)
     rho = rho/(sum(rho)*dx)
-    
-    
-
-    v=scipy.zeros_like(grid)
-    for j in range(len(grid)): 
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/check_pde.py

@@ -28,13 +28,13 @@ import pde
 
 if __name__=="__main__":
     D = 1./2.
-    Dt = 1e-2
+    Dt = 2
     seed = 16
     # discretization parameters
    #h=2e-2 # kde mesh width 
     dx = 1e-3
 #    dxlist= [1e-2,  5e-3, 1e-3, 5e-4] 
-    dxlist= [1e-2, 1e-3]
+    dxlist= [1e-2]
     dt = min(dxlist)**2/(2*D)
     #r= (2.*D*dt)/(dx)**2
     r= (2.*D*dt)/(min(dxlist))**2
@@ -65,13 +65,15 @@ if __name__=="__main__":
         print "rho: ", rho
  
         fp_pde = pde.FokkerPlanck(rho,grid,pde.doublewell,lambd,param)
+        rho_Dt = fp_pde.u_Dt
 
         v=scipy.zeros_like(grid)
         v[200]=-1
         v[300]=1
     
         Jv_pde = fp_pde.applyJacobian(v)    
-        plt.plot(grid, Jv_pde)
+       # plt.plot(grid, Jv_pde)
+        plt.plot(grid, rho_Dt)
         plt.show()
         #rho_Dt_fp = fp_pde.u_Dt
         #plt.plot( grid, rho_Dt_fp)

### 2016-04-15 Problems/WeightedParticles/checkSystem/check_sde.py

@@ -12,7 +12,7 @@ import numpy as np
 from scipy.linalg import norm
 from scipy import sqrt
 
-#import matplotlib.pylab as plt
+import matplotlib.pylab as plt
 
 sys.path.append("../system/")
 sys.path.append("../lifting/")
@@ -33,9 +33,9 @@ import Utils.conditions.probability as probability
 #import Continuer.Continuer as Continuer 
 
 if __name__=="__main__":
-    print 'Start solving sde'
-    D = 1./2.
-    Dt = 100
+
+    D = 1
+    Dt = 2
     seed = 16
     # discretization parameters
    #h=2e-2 # kde mesh width 
@@ -57,7 +57,7 @@ if __name__=="__main__":
 #    factor = int (dx/dx_fp)
 #    print "Discretisation for solving sde is ", factor, " times coarser than the discretisation for solving the pde"
     dt = 1e-2 # ti
-    N=1000
+    N=5000000
     Nlarge= N
     Nsmall =N
     
@@ -66,16 +66,9 @@ if __name__=="__main__":
     rho = scipy.ones_like(grid)
     rho = rho/(sum(rho)*dx)
     
-    
-
     v=scipy.zeros_like(grid)
     for j in range(len(grid)): 
         v[j]= np.sin(j*2*np.pi/len(grid))
-
-#    v=scipy.zeros_like(grid)
-#    v[20]=1
-#    v[30]=-1
-     
     
   #  sampler_param = inv_transform.Sampler.getDefaultParameters()
   #  sampler_param['seed']=0
@@ -145,57 +138,58 @@ if __name__=="__main__":
                 rg_state = lifting.get_rg_state()  #We remember the internal state of the random generator to get new random number for sampling the new particles
                 
                 print "calculate .u_Dt"
-                rho_Dt = fp_sde.u_Dt 
+                rho_Dt_sde = fp_sde.u_Dt 
                 t0=time.time()
                 print "Simulation time for solving sde: " , t0-t1, " seconds"
                 #rho_fine = fp_sde.make_rho_fine(rho_Dt) 
              #   rho_sq[n] = rho_sq[n] + (norm(rho_Dt))**2
             #    E_rho[n] = E_rho[n] + rho_Dt           #matrix
         
+                plt.plot(rho_Dt_sde)
                 print "calculate JVs"
               #  Jv =fp_sde.applyJacobian(v)                                                 
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/check_sde_M.py

@@ -54,7 +54,7 @@ if __name__=="__main__":
     t1 = time.time()
     
     print 'Start solving sde'
-    dx = 1e-1
+    dx = 1e-2
 
     dt = 1e-4 # ti
     N=1000
@@ -96,10 +96,10 @@ if __name__=="__main__":
    # fp_sde = particles.Particles(lifting,restriction,rho,grid, lambd, param=param)       
 
    # Nlist = scipy.array([1000, 2000, 4000, 8000, 16000, 32000, 64000, 128000])
-    Nlist = scipy.array([1000]) #, 800, 1600, 3200])
+    Nlist = scipy.array([10, 100, 1000, 10000]) #, 800, 1600, 3200])
     
     param['eps']= 1e-5 
-    M=100 #number of monte carlo steps  
+    M=10 #number of monte carlo steps  
     
     nN= len(Nlist)
 
@@ -146,9 +146,9 @@ if __name__=="__main__":
                 E_rho_sum[n] = E_rho_sum[n] + rho_Dt           #matrix
         
                 print "calculate JVs"
-                Jv =fp_sde.applyJacobian(v)                                                 
-                Jv_sq_sum[n] = Jv_sq_sum[n] + (norm(Jv))**2
-                E_Jv_sum[n] = E_Jv_sum[n] + Jv 
+              #  Jv =fp_sde.applyJacobian(v)                                                 
+          #      Jv_sq_sum[n] = Jv_sq_sum[n] + (norm(Jv))**2
+         #       E_Jv_sum[n] = E_Jv_sum[n] + Jv 
                 
                 x_prev_sim = fp_sde.x_Dt
               #  print 'len x-vector= ' , len( fp_sde.x_Dt) 
@@ -169,14 +169,14 @@ if __name__=="__main__":
             E_rho[n] = E_rho_sum[n]/Mtot
             sq_E_rho[n]= (norm(E_rho[n]))**2
             
-            Jv_sq[n]=  Jv_sq_sum[n]/Mtot
-            E_Jv[n] = E_Jv_sum[n]/Mtot
-            sq_E_Jv[n]= (norm(E_Jv[n]))**2
+       #     Jv_sq[n]=  Jv_sq_sum[n]/Mtot
+       #     E_Jv[n] = E_Jv_sum[n]/Mtot
+       #     sq_E_Jv[n]= (norm(E_Jv[n]))**2
             
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/post_processing.py

@@ -18,160 +18,28 @@ import matplotlib.pylab as plt
 
 if __name__=="__main__":
 
-
-
-    rho_Dt_fp= np.loadtxt('data/obs/25-11-Rho_pde_dx_e-4_eps-5.out') 
-    Jv_pde= np.loadtxt('data/obs/25-11-Jv_pde_dx_e-4_eps-5.out')
+    residual_directsim = np.loadtxt('Newton/residual_sim_N1e6.out')/sqrt(340) 
+    residual1 = np.loadtxt('Newton/residual1_N1e6.out') 
+    residual2 = np.loadtxt('Newton/residual2_N1e6.out') 
+    residual3 = np.loadtxt('Newton/residual3_N1e6.out')  
     
-    sde_rho_sq = np.loadtxt('data/obs/25-11-rho_sq_eps-5.out')
-    E_rho = np.loadtxt('data/obs/25-11-E_rho_eps-5.out')
-    sq_E_rho  =np.loadtxt('data/obs/25-11-sq_E_rho_eps-5.out')
-    
-    sde_Jv_sq = np.loadtxt('data/obs/25-11-Jv_sq_eps-5.out')
-    E_Jv = np.loadtxt('data/obs/25-11-E_Jv_eps-5.out')   #lots of data, not needed when only interested in the variance
-    sq_E_Jv =  np.loadtxt('data/obs/25-11-sq_E_Jv_eps-5.out' )
-    
-    
-
-            
-    #SDE
-    rho_Dt_sde = E_rho[-1]
-    Jv_sde = E_Jv[-1]
-
-    
-    resize_factor = int (len(rho_Dt_fp)/len(rho_Dt_sde))
-  #  resize_factor = 10
-    print "Discretisation for solving sde is ",  resize_factor , " times coarser than the discretisation for solving the pde"
-    
-    rho_coarse = scipy.zeros(len(rho_Dt_sde))
-    for i in range (0,len(rho_coarse)):
-        bin_av = 0
-        for j in range (0,  resize_factor ):
-            bin_av = bin_av+ rho_Dt_fp[i*resize_factor+j]
-        rho_coarse[i] = bin_av/resize_factor
-        
-    bins =  len(Jv_sde)
-    Jv_coarse = scipy.zeros(bins)
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/post_processing_GMRES.py

@@ -1,0 +1,69 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+import matplotlib.pylab as plt
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+import matplotlib.pylab as plt
+
+
+if __name__=="__main__":
+
+
+    res_e3= np.loadtxt('GMRES/res_Dt_e-2_builtin_N1000.out') 
+    res_e4= np.loadtxt('GMRES/res_Dt_e-2_builtin_N10000.out') 
+    res_e5= np.loadtxt('GMRES/res_Dt_e-2_builtin_N100000.out') 
+    res_e6= np.loadtxt('GMRES/res_Dt_e-2_builtin_N1000000.out') 
+
+    Nlist = scipy.array([10**3, 10**4, 10**5, 10**6])
+       
+    plot_res= plt.plot(res_e3, 'yo' , label= r'$N=10^3$', markersize =5)# % Nlist[0])  #gives the smae result, as expected
+    plot_res= plt.plot(res_e4, 'bo' ,label= r'$N=10^4$', markersize =5) #)% Nlist[1]) 
+    plot_res= plt.plot(res_e5, 'ro' ,label= r'$N=10^5$' , markersize =5)#% Nlist[2]) 
+    plot_res= plt.plot(res_e6, 'go' , label= r'$N=10^6$', markersize =5) #% Nlist[3]) 
+        
+    plt.ylabel('GMRES residual' , fontsize = 11)
+    plt.xlabel('iterations', fontsize = 11)
+    plt.legend([plot_res], loc='best') 
+    plt.legend(bbox_to_anchor=(1, 1), numpoints = 1 )    
+    
+     #   plt.xscale(log_flag)    res_e3= np.loadtxt('GMRES/res_N10000.out') 
+    
+    plt.yscale('log')
+    plt.savefig("plots/GMRES/GMRES_N_Dt_1e-2_builtin.pdf")
+       
+    
+    plt.show()
+    
+    
+    
+    
+    spectrum_e3= np.loadtxt('GMRES/spectrum_Dt_e-2_builtin_N1000.out').view(complex) 
+    spectrum_e4= np.loadtxt('GMRES/spectrum_Dt_e-2_builtin_N10000.out').view(complex) 
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/post_processing_M.py

@@ -22,13 +22,14 @@ if __name__=="__main__":
    
     Mlist= [ 20, 40, 60, 80, 100, 120, 140, 160, 180, 200]
     
-   # Nlist = scipy.array([1000, 2000,4000,8000,16000,32000,64000]) #,128000])
     Nlist = scipy.array([1000, 2000,4000,8000,16000,32000,64000, 128000, 256000])
     Nlist_inv = 1.0/Nlist
     rho_Dt_fp= np.loadtxt('data/25-11-Rho_pde_dx_e-4_eps-5.out')   #,  np.loadtxt('25-11-Rho_pde_dx_e-4_eps-5.out') ,  
     #np.loadtxt('25-11-Rho_pde_dx_e-4_eps-6.out') ,  np.loadtxt('25-11-Rho_pde_dx_e-4_eps-7.out') ]  #these data should be the same
-    Jv_pde= np.loadtxt('data/25-11-Jv_pde_dx_e-4_eps-5.out') 
-    Jv_pde_dxbig= np.loadtxt('data/Jv_pde_dx_5e-4_vsin.out') 
+#    Jv_pde= np.loadtxt('data/25-11-Jv_pde_dx_e-4_eps-5.out') 
+   # Jv_pde_dxbig= np.loadtxt('data/Jv_pde_dx_5e-4_vsin.out') 
+    Jv_coarse = np.loadtxt('Jv_pde_coarse.out')
+    
   #  Jv_pde = Jv_pde_list[1]
     
     #SDE       
@@ -82,30 +83,30 @@ if __name__=="__main__":
 
     bins = len(Jv_sde)
     
-    Jv_coarse = scipy.zeros(bins)
-    for i in range (0,bins):
-        bin_av = 0
-        for j in range (0,  resize_factor ):
-            bin_av = bin_av+ Jv_pde[i*resize_factor+j]
-            Jv_coarse[i] = bin_av/resize_factor
-            
-            
-    resize_factor = int (len(Jv_pde_dxbig)/len(Jv_sde))        
-    Jv_coarse_dxbig = scipy.zeros(bins)
-    for i in range (0,bins):
-        bin_av = 0
-        for j in range (0,  resize_factor ):
-            bin_av = bin_av+ Jv_pde[i*resize_factor+j]
-            Jv_coarse[i] = bin_av/resize_factor
+#    Jv_coarse = scipy.zeros(bins)
+#    for i in range (0,bins):
+#        bin_av = 0
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/post_processing_Newton(N,k).py

@@ -1,0 +1,155 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+import matplotlib.pylab as plt
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+import matplotlib.pylab as plt
+
+
+if __name__=="__main__":
+
+#    residual_directsim = np.loadtxt('Newton/residual_sim_N1e6.out')
+#    residual1 = np.loadtxt('Newton/residual1_N1e6.out') 
+#    residual2 = np.loadtxt('Newton/residual2_N1e6.out') 
+#    residual3 = np.loadtxt('Newton/residual3_N1e6.out')  
+#    
+#    residual_directsim = np.loadtxt('Newton/residual_sim(N)_Dte-1_tol1e-3.out')
+#    residual1 = np.loadtxt('Newton/residual1(N)__Dte-1_tol1e-3.out') 
+#    residual2 = np.loadtxt('Newton/residual2(N)__Dte-1_tol1e-3.out') 
+#    residual3 = np.loadtxt('Newton/residual3(N)_Dte-1_tol1e-3.out')  
+#    
+
+#    resnorm_t100_N1000 = np.loadtxt('Newton/resnorm_Dte-1_N1000')
+#    resnorm_t100_N10000 = np.loadtxt('Newton/resnorm_Dte-1_N10000') 
+#    resnorm_t100_N100000 = np.loadtxt('Newton/resnorm_Dte-1_N100000') 
+#    resnorm_t100_N1000000 = np.loadtxt('Newton/resnorm_Dte-1_N1000000')  
+#    resnorm_t100_N10000000= np.loadtxt('Newton/resnorm_Dte-1_N10000000')  
+    
+    resnorm_t10_tol2_N1000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N1000')
+    resnorm_t10_tol2_N10000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N10000') 
+    resnorm_t10_tol2_N100000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N100000') 
+    resnorm_t10_tol2_N1000000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N1000000')  
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/post_processing_Newton(t).py

@@ -1,0 +1,124 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+import matplotlib.pylab as plt
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+import matplotlib.pylab as plt
+
+if __name__=="__main__":
+
+    residual_directsim = np.loadtxt('Newton/residual_sim(t)_Ne5_tol1e-2.out')
+    residual1 = np.loadtxt('Newton/residual1(t)_Ne5_tol1e-2.out') 
+    residual2 = np.loadtxt('Newton/residual2(t)_Ne5_tol1e-2.out') 
+    residual3 = np.loadtxt('Newton/residual3(t)_Ne5_tol1e-2.out') 
+    
+    residual_directsim_Ne6 = np.loadtxt('Newton/residual_sim(t)_Ne6_tol1e-2.out')
+    residual1_Ne6  = np.loadtxt('Newton/residual1(t)_Ne6_tol1e-2.out') 
+    residual2_Ne6  = np.loadtxt('Newton/residual2(t)_Ne6_tol1e-2.out') 
+    residual3_Ne6  = np.loadtxt('Newton/residual3(t)_Ne6_tol1e-2.out')  
+    
+    residual_directsim_Ne7 = np.loadtxt('Newton/residual_sim(t)_Ne7_tol1e-2.out')
+    residual1_Ne7 = np.loadtxt('Newton/residual1(t)_Ne7_tol1e-2.out') 
+    residual2_Ne7  = np.loadtxt('Newton/residual2(t)_Ne7_tol1e-2.out') 
+    residual3_Ne7  = np.loadtxt('Newton/residual3(t)_Ne7_tol1e-2.out')  
+#    
+    Dtlist = [ 5e-3, 1e-2, 2e-2, 5e-2, 1e-1, 2e-1, 5e-1, 1]
+    
+    
+    ax = plt.subplot(111)  
+    box = ax.get_position()
+    ax.set_position([box.x0, box.y0, box.width * 0.85, box.height])   
+   # Nlist = scipy.array([1e4,5e4, 1e5, 5e5, 1e6, 5e6, 1e7, 5e7]) 
+       
+   # fig = plt.figure()
+   # ax = fig.add_subplot(1,1,1)
+ #   ax = plt.subplot(111)  
+  #  box = ax.get_position()
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/post_processing_Newton.py

@@ -1,0 +1,128 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+import matplotlib.pylab as plt
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+import matplotlib.pylab as plt
+
+
+if __name__=="__main__":
+
+#    residual_directsim = np.loadtxt('Newton/residual_sim_N1e6.out')
+#    residual1 = np.loadtxt('Newton/residual1_N1e6.out') 
+#    residual2 = np.loadtxt('Newton/residual2_N1e6.out') 
+#    residual3 = np.loadtxt('Newton/residual3_N1e6.out')  
+#    
+#    residual_directsim = np.loadtxt('Newton/residual_sim(N)_Dte-1_tol1e-3.out')
+#    residual1 = np.loadtxt('Newton/residual1(N)__Dte-1_tol1e-3.out') 
+#    residual2 = np.loadtxt('Newton/residual2(N)__Dte-1_tol1e-3.out') 
+#    residual3 = np.loadtxt('Newton/residual3(N)_Dte-1_tol1e-3.out')  
+#    
+
+#    resnorm_t100_N1000 = np.loadtxt('Newton/resnorm_Dte-1_N1000')
+#    resnorm_t100_N10000 = np.loadtxt('Newton/resnorm_Dte-1_N10000') 
+#    resnorm_t100_N100000 = np.loadtxt('Newton/resnorm_Dte-1_N100000') 
+#    resnorm_t100_N1000000 = np.loadtxt('Newton/resnorm_Dte-1_N1000000')  
+#    resnorm_t100_N10000000= np.loadtxt('Newton/resnorm_Dte-1_N10000000')  
+    
+    resnorm_t10_tol2_N1000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N1000')
+    resnorm_t10_tol2_N10000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N10000') 
+    resnorm_t10_tol2_N100000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N100000') 
+    resnorm_t10_tol2_N1000000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N1000000')  
+    resnorm_t10_tol2_N10000000= np.loadtxt('Newton/resnorm_Dte-2_tole-2_N10000000')  
+
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/post_processing_Newton_variance.py

@@ -1,0 +1,77 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+import matplotlib.pylab as plt
+from matplotlib.path import Path #to draw trriangles indicating order of convergence
+import matplotlib.patches as patches #to draw trriangles indicating order of convergence
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+import matplotlib.pylab as plt
+
+
+if __name__=="__main__":
+   
+
+    
+    Nlist = scipy.array([1e3, 1e4, 1e5, 1e6])
+    Nlist_inv = 1.0/Nlist
+    rho_sq_tol2 = np.loadtxt('Newton/rho_sq.out')         #M=50
+    rho_sq_tol4 = np.loadtxt('Newton/rho_sq_tol_e-4.out')  
+    rho_sq_tol3 = np.loadtxt('Newton/rho_sq_tol_e-3.out') 
+    rho_sq_tol5 = np.loadtxt('Newton/rho_sq_tol_e-5.out')  
+    #np.loadtxt('25-11-Rho_pde_dx_e-4_eps-6.out') ,  np.loadtxt('25-11-Rho_pde_dx_e-4_eps-7.out') ]  #these data should be the same
+#    Jv_pde= np.loadtxt('data/25-11-Jv_pde_dx_e-4_eps-5.out') 
+   # Jv_pde_dxbig= np.loadtxt('data/Jv_pde_dx_5e-4_vsin.out') 
+    E_rho_tol2 = np.loadtxt('Newton/E_rho.out')       #M=50
+    E_rho_tol4 = np.loadtxt('Newton/E_rho_tol_e-4.out')  
+    E_rho_tol3 = np.loadtxt('Newton/E_rho_tol_e-3.out') 
+    E_rho_tol5 = np.loadtxt('Newton/E_rho_tol_e-5.out') 
+  #  Jv_pde = Jv_pde_list[1]
+    
+    #SDE       
+    variance =  scipy.zeros(len(Nlist))
+    variance4 =  scipy.zeros(len(Nlist))
+    variance3 =  scipy.zeros(len(Nlist))
+    variance1 =  scipy.zeros(len(Nlist))
+    variance5 =scipy.zeros(len(Nlist))
+    
+    for n in range (0, len(Nlist)):
[…]

### 2016-04-15 Problems/WeightedParticles/checkSystem/post_processing_bif.py

@@ -1,0 +1,70 @@
+# -*- coding: utf-8 -*-
+import scipy
+"""
+Created on Wed Mar 23 20:01:01 2016
+
+@author: pieter
+"""
+
+#prev_point = branch.points[0].u
+#last_point = branch.points[1].u
+
+#STOCHASTIC
+
+save_flag= False
+bif_par = zeros( len(branch.points))
+
+for i in range ( len(branch.points)):
+    bif_par[i] = branch.points[i].lambd[0]
+    
+for i in range (len(branch.points)):
+    par_label = r'$a = %g $' % bif_par[i]
+  #  par_label = r'$D = %g $' % bif_par[i]
+    plt.plot(branch.points[i].u, label=par_label)
+    
+plt.legend()
+if save_flag:
+    plt.savefig('plots/fixed_states_sde(a)_1.pdf')
+    
+plt.show()
+    
+#x, y=  branch.getPlot()
+
+
+plt.show()
+    
+#DETERMINISTIC
+bif_par_pde = zeros( len(branch_pde.points))
+
+for i in range ( len(branch_pde.points)):
+    bif_par_pde[i] = branch_pde.points[i].lambd[0]
+    
+for i in range (len(branch_pde.points)):
+  #  par_label = r'$\alpha = %g $' % bif_par_pde[i]
+    par_label = r'$a = %g $' % bif_par_pde[i]
+    plt.plot(branch_pde.points[i].u, label=par_label)
+
+plt.legend()
+if save_flag:
+    plt.savefig('plots/fixed_states_pde(a)_1.pdf')
+plt.show()
+    
+#x_pde, y_pde=  branch_pde.getPlot()
+plt.plot(x_pde, y_pde, 'ro') #, markersize=3)
+plt.ylabel(r'$\left\||\rho^*\right\||$', fontsize = 16)
+plt.xlabel(r'$a$', fontsize = 12)
+plt.savefig('plots/bifurcation_pde(D).pdf')
+
+plt.plot(x_sde, y_sde, 'bo')
+
+if save_flag:
+    plt.savefig('plots/bifurcation_pde_1.pdf')
+
+plt.show()
+
+plt.plot(x_sde, abs(y_sde-y_pde), 'bo')
+
+if save_flag:
+    plt.savefig('plots/bifurcation_bias_1.pdf')
+
+plt.show()

### 2016-04-15 Problems/WeightedParticles/checkSystem/post_processing_eps.py

@@ -10,11 +10,12 @@ import sys
 import numpy as np
 import matplotlib.pylab as plt
 
+
 from scipy.linalg import norm
 from scipy import sqrt
 
 import matplotlib.pylab as plt
-
+plt.rc('text', usetex=True)
 
 if __name__=="__main__":
     D = 1./2
@@ -146,7 +147,7 @@ if __name__=="__main__":
        
     for eps_i in range (0,len(eps_list)):
         plot_var = plt.plot(Nlist_inv, (sde_Jv_sq[eps_i] -sq_E_Jv[eps_i])/bins, lines[eps_i],
-                            label =r'$\varepsilon=10^{-%d}$' %eps_list_exponents[eps_i] , linewidth=2)
+                            label =r'$\varepsilon=10^{-%d}$' %eps_list_exponents[eps_i] , linewidth=3)
     plt.ylabel('Var ($\mathbf{\hat{Jv}} $)', fontsize = 16)
     plt.xlabel('$1/N$', fontsize = 16) 
     plt.xscale(log_flag)
@@ -154,6 +155,8 @@ if __name__=="__main__":
     plt.legend([plot_var], loc='best')
     plt.legend(bbox_to_anchor=(1, 0.5), numpoints = 1 )
     plt.gca().add_patch(triangle_o1var)
+    order= r'$\mathcal{O}(1/N)$'
+    plt.annotate(order,  xy=(2e-4, 2e-2), xytext=(1.1e-4, 5.2e-2), fontsize=11, color='grey')
     if(save_flag): plt.savefig("plots/Var_N_eps.pdf")
     plt.show()
     

### 2016-04-15 Problems/WeightedParticles/gmres/gmres_sde.py

@@ -126,8 +126,8 @@ if __name__=="__main__":
 
     # POINT PARAMETERS
     point_param = Point.Point.getDefaultParameters()
-    point_param['artificial']=[4]
-    point_param['artificial_condition']=[probability.condition]
+  #  point_param['artificial']=[4]
+  #  point_param['artificial_condition']=[probability.condition]
     
     pprev = p
     p = Point.Point(fp_sde,nsolv,None,point_param)

### 2016-04-15 Problems/WeightedParticles/lifting/inv_transform_sampling.py

@@ -24,7 +24,7 @@ class Sampler(object):
     def lift(self,rho,grid,N,skip=1):
         cum,edges = self.rho2cum(rho,grid)
         y = self.rand.uniform(size=N)
-      #  print y
+      #  print "random sampling numbers " , y
    #     print "sampling density with seed : ", self.cur_seed, "first sample: ", y[0]
         return self.inv_cum(y,cum,edges)
     
@@ -39,10 +39,10 @@ class Sampler(object):
         
         cum[0]=0.
         for n in range(ngrid):
-            if rho[n]<0 :
-                print "encountered a negative density : ", rho[n]
+         #   if rho[n]<0 :
+         #       print "encountered a negative density : ", rho[n]
             cum[n+1]=cum[n]+rho[n]*(edges[n+1]-edges[n])
-        print "estimating cumulative density : last value = ", cum[-1], " (should be 1.)"
+     #   print "estimating cumulative density : last value = ", cum[-1], " (should be 1.)"
         if scipy.absolute(cum[-1]-1)>1e-7:
             print "cumulative: ", cum
         return cum,edges
@@ -104,5 +104,25 @@ if __name__ == "__main__":
     indices = scipy.argsort(y)
     xsort = sampler.inv_cum(ysort,cum,edges)
     print scipy.amax(xsort-x[indices])   
+    
+    dx= 1e-1    
+    xL = -1.7
+    xR =1.7
+    grid = scipy.arange(xL+dx/2.,xR,dx)
+    rho_newton = scipy.array([ 0.00056236,  0.00056236,  0.00056236,  0.00075873,  0.00780793,
+        0.03639701,  0.1223761 ,  0.25240819,  0.39957626,  0.46229259,
+        0.45699236,  0.42253595,  0.33718642,  0.27315376,  0.22453869,
+        0.2013031 ,  0.18209444,  0.18183162,  0.19075894,  0.23761909,
+        0.2595615 ,  0.34807625,  0.43138754,  0.47717044,  0.48015413,
+        0.39398931,  0.25465005,  0.118413  ,  0.03884051,  0.00526264,
+        0.00126285,  0.00056236,  0.00056236,  0.00056236])
+    rho_newton = rho_newton/(sum(rho_newton)*dx)
+    param = Sampler.getDefaultParameters()
+    cum,edges = sampler.rho2cum(rho_newton,grid)
+    y = scipy.random.uniform(size=1000)
+    x = sampler.inv_cum(y,cum,edges)
+    
+        
+        
 
     

### 2016-04-15 Problems/WeightedParticles/system/particles.py

@@ -3,6 +3,8 @@ from scipy.linalg import norm
 from scipy import sqrt
 import numpy as np
 import sys         
+import matplotlib.pylab as plt
+
 sys.path.append('../')   #added to resolve importerror of nephew-directories
 sys.path.append('../../')
 sys.path.append('../../../')
@@ -117,39 +119,62 @@ class Particles(TSystem.TimestepperSystem):
         x = self.lifting.lift(rho,grid,self.param['Nsmall'],self.skip)
         return x
     
-    def simulate(self,x0,seedlist,lambd):
-        #print "in simulate : ", lambd
-        x_Dt = scipy.zeros_like(x0)
-        for i in range(len(x0)):
-            x_Dt[i]=self.simulate_particle(x0[i],seedlist[i],lambd)
-#            if (i==42): 
-#                print 'position diff for particle 42 =' , x_Dt[i]-x0[i]
-        return x_Dt
-    
-    def simulate_particle(self,x,seed,lambd):
-        tstart = 0
-        tcur = tstart
+#    def simulate(self,x0,seedlist,lambd):
+#        #print "in simulate : ", lambd
+#        x_Dt = scipy.zeros_like(x0)
+#        for i in range(len(x0)):
+#            x_Dt[i]=self.simulate_particle(x0[i],seedlist[i],lambd)
+##            if (i==42): 
+##                print 'position diff for particle 42 =' , x_Dt[i]-x0[i]
+#        return x_Dt
+#    
+#    def simulate_particle(self,x,seed,lambd):
+#        tstart = 0
+#        tcur = tstart
+#        Dt = self.param['Dt']
+#        # Dt needs to be a multiple of param['Dt']
+#        dt = self.param['dt']
+#        D = lambd[1]
+#        a = lambd[0]
+#       # print "seed", seed
+#      #  self.seed_particle(seed)    # is this why every particle gets different random numbers?
+#        while (tcur < tstart + Dt - dt/2 ):
+#            tcur += dt
+#            # the random number
+#            dW = self.getBrownianIncrement()
+#     #       if tcur == dt:    #only print random number for first time step
+#            #    print dW
+#            # the process
+#            drift_term = a * self.drift(x)
+#            x=x+drift_term*dt+sqrt(2*D*dt)*dW
+#            # and reflecting boundary conditions
+#            if (x>self.domain[1]):
[…]

### 2016-04-15 Problems/WeightedParticles/system/pde.py

@@ -11,7 +11,7 @@ sys.path.append('../../../')
 import System.TimestepperSystem as TSystem
 
 def doublewell(x):
-    return 2*x-4*x**3   
+    return  2*x-4*x**3   
 
 
 def d2udx2(u,dx):

### 2016-04-15 Solver/GMRESLinearSolver.py

@@ -5,6 +5,9 @@ from scipy.sparse.linalg import gmres as sgmres
 from scipy.linalg import norm, lstsq
 from scipy import zeros,dot,r_
 import LinearSolver
+import matplotlib.pylab as plt
+import scipy
+
 
 class GMRESLinearSolver(LinearSolver.LinearSolver):
 
@@ -23,12 +26,13 @@ class GMRESLinearSolver(LinearSolver.LinearSolver):
         """
         LinearSolver.LinearSolver.__init__(self,param)
         self.resid = zeros((0,))
+        self.kappa =  zeros((0,))
         self.createCallback()
     
     def createCallback(self):
         def callback(res):
             self.resid = r_[self.resid,res]
-            if self.param['print']=='short':
+            if self.param['print']=='long':
                 print "gmres residual: ", len(self.resid), res
         self.callback=callback
     
@@ -66,7 +70,7 @@ class GMRESLinearSolver(LinearSolver.LinearSolver):
              k>0 : maximum number of iterations reached
              k<0 : illegal input or breakdown
         """
-        self.resid = zeros((0,))
+        #self.resid = zeros((0,))
         tol = self.param['tol']
         if self.param['relative']:
             tol = max(self.param['rel_tol']*norm(rhs),tol)
@@ -87,6 +91,15 @@ class GMRESLinearSolver(LinearSolver.LinearSolver):
         else:
             x,info=self.gmres(rhs,tol=tol,restrt=restart,\
                 maxiter=maxiter)
+        if print_it == 'long':  
+            print 'Convergence in GMRES!'
+            plt.title('Final GMRES-residual')
+            plt.plot(rhs-self.point.system.applyJacobian(x))        
+            plt.show()
+            print ' for vector '
+            plt.plot(x)
+            plt.title(r'$\delta_{final}$')
+            plt.show()
         return (x,info)
 
 
@@ -143,11 +156,13 @@ class GMRESLinearSolver(LinearSolver.LinearSolver):
         r = rhs - self.point.matvec(x)
         beta = norm(r)
         myres = beta/len(r)
-        V[:,0]=r/beta
-        print "in gmres",
-        print myres,tol
+        V[:,0]=r/beta 
+        print_it = self.param['print']
+        if print_it == 'long': 
+            print "in gmres",
+            print myres,tol
         while j < maxiter and myres > tol:
-            z = self.point.psolve(V[:,j])
+            z = self.point.psolve(V[:,j])   
[…]

### 2016-04-15 Solver/NewtonSolver.py

@@ -1,6 +1,10 @@
 import scipy
 import Solver
 import LinearSolver
+from scipy import zeros,dot,r_
+
+
+import matplotlib.pylab as plt
 
 class NewtonSolver(Solver.Solver):
 
@@ -26,8 +30,13 @@ class NewtonSolver(Solver.Solver):
         else:
             raise TypeError, "input argument " + linear_solver \
                 + " should be a linear solver"
-        self.nb_newt = 0
-    
+        self.nb_newt = 0 
+        self.newton_residual = zeros((0,))
+        self.newton_res_norm = zeros(0,)
+       # self.newton_states = zeros((  param['max_iter'],len( self.point.getCurrentGuess()) ))
+        self.newton_states = zeros((  0 ))
+      #  self.iterations  = 0
+
     def getDefaultParameters():
         """  
         Returns a dictionary of default parameters 
@@ -66,20 +75,32 @@ class NewtonSolver(Solver.Solver):
         print_it = self.param['print']
         alpha = self.param['damping']
         stop = self.param['stop']
-        
-        res = self.point.getResidual()
         x = self.point.getCurrentGuess()
-        res_norm = scipy.linalg.norm(res)
+        self.newton_states = x
+        res = self.point.getResidual()
+        res_norm = scipy.linalg.norm(res)/len(res)
+        self.newton_res_norm  =res_norm
+        print 'res norm init ', self.newton_res_norm
+        grid = self.point.system.grid
+        grid_dx = grid[-1] - grid[-2]
         if stop == "step":
             stop_norm = 0
         elif stop == "res":
             stop_norm = res_norm
             
-        while iter==0 or (iter < max_iter and stop_norm > abs_tol):
-
+        #while (iter < max_iter and (stop_norm > abs_tol or iter ==0)): geprobeerd om iedentieke toestanden te vermijden in opl van pde, maar als het residu te klein is ivm de tolerantie, dan heeft een newtonstap geen zin want dan wordt dx toch 0
+        while (iter < max_iter and (stop_norm > abs_tol)):
             iter += 1
             self.nb_newt += 1
-            dx, status = self.linsolv.solve(-res)
+            dx, status = self.linsolv.solve(-res)        # res = rho - rho.Dt
+            if print_it == 'long':  
+                plt.plot(res)                  
+                plt.title(r'Res before newton step = $ \rho - \phi(\rho) $ ')
+                plt.show()
+                plt.plot(dx)
[…]

### 2016-04-15 System/TimestepperSystem.py

@@ -43,13 +43,20 @@ class TimestepperSystem(System.System):
         self.lambd = lambd
         self.u_Dt = self.integrate(u,lambd)
        # self.x_prev = x_prev
-        print self.u_Dt
+        #print self.u_Dt
         print "leaving setState"
+        
+    def updateState(self, u, lambd):               #needed in Newton-solver?
+        self.u = self.integrate(u,lambd)
+        self.lambd = lambd
+                
+        
     def getResidual(self):
         """
         This method determines what the residual for the system is.
         """
         res = self.u - self.u_Dt
+      #  print '         Residual = ', res
         return res
     
     def computeJacobian(self):
@@ -57,7 +64,7 @@ class TimestepperSystem(System.System):
         returns the Jacobian of the system for the given state and
         parameter vectors
         """
-        print "ATTENTION: RUNNING COMPUTE JACOBIAN ON TIMESTEPPER SYSTEM"
+        print "ATTENTION: RUNNING COMPUTE JACOBIAN ON TIMESTEPPER SYSTEM!"
         n = len(self.u)
         A = scipy.zeros((n,n))
         for i in range(n):
@@ -72,11 +79,18 @@ class TimestepperSystem(System.System):
         approximation to) (1-J)*v 
         """
         
-        print "WE ARE IN FP"
+    #    print "we are in FP"
         eps = self.param['eps']
         u_eps = self.u + eps * v/norm(v)
+        
         u_eps_Dt = self.integrate(u_eps,self.lambd)
-        result = (u_eps_Dt - self.u_Dt)/eps*norm(v)
+        
+#        print 'checksum u_eps_Dt = ' , sum(u_eps_Dt)
+#        print 'checksum u__Dt = ' , sum(self.u_Dt)
+#        print 'checksum Jv= ' , sum( (u_eps_Dt -self.u_Dt)/eps*norm(v))
+        result = v - (u_eps_Dt - self.u_Dt)/eps*norm(v)
+#        print 'checksum v= ' , sum(v)
+#        print 'check j-Jv' , sum(result)
         return result   
     
     def getParameterDerivative(self,i):

### 2016-05-02 Documentation/draft.tex
Added errorbars on NK-solution

@@
 
-We investigate the emergence of coarse equilibrium states in a stochastic system of p
+In this project, we develop a Newton-Krylov method that is able to compute efficiently coarse fixed points when the underlying fine-scale dynamics is stochastic.
+We investigate the use of correlated samples to perform a more accurate approximation of the Jacobian vector products that are needed in a Krylov solution in each Newton iteration.
+The main novelty of the algorithm is in the elimination of the noise that is generated when estimating Jacobian-vector products using time-integration of perturbed initial conditions.
+We present numerical results that demonstrate the convergence properties of the numerical method, and use the method to show   the emergence of coarse equilibrium state
 
- system of diffusion processes have a stabilizing force acting on each of them, corresponding to a bistable potential.
+%system of diffusion processes have a stabilizing force acting on each of them, corresponding to a bistable potential.
 
-We develop a Newton-Krylov method that is able to compute accurately and efficiently coarse fixed points when the underlying fine-scale dynamics is stochastic.
-The main novelty of the algorithm is in the elimination of the noise that is generated when estimating Jacobian-vector products using time-integration of perturbed initial conditions.
-We present numerical results that demonstrate
-%the convergence properties of the numerical method, and use the method to show that macroscopic
+%nd use the method to show that macroscopic
 %fronts in this model destabilise at a coarse symmetry-breaking bifurcation.
@@
 
-\section{Problem}
+\section{Model problem}
 
- As a toy problem we use a model where the closed model is exact.
-In general, we are interested in performing a bifurcation analysis for a model for which at which an exact, closed model at the macorscopic level is not available.
-\subsection{Model problem}
-
-\input{fokkerplanck_update.tex}
-
-\subsection{Discretization}
+\input{model_problem.tex}
 
@@
 
-The convergence of the density to the real solution, strongly depends on the number of particles $N$ and on the choise of $\epsilon$ in the GMRES-method.
+The convergence of the density to the real solution, strongly depends on the number of particles $N$ and on the choice of $\epsilon$ in the GMRES-method.
 
@@
 Until now, we considered a system of non-interacting particles.
-We will now extend our model to a mean field model by introducing a third parameter $\alpha$, which is the degree of interaction or cooporation in the system.
+We will now extend our model to a mean field model by introducing a third parameter $\alpha$, which is the degree of interaction or cooperation in the system.
 A simple form of cooperative behaviour is the case where each agent tends to follow the state of the majority (or, each particle feels an attractive force towards the mean state of the system).
@@
 The evolution of the system is now characterized by  by the initial conditions, the three parameters ($\mu$, $\sigma$, $\alpha$) and by the system size $N$.
-Figure \ref{fig:sysrisk}  illustrates the behavior of the empiral mean $\bar{x}$.
+Figure \ref{fig:sysrisk}  illustrates the behaviour of the empirical mean $\bar{x}$.
 The simulations were performed with all agents initially in the normal state.
 Nevertheless, if randomness dominates the interaction, the agents can move immediately to the other potential well.
-The system then behaves like $N$ independent difussions, and hence, by the symmetry of the potential, the mean state will be attracted to a single mixed state $\bar{x}=0$.
-Upon increasing the interaction parameter $\alpha$, however, we find two new macroscopic states, suggesting the precence of a pitchfork bifurcation at the macoscopic level.
-These solutions are no stable steady states, but rather coarse matastable states.
+The system then behaves like $N$ independent diffusions, and hence, by the symmetry of the potential, the mean state will be attracted to a single mixed state $\bar{x}=0$.
+Upon increasing the interaction parameter $\alpha$, however, we find two new macroscopic states, suggesting the precence of a pitchfork bifurcation at the macroscopic level.
+These solutions are no stable steady states, but rather coarse metastable states.
 Their lifetime is linked to the finite system size.
@@
 
+\subsection{Results for the sde}
+We chose $\Delta \sigma = 0.05$ as continuation step seize.
+Performing a bifurcation with smaller steps is pointless, because the difference in densities will become too small in comparison with the noise on the residual.
+
+\begin{figure}
+\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
+\caption{  Bifurcation diagram of the steady states calculated with the Newton-Krylov-Solver for the SDE.
+$\Delta t = 10^{-3}, \Delta T = 10^{-1}, \Delta x = 10^{-2}, \epsilon_{GMRES}=10^{-5},  \delta_{Newton} = 10^{-7}, \epsilon=10^{-5}$
+}
+\end{figure}
+
 %\begin{center}
@@
 
-\section{Newton convergence}
+\section{Convergence of the Newton Method}
+\input{newton_convergence.tex}
 

### 2016-05-02 Documentation/newton_convergence.tex
Added errorbars on NK-solution

Numerical experiments that show how the Newton method converges to the fixed point:

What is the maximal accuracy that can be obtained? (Relate to the variance of the coarse time-stepper.) How to cheaply estimate the stopping criterion?
Is the resulting fixed point biased?
Can we reduce variance by averaging over a number of 'converged' Newton iterations?
Additional questions of interest:

Can we stop the Krylov iterations before reaching machine precision (knowing that the end result will contain noise anyway)? Does an inaccuracy in the solution of the linear systems increase the required number of Newton iterations?
Do we keep the particles fixed throughout the Newton iterations as well? (We can then probably converge the Newton procedure to machine precision.) What is then the probability distribution of the obtained fixed points? Bias? Variance?

### 2016-05-02 Continuer/Continuer.py

@@ -224,7 +224,7 @@ class Continuer(object):
         for i in range(len(self.points)):
             x[i] = self.get(self.points[i],"x")
             y[i] = self.get(self.points[i],"y")
-            print x[i],y[i]
+        #    print x[i],y[i]
         return x,y
         
     def getIterations(self):

### 2016-05-02 Problems/WeightedParticles/checkSystem/check_bif_pde.py

@@ -37,22 +37,25 @@ import Continuer.Continuer as Continuer
 
 
 
+
 if __name__=="__main__":
    # D = 1./2.
-    sigma =1
-    D= 1#1/2.
+    mu = 1.0
+    sigma=2.0
+    D=0.5*sigma**2
+  #  D = 2.0
+   # sigma = 2*np.sqrt(D)
     Dt = 1e-2
   #  Dtlist = [1e-5,2e-5, 5e-5, 1e-4, 2e-4, 5e-4, 1e-3, 2e-3, 5e-3 , 1e-2, 2e-2, 5e-2 , 1e-1, 2e-1, 5e-1, 1, 2, 5, 10]
     Dtlist = [Dt]
  #   Dtlist = [2e-3]
-    seed = 15
-    # discretization parameters
+    # discretizatio    seed = 15n parameters
    #h=2e-2 # kde mesh width 
     dx = 1e-2
 #    dxlist= [1e-2,  5e-3, 1e-3, 5e-4] 
   #  dxlist= [1e-2, 1e-3]
     #dt = min(dxlist)**2/(2*D)
-    dt=5e-5
+    dt=1e-5
     r= (2.*D*dt)/(dx)**2
     print'r = ' ,  r
 
@@ -62,12 +65,10 @@ if __name__=="__main__":
         sys.exit("Ending Script. Make sure that r<1") 
     xL = -1.7
     xR = 1.7
-    
-    a = 0.2
     zeta = 0.
     alpha=1
     beta = -1
-    lambd = scipy.array([a,D,alpha])
+    lambd = scipy.array([mu, sigma])
     param=pde.FokkerPlanck.getDefaultParameters()
     param['Dt'] = Dt
     param['dt']=dt         
@@ -143,8 +144,8 @@ if __name__=="__main__":
         newton_states = nsolv.newton_states
         nstates =  newton_states.reshape(nsolv.nb_newt+1, len(grid))
         
-        norm_c = sum( np.exp( (-grid**4 + grid**2)*a/D))*dx
-        rho_ss = np.exp( (-grid**4 + grid**2 )*a/D) /norm_c
+        norm_c = sum( np.exp( (-grid**4 + grid**2)*mu/D))*dx
+        rho_ss = np.exp( (-grid**4 + grid**2 )*mu/D) /norm_c
         label2= r'$\frac{ \exp{\left[-\frac{V(x)}{\sigma^2}\right]}}{\mathcal{N}}$'
[…]

### 2016-05-02 Problems/WeightedParticles/checkSystem/check_bif_sde.py

@@ -38,7 +38,12 @@ import Continuer.Continuer as Continuer
 
 if __name__=="__main__":
     print 'Start solving sde'
-    D = 1
+ #   sigma=2
+    mu = 1.0
+  #  D=0.5*sigma**2
+    D = 2.0
+    sigma=1.0
+   # sigma = 2*np.sqrt(D)
     Dt = 1e-1  #Dt influences number of GMRES-iterations strongly
     Dtlist = [ Dt]
     #Dtlist = [ 5e-3, 1e-2, 2e-2, 5e-2, 1e-1, 2e-1, 5e-1, 1]
@@ -49,11 +54,11 @@ if __name__=="__main__":
    # dt=1e-6
     xL = -1.7
     xR = 1.7
-    a = 0.2
+
     zeta = 0.
     alpha=1
     beta = -1
-    lambd = scipy.array([a,D,alpha,beta,zeta])
+    lambd = scipy.array([mu, sigma])
        
     #SDE
     t1 = time.time()
@@ -77,7 +82,11 @@ if __name__=="__main__":
   #  lifting =  inv_transform.Sampler(sampler_param)
     
     h=2e-2 # kde mesh width     
-    M=1 #number of monte carlo steps                
+    M=10 #number of monte carlo steps         
+    y_mean = scipy.zeros(M)
+    y_sde = scipy.zeros(M)
+
+    branch_list = [] #scipy.array([M])              
     param_histogram = histogram.Histogram.getDefaultParameters()
     param_histogram['h']=h
     restriction = histogram.Histogram(param_histogram)
@@ -92,7 +101,7 @@ if __name__=="__main__":
 
  #   Nlist = scipy.array([1000, 2000, 4000, 8000, 16000, 32000, 64000, 128000])
 #    Nlist = scipy.array([1e3, 1e4, 1e5, 1e6, 1e7, 1e8]) 
-    Nlist = scipy.array([1e5]) #, 800, 1600, 3200])
+    Nlist = scipy.array([1e6]) #, 800, 1600, 3200])
     eps_list_exponents=[5]
     eps_list= [1e-5]
     param['eps']= eps_list[-1]
@@ -115,6 +124,7 @@ if __name__=="__main__":
     residual2 = scipy.zeros(len(Dtlist))
     residual3 = scipy.zeros(len(Dtlist))
     residual_directsim = scipy.zeros(len(Dtlist)) 
+
 #    
             
[…]

### 2016-05-02 Problems/WeightedParticles/checkSystem/check_pde.py

@@ -27,16 +27,18 @@ import pde
 
 
 if __name__=="__main__":
-    D = 1./2.
-    Dt = 2
+    sigma=1
+    D=0.5*sigma**2
+    Dt = 1
     seed = 16
     # discretization parameters
    #h=2e-2 # kde mesh width 
-    dx = 1e-3
+    dx = 1e-2
 #    dxlist= [1e-2,  5e-3, 1e-3, 5e-4] 
-    dxlist= [1e-2]
-    dt = min(dxlist)**2/(2*D)
+    dxlist= [dx]
+    #dt = min(dxlist)**2/(2*D)
     #r= (2.*D*dt)/(dx)**2
+    dt=1e-5
     r= (2.*D*dt)/(min(dxlist))**2
     if r>1: 
         print 'Stability condition not fulfilled, because r=',r #(see https://en.wikipedia.org/wiki/FTCS_scheme)
@@ -44,11 +46,11 @@ if __name__=="__main__":
     xL = -1.7
     xR = 1.7
     
-    a = 1
+    mu = 1
     zeta = 0.
     alpha=1
     beta = -1
-    lambd = scipy.array([a,D,alpha,beta,zeta])
+    lambd = scipy.array([mu, sigma])
     param=pde.FokkerPlanck.getDefaultParameters()
     param['Dt'] = Dt
     param['dt']=dt         
@@ -67,11 +69,8 @@ if __name__=="__main__":
         fp_pde = pde.FokkerPlanck(rho,grid,pde.doublewell,lambd,param)
         rho_Dt = fp_pde.u_Dt
 
-        v=scipy.zeros_like(grid)
-        v[200]=-1
-        v[300]=1
-    
-        Jv_pde = fp_pde.applyJacobian(v)    
+        
+     #   Jv_pde = fp_pde.applyJacobian(v)    
        # plt.plot(grid, Jv_pde)
         plt.plot(grid, rho_Dt)
         plt.show()

### 2016-05-02 Problems/WeightedParticles/checkSystem/check_sde.py

@@ -34,20 +34,21 @@ import Utils.conditions.probability as probability
 
 if __name__=="__main__":
 
-    D = 1
-    Dt = 2
+    sigma = 1
+    Dt = 1
     seed = 16
+    D=0.5*sigma**2
     # discretization parameters
    #h=2e-2 # kde mesh width 
    # dt=1e-6
     xL = -1.7
     xR = 1.7
     
-    a = 1
+    mu = 1
     zeta = 0.
-    alpha=1
-    beta = -1
-    lambd = scipy.array([a,D,alpha,beta,zeta])
+    alpha=0
+    beta = 0
+    lambd = scipy.array([mu, sigma])
             
     #SDE
     t1 = time.time()
@@ -57,7 +58,7 @@ if __name__=="__main__":
 #    factor = int (dx/dx_fp)
 #    print "Discretisation for solving sde is ", factor, " times coarser than the discretisation for solving the pde"
     dt = 1e-2 # ti
-    N=5000000
+    N=1e5
     Nlarge= N
     Nsmall =N
     

### 2016-05-02 Problems/WeightedParticles/checkSystem/check_sde_no_weights.py

@@ -1,193 +1,0 @@
-"""
-Created on Wed Oct 28 10:05:37 2015
-
-@author: pieter
-"""
-import time 
-import tables
-import scipy
-import sys
-import numpy as np
-
-from scipy.linalg import norm
-from scipy import sqrt
-
-#import matplotlib.pylab as plt
-
-sys.path.append("../system/")
-sys.path.append("../lifting/")
-sys.path.append("../restriction/")
-sys.path.append("..")
-sys.path.append("../..")
-
-import pde
-import histogram
-import particles
-import inv_transform_sampling as inv_transform
-
-if __name__=="__main__":
-    D = 1./2.
-    Dt = 1e-2
-    seed = 16
-    # discretization parameters
-   #h=2e-2 # kde mesh width 
-    dx_fp = 5e-4
-   # dt=1e-6
-    xL = -1.7
-    xR = 1.7
-    
-    a = 1
-    zeta = 0.
-    alpha=1
-    beta = -1
-    lambd = scipy.array([a,D,alpha,beta,zeta])
-  
-    
-    #PDE
-#    print 'Start solving pde'
-#    fp_pde = pde.FokkerPlanck(rho,grid,pde.doublewell,lambd,param)
-#    Jv_pde = fp_pde.applyJacobian(v)    
-#    rho_Dt_fp = fp_pde.u_Dt
- #   Jv_pde= np.loadtxt('Jv_fp_dx_1e-4.out')
-#    rho_Dt_fp= np.loadtxt('rho_fp_dx_1e-4.out') 
- #   np.savetxt('rho_fp_dx_5e-5.out', rho_Dt_fp)
- #   np.savetxt('rho_Jv_dx_5e-5.out', Jv_pde)
-            
-    #SDE
-    t1 = time.time()
-    
-    print 'Start solving sde'
-    dx = 1e-2
-    factor = int (dx/dx_fp)
-    print "Discretisation for solving sde is ", factor, " times coarser than the discretisation for solving the pde"
-    dt = 1e-4 # ti
-    N=1000
-    Nlarge= N
-    Nsmall =N
-    
-    
-    grid = scipy.arange(xL+dx/2.,xR,dx)
-    rho = scipy.ones_like(grid)
-    rho = rho/(sum(rho)*dx)
-    
-    
-
-    v=scipy.zeros_like(grid)
-    for j in range(len(grid)): 
-        v[j]= np.sin(j*2*np.pi/len(grid))
-
[…]

### 2016-05-02 Problems/WeightedParticles/checkSystem/post_processing_bif.py

@@ -1,70 +1,198 @@
 # -*- coding: utf-8 -*-
 import scipy
+import numpy as np
+from pylab import *
+from scipy.linalg import norm
 """
 Created on Wed Mar 23 20:01:01 2016
 
 @author: pieter
 """
 
-#prev_point = branch.points[0].u
-#last_point = branch.points[1].u
+dx=1e-2
+xL = -1.7
+xR = 1.7
+grid = scipy.arange(xL+dx/2.,xR,dx)
 
 #STOCHASTIC
 
-save_flag= False
-bif_par = zeros( len(branch.points))
+save_flag= True
+branch = branch_list[-1]
+
+bif_par = scipy.zeros( len(branch.points))
 
 for i in range ( len(branch.points)):
-    bif_par[i] = branch.points[i].lambd[0]
+    bif_par[i] = branch.points[i].lambd[1]
     
 for i in range (len(branch.points)):
-    par_label = r'$a = %g $' % bif_par[i]
-  #  par_label = r'$D = %g $' % bif_par[i]
-    plt.plot(branch.points[i].u, label=par_label)
-    
+  #  par_label = r'$a = %g $' % bif_par[i]
+    par_label = r'$\sigma= %g $' % bif_par[i]
+    plt.plot(grid, branch.points[i].u,  label=par_label)
+plt.xlabel(r'$\sigma$', fontsize = 15)
 plt.legend()
 if save_flag:
-    plt.savefig('plots/fixed_states_sde(a)_1.pdf')
+    plt.savefig('plots/bif/fixed_states_sde(sigma)_Ne6.pdf')
     
 plt.show()
+
+
+for b in range ( len(branch_list)):
+    plt.plot(grid, branch_list[b].points[-1].u, label=b)
+plt.xlabel(r'$\sigma$', fontsize = 15)
+plt.show()
+    
     
 #x, y=  branch.getPlot()
 
+# MEAN FIXED STATE
+M= len(branch_list)
+rho_mean =  scipy.zeros((len(branch.points), len(branch.points[-1].u)))
 
-plt.show()
-    
-#DETERMINISTIC
-bif_par_pde = zeros( len(branch_pde.points))
+for b in range (M):
+    for i in range (len(branch_list[b].points)):
[…]

### 2016-05-02 Problems/WeightedParticles/checkSystem/post_processing_bif_obs.py

@@ -1,0 +1,86 @@
+# -*- coding: utf-8 -*-
+import scipy
+import numpy as np
+from pylab import *
+from scipy.linalg import norm
+"""
+Created on Wed Mar 23 20:01:01 2016
+
+@author: pieter
+"""
+
+dx=1e-2
+xL = -1.7
+xR = 1.7
+grid = scipy.arange(xL+dx/2.,xR,dx)
+
+M=10
+rho_mean = np.loadtxt('bifurcation_data/fixed_states_sde_Ne6_mean_M10.txt')
+y_mean =  np.loadtxt('bifurcation_data/norm_fixed_states_sde_Ne6_mean_M10_RL.txt')
+y_var = np.loadtxt('bifurcation_data/variance_fixed_states_sde_Ne6_mean_M10_RL.txt')
+y_error = sqrt(y_var)/sqrt(M)
+
+#STOCHASTIC
+
+save_flag= False
+N_points = len(y_mean)
+
+sigma_list= scipy.arange(2, 0.9, -0.1)
+        
+for i in range (N_points):
+    par_label = r'$\sigma= %g $' % sigma_list[i]
+    plt.plot(grid, rho_mean[i], label=par_label)   
+plt.xlabel(r'$\sigma$', fontsize = 15)
+plt.legend()
+if save_flag:
+    plt.savefig('plots/bif/fixed_states_sde(sigma)_Ne6_mean_M10.pdf')  
+plt.show()
+
+#ANALYTIC
+mu=1.0
+sigma_list= scipy.arange(1.5, 0.4, -0.1)
+sigma_list= scipy.arange(2, 0.9, -0.1)
+D_list= scipy.arange(1, 0, -0.1)
+y_anal = []
+rho_norm_D = []
+for i in range ( len(sigma_list)):
+    norm_c = sum( np.exp( (-grid**4 + grid**2)*2*mu/(sigma_list[i]**2)))*dx
+    rho_ss = np.exp( (-grid**4 + grid**2 )*2*mu/(sigma_list[i]**2)) /norm_c
+    y_anal.append(norm(rho_ss))
+
+sigma_list_fine= scipy.arange(2.1, 0.9, -0.001)
+y_anal_fine = []
+for i in range ( len(sigma_list_fine)):
+    norm_c = sum( np.exp( (-grid**4 + grid**2)*2*mu/(sigma_list_fine[i]**2)))*dx
[…]

### 2016-05-02 Problems/WeightedParticles/system/particles.py

@@ -11,7 +11,7 @@ sys.path.append('../../../')
 
 import System.TimestepperSystem as TSystem
 
-def create_doublewell(alpha,beta):
+def create_doublewell():
     def doublewell(x):
        # return -4*x**3+3*(alpha+beta)*x**2-2*alpha*beta*x
         return 2*x-4*x**3  
@@ -32,7 +32,8 @@ class Particles(TSystem.TimestepperSystem):
         if param == None :
             param = Particles.getDefaultParameters()
         self.param = param
-        self.drift=create_doublewell(alpha=lambd[2],beta=lambd[3])
+     #   self.drift=create_doublewell(alpha=lambd[2],beta=lambd[3])
+        self.drift=create_doublewell()
         self.control= control
 #        if control == None:
 #            self.N=param['Nlarge']
@@ -157,9 +158,10 @@ class Particles(TSystem.TimestepperSystem):
     def simulate(self,x, seedlist, lambd):   #using this for-loop works 100 times faster in comparison with the implementation in 
         Dt = self.param['Dt']
         dt = self.param['dt']
-        D= lambd[1]
-        sigma = np.sqrt(D)
-        a = lambd[0]
+       # sigma= sqrt(2*lambd[1])   
+        sigma = lambd[1]
+      #  sigma = np.sqrt(D)
+        mu = lambd[0]
      #   alpha =lambd[2]
         
         n_steps = int(Dt/dt)
@@ -170,7 +172,7 @@ class Particles(TSystem.TimestepperSystem):
            # if tcur == 1:    #only print random number for first time step
             #    print 'dW =',  dW            
                     # the process
-            drift_term = a * self.drift(x)
+            drift_term = mu * self.drift(x)
             x=x+ drift_term*dt+dW 
             # and reflecting boundary conditions
             scipy.where(x>self.domain[1],2*self.domain[1]-x,x)

### 2016-05-02 Problems/WeightedParticles/system/pde.py

@@ -23,7 +23,9 @@ class FokkerPlanck(TSystem.TimestepperSystem):
     """
     Implements a PDE based time-stepper for the PDE
     
-    rho_t + [a*(x-x**3) \rho]_x = D \rho_xx 
+    originally rho_t + [a*(x-x**3) \rho]_x = D \rho_xx 
+    
+    replaced by  rho_t + [mu*(x-x**3) \rho]_x = 0.5*sigma^2 \rho_xx 
     """
     def __init__(self,rho,grid,drift,lambd,param=None):
         self.grid = grid
@@ -56,10 +58,13 @@ class FokkerPlanck(TSystem.TimestepperSystem):
     def rhs(self,rho,lambd):
         grid = self.grid
         dx = grid[1]-grid[0]
-        D = lambd[1]
-        a = lambd[0]
+   #     sigma = 2*np.sqrt(lambd[1])
+      #  D=lambd[1]
+        sigma=lambd[1]
+        mu = lambd[0]
         flux = self.flux(rho)
-        rhs = -a*(flux[1:]-flux[:-1])/dx+D*d2udx2(rho,dx)
+        rhs = -mu*(flux[1:]-flux[:-1])/dx+0.5*sigma**2*d2udx2(rho,dx)
+      #  rhs = -mu*(flux[1:]-flux[:-1])/dx+D*d2udx2(rho,dx)
         return rhs
 
     def flux(self,rho):

### 2016-05-09 Problems/WeightedParticles/checkSystem/check_bif_sde.py
Added errorbars on NK-solution

@@ -24,7 +24,7 @@ import histogram
 import particles
 import inv_transform_sampling as inv_transform
 
-import matplotlib.pylab as plt
+#import matplotlib.pylab as plt
 
 import Point.Point as Point
 import Solver.NewtonSolver as NewtonSolver
@@ -82,7 +82,7 @@ if __name__=="__main__":
   #  lifting =  inv_transform.Sampler(sampler_param)
     
     h=2e-2 # kde mesh width     
-    M=10 #number of monte carlo steps         
+    M=2 #number of monte carlo steps         
     y_mean = scipy.zeros(M)
     y_sde = scipy.zeros(M)
 
@@ -101,7 +101,7 @@ if __name__=="__main__":
 
  #   Nlist = scipy.array([1000, 2000, 4000, 8000, 16000, 32000, 64000, 128000])
 #    Nlist = scipy.array([1e3, 1e4, 1e5, 1e6, 1e7, 1e8]) 
-    Nlist = scipy.array([1e6]) #, 800, 1600, 3200])
+    Nlist = scipy.array([1e5]) #, 800, 1600, 3200])
     eps_list_exponents=[5]
     eps_list= [1e-5]
     param['eps']= eps_list[-1]
@@ -124,6 +124,7 @@ if __name__=="__main__":
     residual2 = scipy.zeros(len(Dtlist))
     residual3 = scipy.zeros(len(Dtlist))
     residual_directsim = scipy.zeros(len(Dtlist)) 
+    save_flag= False
 
 #    
             
@@ -194,9 +195,11 @@ if __name__=="__main__":
                 newt_param = NewtonSolver.NewtonSolver.getDefaultParameters()
                 newt_param['rel_tol']=1e-5
                 if (N==1e6):
-                    newt_param['abs_tol']=45e-5
+                    newt_param['abs_tol']=45e-5     #Open question: How to set the tolerance a priori (what is the dependence of the noise on the number of particles?)
                 else:
                     newt_param['abs_tol']=140e-5  
+                if (N==1e7):
+                    newt_param['abs_tol']=14e-5
                 #newt_param['abs_tol']=43e-5
                 newt_param['print']='short'
                 newt_param['max_iter']=5
@@ -331,6 +334,30 @@ if __name__=="__main__":
 
[…]

### 2016-05-09 Problems/WeightedParticles/checkSystem/post_processing_bif.py

@@ -16,7 +16,7 @@ grid = scipy.arange(xL+dx/2.,xR,dx)
 
 #STOCHASTIC
 
-save_flag= True
+save_flag= False
 branch = branch_list[-1]
 
 bif_par = scipy.zeros( len(branch.points))
@@ -31,16 +31,16 @@ for i in range (len(branch.points)):
 plt.xlabel(r'$\sigma$', fontsize = 15)
 plt.legend()
 if save_flag:
-    plt.savefig('plots/bif/fixed_states_sde(sigma)_Ne6.pdf')
+    plt.savefig('plots/bif/fixed_states_sde(sigma)_Ne5_2.pdf')
     
 plt.show()
 
 
-for b in range ( len(branch_list)):
-    plt.plot(grid, branch_list[b].points[-1].u, label=b)
-plt.xlabel(r'$\sigma$', fontsize = 15)
-plt.show()
-    
+#for b in range ( len(branch_list)):
+#    plt.plot(grid, branch_list[b].points[-1].u, label=b)
+#plt.xlabel(r'$\sigma$', fontsize = 15)
+#plt.show()
+#    
     
 #x, y=  branch.getPlot()
 
@@ -53,7 +53,7 @@ for b in range (M):
         rho_mean[i] =  rho_mean[i] + branch_list[b].points[i].u/M
 
 if save_flag:
-    np.savetxt('bifurcation_data/fixed_states_sde_Ne6_mean_M10.txt',rho_mean)
+    np.savetxt('bifurcation_data/fixed_states_sde_Ne5_mean_M10_2.txt',rho_mean)
         
         
 for i in range (len(rho_mean)):
@@ -62,27 +62,60 @@ for i in range (len(rho_mean)):
 plt.xlabel(r'$\sigma$', fontsize = 15)
 plt.legend()
 if save_flag:
-    plt.savefig('plots/bif/fixed_states_sde(sigma)_Ne6_mean_M10.pdf')  
+    plt.savefig('plots/bif/fixed_states_sde(sigma)_Ne5_mean_M10_2.pdf')  
 plt.show()
 
 #ANALYTIC
-mu=1.0
-sigma_list= scipy.arange(1.5, 0.4, -0.1)
-sigma_list= scipy.arange(2, 0.9, -0.1)
-D_list= scipy.arange(1, 0, -0.1)
-y_anal = []
-rho_norm_D = []
-for i in range ( len(sigma_list)):
[…]

### 2016-05-09 Problems/WeightedParticles/checkSystem/post_processing_bif_obs.py

@@ -16,8 +16,8 @@ grid = scipy.arange(xL+dx/2.,xR,dx)
 
 M=10
 rho_mean = np.loadtxt('bifurcation_data/fixed_states_sde_Ne6_mean_M10.txt')
-y_mean =  np.loadtxt('bifurcation_data/norm_fixed_states_sde_Ne6_mean_M10_RL.txt')
-y_var = np.loadtxt('bifurcation_data/variance_fixed_states_sde_Ne6_mean_M10_RL.txt')
+y_mean =  np.loadtxt('bifurcation_data/norm_fixed_states_sde_Ne6_mean_M10_LR.txt')
+y_var = np.loadtxt('bifurcation_data/variance_fixed_states_sde_Ne6_mean_M10_LR.txt')
 y_error = sqrt(y_var)/sqrt(M)
 
 #STOCHASTIC
@@ -25,7 +25,8 @@ y_error = sqrt(y_var)/sqrt(M)
 save_flag= False
 N_points = len(y_mean)
 
-sigma_list= scipy.arange(2, 0.9, -0.1)
+#sigma_list= scipy.arange(2, 0.9, -0.1)   #voor RL
+sigma_list= scipy.arange(1, 2.1, 0.1)  #voor LR
         
 for i in range (N_points):
     par_label = r'$\sigma= %g $' % sigma_list[i]
@@ -33,13 +34,13 @@ for i in range (N_points):
 plt.xlabel(r'$\sigma$', fontsize = 15)
 plt.legend()
 if save_flag:
-    plt.savefig('plots/bif/fixed_states_sde(sigma)_Ne6_mean_M10.pdf')  
+    plt.savefig('plots/bif/fixed_states_sde(sigma)_Ne6_mean_M10_LR.pdf')  
 plt.show()
 
 #ANALYTIC
 mu=1.0
-sigma_list= scipy.arange(1.5, 0.4, -0.1)
-sigma_list= scipy.arange(2, 0.9, -0.1)
+#sigma_list= scipy.arange(1.5, 0.4, -0.1)
+#sigma_list= scipy.arange(2, 0.9, -0.1)
 D_list= scipy.arange(1, 0, -0.1)
 y_anal = []
 rho_norm_D = []
@@ -62,7 +63,7 @@ for i in range ( len(sigma_list_fine)):
     
 plt.plot( sigma_list_fine,y_anal_fine,  'r', markersize=4,label= 'Analytic solution')
[…]

### 2016-05-23 Documentation/draft.tex
Added errorbars on NK-solution
nice results for Newton convergence if eps_gmres is smalle enough

@@
 
-\subsection{Newton-Krylov solver} \label{sec:Newton-Krylov}
+\subsection{Fixed-point algorithm}
+
+\subsubsection{Newton-Krylov solver} \label{sec:Newton-Krylov}
 
@@
 
-\input{GMRES}
-
-\subsection{Variance reduced Jacobian-vector products}
+\subsubsection{Variance reduced Jacobian-vector products}
 
@@
 \section{Analysis}
+
+Numerical experiments that show how bias and variance of the Jacobian-vector product depend on the number of particles used.
+
+Watch out: as the reference solution we use the deterministic Jacobian-vector product.
+That, in itself, has a discretisation error that should not be interpreted as a bias of the stochastic Jacobian-vector product.
+
 \subsection{Variance}
@@
 
+\subsection{bias}
+
 The convergence of the density to the real solution, strongly depends on the number of particles $N$ and on the choice of $\epsilon$ in the GMRES-method.
 
-\subsection{}
-
-\subsection{bias}
+\input{GMRES}
 
@@
 \begin{figure}
-\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
+%\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
 \caption{  Bifurcation diagram of the steady states calculated with the Newton-Krylov-Solver for the PDE.
@@
 \begin{figure}
-\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
+%\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
 \caption{  Bifurcation diagram of the steady states calculated with the Newton-Krylov-Solver for the SDE.

### 2016-05-23 Documentation/newton_convergence.tex
nice results for Newton convergence if eps_gmres is smalle enough

@@
-Numerical experiments that show how the Newton method converges to the fixed point:
+We show some numerical experiments that show how the Newton method converges to the fixed point.
 
-What is the maximal accuracy that can be obtained? (Relate to the variance of the coarse time-stepper.) How to cheaply estimate the stopping criterion?
-Is the resulting fixed point biased?
-Can we reduce variance by averaging over a number of 'converged' Newton iterations?
-Additional questions of interest:
+\subsection{Maximal accuracy}
 
-Can we stop the Krylov iterations before reaching machine precision (knowing that the end result will contain noise anyway)?
-Does an inaccuracy in the solution of the linear systems increase the required number of Newton iterations?
-Do we keep the particles fixed throughout the Newton iterations as well?
+The accuracy on the Newton-Krylov solution is inevitably limited by the noise on the stochastic coarse-time-stepper, which appears in the evaluation of the residual and in the evaluation of the Jacobian-vector-product.
+The variance on the coarse-time-stepper is of order $\mathcal{O}(\frac{1}{N})$, as shown in fig.
+\ref{fig:Variance_on_cts(N-1)}.
+If we use weighted restriction, the variance on the Jacobian-vector-product converges in the same way.
+Therefore, the accuracy on the Newton-Krylov solution will depend on the number of particles used to calculate the coarse time-stepper.
+This is confirmed in fig.
+\ref{fig:Newton_sde_res(k)}, which shows the convergence of the Newton-residuals.
+After the Newton-Krylov solution is converged, it stays oscillating around the true solution with a standard deviation depending on the number of particles.
+Because the solution only depends on the previous density and on the random choices in the coarse time stepper, we can interpret this as a Markov chain.
+The residual averaged out over the converged Newton steps is a good indication of the best tolerance we can achieve.
+It converges as $\mathcal{O}(\frac{1}{\sqrt{N}})$.
+Given the number of particles $N$,  we will use this tolerance as the stopping criterion for the Newton-method.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\linewidth]{../Problems/WeightedParticles/checkSystem/plots/Variance_on_cts(N-1)}
+\caption{The variance of the coarse time-stepper is of $\mathcal{O}(\frac{1}{N})$ }
+\label{fig:Variance_on_cts(N-1)}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Newton_sde_res(k)_Dt_e-2_tol1e-7.pdf}
+\includegraphics[width=0.49\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Tolerance_on_NK-solution_converges_N-1_tol_1e-7.pdf}
+
+\caption{ The non-linear solver converges after 1 or 2 Newton iterations, up to a tolerance which depends on the number of particles $N$ (\textit{left}).
+This best achieved tolerance converges to zero with  $\mathcal{O}(\frac{1}{\sqrt{N}})$ (\textit{right}).
+Parameter values:  $\Delta t = 10^{-3}, \Delta T = 10^{-1} , \epsilon_{\texttt{GMRES}}=10^{-7}$
+}.
+\label{fig:Newton_sde_res(k)}
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.5\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Newton_sde_res(k)_Dt_e-2_tol1e-2}
+\includegraphics[width=0.49\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Tolerance_on_NK-solution_converges_N-1}
+
+\caption{ The best achieved tolerance is limited because of the size of the GMRES-tolerance ($\epsilon_{\texttt{GMRES}}=10^{-4}$)
+}
+\label{fig:Newton_sde_res(k)_tol1e-4}
+\end{figure}
+
+%->Insert result N=1e7 (on the cluster now)
+
+%How to cheaply estimate the stopping criterion?
+%Is the resulting fixed point biased?
+%Can we reduce variance by averaging over a number of 'converged' Newton iterations?
+%Additional questions of interest:
+%
+%Can we stop the Krylov iterations before reaching machine precision (knowing that the end result will contain noise anyway)? 
+%
+%Does an inaccuracy in the solution of the linear systems increase the required number of Newton iterations?
+%
+%Do we keep the particles fixed throughout the Newton iterations as well?
 (We can then probably converge the Newton procedure to machine precision.) What is then the probability distribution of the obtained fixed points?
@@
 Variance?
+
+%\subsection{Results for the pde}
+%Ook hier zien we dat de keuze van de newtontolerantie belangrijk is (te groot = kans op dezelfde toestanden , te klein= failed iterations).
+%Dit kan echter opgevangen worden door de oplossing van dx nauwkeuriger te maken door de tolerantie van de GMRES te verfijnen (het probleem was dat er dx=0 oplossingen werden teruggegeven)
+
+%\begin{figure}
+%%\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
+%\caption{  Bifurcation diagram of the steady states calculated with the Newton-Krylov-Solver for the PDE.
+$\Delta t = 10^{-4}, \Delta T = 10^{-2}, \Delta x = 10^{-2}, \Delta D = 0.01, \epsilon_{GMRES}=10^{-5},  \delta_{Newton} = 10^{-7}, \epsilon=10^{-5}$
+%}
+%\end{figure}
+
+%\subsection{Results for the sde}
+
+\subsection{Continuation}
+ We start from a known solution to the steady-state problem, increment the continuation parameter and solve a new problem using the previous solution as an initial guess.
+We chose $\Delta \sigma = 0.05$ as continuation step seize.
+Performing a bifurcation with smaller steps is pointless, because the difference in densities will become too small in comparison with the noise on the residual.
+
+\begin{figure}[h]
+\centering
+%\begin{subfigure}{0.5\textwidth}
+\includegraphics[width=0.49\linewidth]{../Problems/WeightedParticles/checkSystem/plots/bif/fixed_states_sde(sigma)_Ne6_mean_M10_LR}
+\includegraphics[width=0.49\linewidth]{../Problems/WeightedParticles/checkSystem/plots/bif/fixed_states(sigma)_analytic}
+\includegraphics[width=0.9\linewidth]{../Problems/WeightedParticles/checkSystem/plots/bif/bifurcation_sde_Ne6_anal(sigma)_LR}
+\caption{The steady states calculated with the Newton-Krylov-solver for the SDE  (\textit{top left}) are compared with the analytical solutions (\textit{top right}).
+The fixed points are visualized as the 2-norm of the density, and plotted as a function of the continuation parameter $\sigma$ (\textit{bottom}).
+The branch of fixed points appears to be in correspondence with the 2-norm of the analytic function $\rho^*(\sigma) =  \exp{\left[-\frac{2 V(x)}{\sigma^2}\right]} $.
+%{\mathcal{N}} $.
+Parameter values: $\Delta \sigma =0.1$, $N=10^6$, $M=10$, $\Delta t = 10^{-3}, \Delta T = 10^{-1}, \Delta x = 10^{-2}, \epsilon_{\texttt{GMRES}}=10^{-5},  \delta_{\texttt{Newton}} = 45 \cdot 10^{-5}, \varepsilon_J=10^{-5}$}.
+\label{fig:fixed_states_sde(sigma)_Ne6_mean_M10_LR}
+\end{figure}
+
+%
+%\subsection{timings}
+%Simulation time for solving sde, ten Newton iterations, with N=1e8, tol=1e-7: Time = 34u26m15s
+%

### 2016-05-23 Problems/WeightedParticles/checkSystem/check_bif_pde.py

@@ -196,7 +196,7 @@ if __name__=="__main__":
   #  plt.plot(Dtlist, residual2, "r--" )
     plt.plot(Dtlist, residual_directsim, "c--" )
     plt.xlabel(r'$\Delta T$', fontsize = 14)
-    plt.ylabel(r'$||  \rho^{*}   - \frac{ \exp{\left[-\frac{V(x)}{\sigma^2}\right]}}{\mathcal{N}}  ||$'  , fontsize = 15)
+    plt.ylabel(r'$||  \rho^{*}   - \frac{ \exp{\left[-2\frac{V(x)}{\sigma^2}\right]}}{\mathcal{N}}  ||$'  , fontsize = 15)
    # plt.savefig("plots/Newton_pde_res(Dt)_dt=1e-5.pdf")
               
     

### 2016-05-23 Problems/WeightedParticles/checkSystem/check_newton_sde.py

@@ -44,7 +44,7 @@ if __name__=="__main__":
     D = 2.0
     sigma=1.0
    # sigma = 2*np.sqrt(D)
-    Dt = 1e-1  #Dt influences number of GMRES-iterations strongly
+    Dt = 1e-2  #Dt influences number of GMRES-iterations strongly
     Dtlist = [ Dt]
     #Dtlist = [ 5e-3, 1e-2, 2e-2, 5e-2, 1e-1, 2e-1, 5e-1, 1]
     #Dtlist = [1e-5, 5e-5, 1e-4,  5e-4, 1e-3,  5e-3 , 1e-2, 5e-2 , 1e-1, 5e-1, 1]
@@ -82,7 +82,7 @@ if __name__=="__main__":
   #  lifting =  inv_transform.Sampler(sampler_param)
     
     h=2e-2 # kde mesh width     
-    M=2 #number of monte carlo steps         
+    M=1 #number of monte carlo steps         
     y_mean = scipy.zeros(M)
     y_sde = scipy.zeros(M)
 
@@ -100,8 +100,8 @@ if __name__=="__main__":
    # fp_sde = particles.Particles(lifting,restriction,rho,grid, lambd, param=param)       
 
  #   Nlist = scipy.array([1000, 2000, 4000, 8000, 16000, 32000, 64000, 128000])
-#    Nlist = scipy.array([1e3, 1e4, 1e5, 1e6, 1e7, 1e8]) 
-    Nlist = scipy.array([1e5]) #, 800, 1600, 3200])
+    Nlist = scipy.array([ 1e5]) 
+  #  Nlist = scipy.array([1e5]) #, 800, 1600, 3200])
     eps_list_exponents=[5]
     eps_list= [1e-5]
     param['eps']= eps_list[-1]
@@ -116,23 +116,24 @@ if __name__=="__main__":
     sq_E_Jv=scipy.zeros(nN)
     E_rho= scipy.zeros((nN,len(rho)))
     E_Jv= scipy.zeros((nN,len(rho)))
+    
    # = scipy.zeros((nN,len(rho)))
     
     
         
-    residual1 = scipy.zeros(len(Dtlist))
-    residual2 = scipy.zeros(len(Dtlist))
-    residual3 = scipy.zeros(len(Dtlist))
-    residual_directsim = scipy.zeros(len(Dtlist)) 
+#    residual1 = scipy.zeros(len(Dtlist))
+#    residual2 = scipy.zeros(len(Dtlist))
[…]

### 2016-05-23 Problems/WeightedParticles/checkSystem/check_sde.py

@@ -35,7 +35,7 @@ import Utils.conditions.probability as probability
 if __name__=="__main__":
 
     sigma = 1
-    Dt = 1
+    Dt = 1e-1
     seed = 16
     D=0.5*sigma**2
     # discretization parameters
@@ -57,8 +57,8 @@ if __name__=="__main__":
     dx = 1e-2
 #    factor = int (dx/dx_fp)
 #    print "Discretisation for solving sde is ", factor, " times coarser than the discretisation for solving the pde"
-    dt = 1e-2 # ti
-    N=1e5
+    dt = 1e-3 # ti
+    N=1e6
     Nlarge= N
     Nsmall =N
     
@@ -76,7 +76,7 @@ if __name__=="__main__":
   #  lifting =  inv_transform.Sampler(sampler_param)
     
     h=2e-2 # kde mesh width     
-    M=1 #number of monte carlo steps                
+    M=10 #number of monte carlo steps                
     param_histogram = histogram.Histogram.getDefaultParameters()
     param_histogram['h']=h
     restriction = histogram.Histogram(param_histogram)
@@ -89,8 +89,8 @@ if __name__=="__main__":
 #    param['eps']=1e-5
    # fp_sde = particles.Particles(lifting,restriction,rho,grid, lambd, param=param)       
 
-  #  Nlist = scipy.array([1000, 2000, 4000, 8000, 16000, 32000, 64000, 128000])
-    Nlist = scipy.array([N]) #, 800, 1600, 3200])
+    Nlist = scipy.array([1e3, 1e4, 1e5, 1e6, 1e7])
+  #  Nlist = scipy.array([N]) #, 800, 1600, 3200])
     eps_list_exponents=[5]
     eps_list= [1e-5]
    # Nlist = scipy.array([5,10, 200]) #,400])
@@ -139,14 +139,14 @@ if __name__=="__main__":
                 rg_state = lifting.get_rg_state()  #We remember the internal state of the random generator to get new random number for sampling the new particles
                 
                 print "calculate .u_Dt"
-                rho_Dt_sde = fp_sde.u_Dt 
+                rho_Dt = fp_sde.u_Dt 
                 t0=time.time()
                 print "Simulation time for solving sde: " , t0-t1, " seconds"
                 #rho_fine = fp_sde.make_rho_fine(rho_Dt) 
[…]

### 2016-05-23 Problems/WeightedParticles/checkSystem/post_processing_Newton.py

@@ -31,16 +31,9 @@ if __name__=="__main__":
 
 #    resnorm_t100_N1000 = np.loadtxt('Newton/resnorm_Dte-1_N1000')
 #    resnorm_t100_N10000 = np.loadtxt('Newton/resnorm_Dte-1_N10000') 
-#    resnorm_t100_N100000 = np.loadtxt('Newton/resnorm_Dte-1_N100000') 
-#    resnorm_t100_N1000000 = np.loadtxt('Newton/resnorm_Dte-1_N1000000')  
-#    resnorm_t100_N10000000= np.loadtxt('Newton/resnorm_Dte-1_N10000000')  
-    
-    resnorm_t10_tol2_N1000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N1000')
-    resnorm_t10_tol2_N10000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N10000') 
-    resnorm_t10_tol2_N100000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N100000') 
-    resnorm_t10_tol2_N1000000 = np.loadtxt('Newton/resnorm_Dte-2_tole-2_N1000000')  
-    resnorm_t10_tol2_N10000000= np.loadtxt('Newton/resnorm_Dte-2_tole-2_N10000000')  
-
+    resnorm_t100_N100000 = np.loadtxt('Newton/resnorm_Dte-1_N100000') 
+    resnorm_t100_N1000000 = np.loadtxt('Newton/resnorm_Dte-1_N1000000')  
+    resnorm_t100_N10000000= np.loadtxt('Newton/resnorm_Dte-1_N10000000')  
     
     resnorm_t10_N1000 = np.loadtxt('Newton/resnorm_Dte-2_N1000')
     resnorm_t10_N10000 = np.loadtxt('Newton/resnorm_Dte-2_N10000') 
@@ -48,26 +41,32 @@ if __name__=="__main__":
     resnorm_t10_N1000000 = np.loadtxt('Newton/resnorm_Dte-2_N1000000')  
     resnorm_t10_N10000000= np.loadtxt('Newton/resnorm_Dte-2_N10000000')  
     
+ 
     
-    
+        
+   # resnorm_t10_tol4_N10000 = np.loadtxt('Newton/17-05_resnorm_Dte-2_tole-4_N10000') 
+    resnorm_t10_tol4_N100000 = np.loadtxt('Newton/17-05_resnorm_Dte-2_tole-4_N100000')  
[…]

### 2016-05-23 Problems/WeightedParticles/checkSystem/post_processing_bif.py

@@ -196,7 +196,7 @@ plt.errorbar(x_sde, y_mean, yerr=y_error,   fmt='bo', ecolor='b', markersize=3, 
 plt.ylabel(r'$\left\||\rho^*\right\||$', fontsize = 15)
 plt.xlabel(r'$\sigma$', fontsize = 11)
 plt.xlim([0.9,2.1])
-plt.legend(prop={'size':8})
+
 #plt.plot(x_sde, y_sde, 'go', markersize=4)
 
 if save_flag:

### 2016-05-23 Problems/WeightedParticles/checkSystem/post_processing_bif_obs.py

@@ -22,17 +22,22 @@ y_error = sqrt(y_var)/sqrt(M)
 
 #STOCHASTIC
 
-save_flag= False
+save_flag= True
 N_points = len(y_mean)
 
 #sigma_list= scipy.arange(2, 0.9, -0.1)   #voor RL
 sigma_list= scipy.arange(1, 2.1, 0.1)  #voor LR
         
 for i in range (N_points):
-    par_label = r'$\sigma= %g $' % sigma_list[i]
-    plt.plot(grid, rho_mean[i], label=par_label)   
-plt.xlabel(r'$\sigma$', fontsize = 15)
-plt.legend()
+    if (i % 2 == 0): 
+        par_label = r'$\sigma= %g $' % sigma_list[i]
+        plt.plot(grid, rho_mean[i], label=par_label)   
+plt.xlim([-1.7,1.7])
+plt.ylim([0,0.6])
+plt.xlabel('$x$', fontsize = 12)
+plt.ylabel(r'$\rho^*$', fontsize = 12)
+plt.legend(bbox_to_anchor=(1,0.6),  numpoints = 1 )
+plt.legend(prop={'size':9})
 if save_flag:
     plt.savefig('plots/bif/fixed_states_sde(sigma)_Ne6_mean_M10_LR.pdf')  
 plt.show()
@@ -48,6 +53,20 @@ for i in range ( len(sigma_list)):
     norm_c = sum( np.exp( (-grid**4 + grid**2)*2*mu/(sigma_list[i]**2)))*dx
     rho_ss = np.exp( (-grid**4 + grid**2 )*2*mu/(sigma_list[i]**2)) /norm_c
     y_anal.append(norm(rho_ss))
+    if (i % 2 == 0): 
+        par_label = r'$\sigma= %g $' % sigma_list[i]
+        plt.plot(grid, rho_ss, label=par_label)
+
+plt.xlim([-1.7,1.7])
+plt.ylim([0,0.6])
+plt.xlabel('$x$', fontsize = 12)
+plt.ylabel(r'$\rho^*$', fontsize = 12)
+plt.legend(bbox_to_anchor=(1,0.6),  numpoints = 1 )
+plt.legend(prop={'size':9})
+if save_flag:
+    plt.savefig('plots/bif/fixed_states(sigma)_analytic.pdf')  
+plt.show()
+
 
 sigma_list_fine= scipy.arange(2.1, 0.9, -0.001)
 y_anal_fine = []
[…]

### 2016-05-23 Problems/WeightedParticles/checkSystem/post_processing_eps.py

@@ -51,7 +51,7 @@ if __name__=="__main__":
  
     #rho_Dt_sde = E_rho[-1]
     Jv_sde = E_Jv[-1][-1]  #dim n_x
-
+    bins = len(Jv_sde)
     
     resize_factor = int (len(rho_Dt_fp)/len(Jv_sde))
   #  resize_factor = 10
@@ -77,8 +77,6 @@ if __name__=="__main__":
 #            bin_av = bin_av+ rho_Dt_fp[i*resize_factor+j]
 #        rho_coarse[i] = bin_av/resize_factor
             
-
-    bins = len(Jv_sde)
     
     Jv_coarse = scipy.zeros((len(eps_list),bins))
     for eps_i in range (0,len(eps_list)):

### 2016-05-23 Problems/WeightedParticles/checkSystem/post_processing_variance.py

@@ -1,0 +1,115 @@
+"""
+Created on Wed Oct 28 10:05:37 2015
+
+@author: pieter
+"""
+import time 
+import tables
+import scipy
+import sys
+import numpy as np
+import matplotlib.pylab as plt
+
+
+from scipy.linalg import norm
+from scipy import sqrt
+
+import matplotlib.pylab as plt
+plt.rc('text', usetex=True)
+
+if __name__=="__main__":
+    D = 1./2
+    Dt = 1e-2    
+    a = 1
+    zeta = 0.
+    alpha=1
+    beta = -1
+    lambd = scipy.array([a,D,alpha,beta,zeta])
+   # param['eps'] = 1e-5
+    #param['Dt'] = Dt
+    #param['dt']=dt         
+
+    
+    #SDE       
+#    rho_sq = np.loadtxt('data/obs/25-11-rho_sq_eps-5.out')
+#    E_rho = np.loadtxt('data/obs/25-11-E_rho_eps-5.out')
+#    sq_E_rho  =np.loadtxt('data/obs/25-11-sq_E_rho_eps-5.out')
+#    
+    rho_sq = np.loadtxt('data/09-05-rho_sq_N=e^3-7.out')
+    E_rho = np.loadtxt('data/09-05-E_rho_N=e^3-7.out')
+    #sq_E_rho  =np.loadtxt('data/09-05-sq_E_rho.out')
+    sq_E_rho  =np.loadtxt('data/09-05-sq_E_rho_N=e^3-7.out')
+  
+    
+#            np.savetxt('data/09-05-rho_sq.out' , rho_sq)
+#        np.savetxt('data/09-05-E_rho.out' , E_rho )
+#        np.savetxt('data/09-05-sq_E_rho.out',  sq_E_rho )
+        
+     
+    Nlist = scipy.array([1000, 2000,4000,8000,16000,32000,64000])
+    Nlist = scipy.array([1e3, 1e4, 1e5, 1e6, 1e7])
+    Nlist_inv = 1.0/Nlist
+    
+    log_flag = 'log'
+  #  log_flag = 'linear'
+    save_flag = False
+ 
+
+    bins = len(E_rho[-1])
+    
+    variance = (rho_sq - sq_E_rho)
+    sdev = sqrt(variance)
+       
+    plot_var = plt.plot(Nlist_inv, variance, label = '$Variance$')
[…]

### 2016-05-24 Documentation/newton_convergence.tex
added nice results for Newton convergence if eps_gmres is smalle enough

@@
 
+\subsubsection{Estimating stopping criterion for the non-linear solver}
 The accuracy on the Newton-Krylov solution is inevitably limited by the noise on the stochastic coarse-time-stepper, which appears in the evaluation of the residual and in the evaluation of the Jacobian-vector-product.
@@
 If we use weighted restriction, the variance on the Jacobian-vector-product converges in the same way.
-Therefore, the accuracy on the Newton-Krylov solution will depend on the number of particles used to calculate the coarse time-stepper.
+Therefore, the accuracy on the Newton-Krylov solution will depend on the number of particles $N$ used to calculate the coarse time-stepper.
 This is confirmed in fig.
 \ref{fig:Newton_sde_res(k)}, which shows the convergence of the Newton-residuals.
-After the Newton-Krylov solution is converged, it stays oscillating around the true solution with a standard deviation depending on the number of particles.
+When the Newton-Krylov solution is converged, it stays oscillating around the true solution with a standard deviation depending on the number of particles.
 Because the solution only depends on the previous density and on the random choices in the coarse time stepper, we can interpret this as a Markov chain.
-The residual averaged out over the converged Newton steps is a good indication of the best tolerance we can achieve.
+The residual averaged out over the converged Newton iterations is a good indication of the best tolerance we can achieve.
 It converges as $\mathcal{O}(\frac{1}{\sqrt{N}})$.
@@
 
-\begin{figure}[h]
+\subsubsection{Estimating stopping criterion for the linear solver}
+Because the solution of the Newton-Krylov solution will contain noise anyway, it is pointless to iterate the linear solver until reaching machine precision.
+In fact, for low particle numbers (and a high variance on the left and right side of the linear equation we need to solve), we observe more failed Newton-iterations if the GMRES-tolerance is chosen too small.
+A question of interest is then `what is a good value for the GMRES-tolerance?'.
+The answer depends on the desired accuracy of the calculated fixed points.
+For instance in fig.
+\ref{fig:Newton_sde_res(k)_tol1e-4}, we observe that we  can already stop the Krylov iterations if the density is calculated to an accuracy of %$\varepsilon_{\texttt{GMRES}}=
+$10^{-4}$ when $N \leq 10^7$, but that we need a smaller GMRES-tolerance if a higher accuracy is desired (by using more particles).
+Therefore, in the simulations we  will use $\varepsilon_{\texttt{GMRES}}= 10/N$.
+
+\begin{figure}[H!]
 \centering
-\includegraphics[width=0.5\linewidth]{../Problems/WeightedParticles/checkSystem/plots/Variance_on_cts(N-1)}
-\caption{The variance of the coarse time-stepper is of $\mathcal{O}(\frac{1}{N})$ }
+\includegraphics[width=0.49\linewidth]{../Problems/WeightedParticles/checkSystem/plots/Variance_on_cts(N-1)}
+\includegraphics[width=0.49\linewidth]{../Problems/WeightedParticles/checkSystem/plots/Variance_on_Jv}
+\caption{The variance of the coarse time-stepper and the variance on the Jacobian vector-product converge to zero with $\mathcal{O}(\frac{1}{N})$ }
 \label{fig:Variance_on_cts(N-1)}
@@
 
-\begin{figure}[h]
+\begin{figure}[H!]
 \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[H!]
 \centering
-\includegraphics[width=0.5\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Newton_sde_res(k)_Dt_e-2_tol1e-2}
-\includegraphics[width=0.49\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Tolerance_on_NK-solution_converges_N-1}
+\includegraphics[width=0.5\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Newton_sde_res(k)_Dt_e-2_tol1e-4}
+\includegraphics[width=0.49\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Tolerance_on_NK-solution_converges_N-1_tol_1e-4}
 
-\caption{ The best achieved tolerance is limited because of the size of the GMRES-tolerance ($\epsilon_{\texttt{GMRES}}=10^{-4}$)
+\caption{ The best achieved tolerance for $N=10^8$ is limited because of the size of the GMRES-tolerance ($\epsilon_{\texttt{GMRES}}=10^{-4}$)
 }
@@
 
-%->Insert result N=1e7 (on the cluster now)
+%How to cheaply estimate the stopping criterion?
 
-%How to cheaply estimate the stopping criterion?
 %Is the resulting fixed point biased?
@@
  We start from a known solution to the steady-state problem, increment the continuation parameter and solve a new problem using the previous solution as an initial guess.
-We chose $\Delta \sigma = 0.05$ as continuation step seize.
+We chose $\Delta \sigma = 0.05$ as continuation step size.
 Performing a bifurcation with smaller steps is pointless, because the difference in densities will become too small in comparison with the noise on the residual.

### 2016-05-24 Problems/WeightedParticles/checkSystem/post_processing_Newton.py

@@ -41,9 +41,6 @@ if __name__=="__main__":
     resnorm_t10_N1000000 = np.loadtxt('Newton/resnorm_Dte-2_N1000000')  
     resnorm_t10_N10000000= np.loadtxt('Newton/resnorm_Dte-2_N10000000')  
     
- 
-    
-        
    # resnorm_t10_tol4_N10000 = np.loadtxt('Newton/17-05_resnorm_Dte-2_tole-4_N10000') 
     resnorm_t10_tol4_N100000 = np.loadtxt('Newton/17-05_resnorm_Dte-2_tole-4_N100000')  
     resnorm_t10_tol4_N1000000= np.loadtxt('Newton/17-05_resnorm_Dte-2_tole-4_N1000000')  
@@ -104,7 +101,7 @@ if __name__=="__main__":
   #  tol3Artist = plt.Line2D((0,1),(0,0), color='k', linestyle='--')
   #  tol2Artist = plt.Line2D((0,1),(0,0), color='k', linestyle='-')
     
-    ax.legend( numpoints = 1,  prop={'size':12} )
+    ax.legend( numpoints = 1,  prop={'size':8} )
 #    ax.legend([handle for i,handle in enumerate(handles) if i in display]+[tol3Artist,tol2Artist],
 #          [label for i,label in enumerate(labels) if i in display]+[r'$\epsilon_{GMRES} = 10^{-5}$', r'$\epsilon_{GMRES} = 10^{-2}$'], bbox_to_anchor=(1.39
 #          , 1), prop={'size':10})
@@ -128,6 +125,7 @@ if __name__=="__main__":
     
     points_o1var = [[1e-3, 1e-3], [1e-3, 5e-3], [5e-3, 5e-3]]   #label if  plotting as a function of 1/sqrt(N)
     points_o1var = [[1e-6, 1e-3], [1e-6, 2e-3], [4e-6, 2e-3]] 
+    points_o1var = [[1e-7, 1e-3], [1e-7, 3e-3], [9e-7, 3e-3]] 
     order= r'$\mathcal{O}(1/\sqrt{N})$'
 
  #   points_o1var = [[1e-5, 1e-2], [2e-5, 2e-2], [1e-5, 2e-2]] 
@@ -142,8 +140,8 @@ if __name__=="__main__":
     plt.gca().add_patch(triangle_o1var)
     plt.ylabel(r'$||  \rho^* - \Phi^N_T(\rho^*)||$'  , fontsize = 15)
[…]

### 2016-05-24 Problems/WeightedParticles/checkSystem/post_processing_eps.py

@@ -101,7 +101,7 @@ if __name__=="__main__":
     
     log_flag = 'log'
   #  log_flag = 'linear'
-    save_flag = True
+    save_flag = False
     
    # Nlist = scipy.array([1000,2000,4000,8000,16000])
        
@@ -156,6 +156,21 @@ if __name__=="__main__":
     order= r'$\mathcal{O}(1/N)$'
     plt.annotate(order,  xy=(2e-4, 2e-2), xytext=(1.1e-4, 5.2e-2), fontsize=11, color='grey')
     if(save_flag): plt.savefig("plots/Var_N_eps.pdf")
+    plt.savefig("plots/Variance_on_Jv.pdf")
+    plt.show()
+    
+        
+    points_o1var = [[1e-4, 2e-2], [1e-4, 1e-1], [5e-4, 1e-1]] 
+    triangle_o1var = plt.Polygon(points_o1var, fill=None  ,edgecolor='grey') 
+    plot_var = plt.plot(Nlist_inv, (sde_Jv_sq[1] -sq_E_Jv[1])/bins, linewidth=2)
+    plt.ylabel('Var ($\mathbf{\hat{Jv}} $)', fontsize = 12)
+    plt.xlabel('$1/N$', fontsize = 12) 
+    plt.xscale(log_flag)
+    plt.yscale(log_flag)
+    plt.gca().add_patch(triangle_o1var)
+    order= r'$\mathcal{O}(1/N)$'
+    plt.annotate(order,  xy=(2e-4, 2e-2), xytext=(1.1e-4, 5.2e-2), fontsize=11, color='grey')
+    plt.savefig("plots/Variance_on_Jv.pdf") #This is used to refer to in Newton convergence analysis
     plt.show()
     
      

### 2016-05-24 Problems/WeightedParticles/checkSystem/post_processing_variance.py

@@ -59,12 +59,19 @@ if __name__=="__main__":
     
     variance = (rho_sq - sq_E_rho)
     sdev = sqrt(variance)
-       
-    plot_var = plt.plot(Nlist_inv, variance, label = '$Variance$')
+    
+    points_o1var = [[1e-6, 2e-2], [1e-6, 1e-1], [5e-6, 1e-1]] 
+ #   points_o1var = [[1e-5, 1e-2], [2e-5, 2e-2], [1e-5, 2e-2]] 
+    triangle_o1var = plt.Polygon(points_o1var, fill=None  ,edgecolor='grey') 
+        
+    plot_var = plt.plot(Nlist_inv, variance, label = '$Variance$', linewidth=2)
     plt.ylabel(r'Var$(\Phi^N_T(\rho))$' , fontsize = 12)
     plt.xlabel('$1/N$', fontsize = 12)
     plt.xscale(log_flag)
     plt.yscale(log_flag)
+    plt.gca().add_patch(triangle_o1var)
+    order= r'$\mathcal{O}(1/N)$'
+    plt.annotate(order,  xy=(2e-6, 2.5e-2), xytext=(1.1e-6, 5.7e-2), fontsize=11, color='grey')
     plt.savefig('plots/Variance_on_cts(N-1).pdf')
     plt.show()
     

### 2016-07-04 Documentation/draft.tex
last bifurcation results

@@
 
+(Under construction)
 \begin{abstract}
@@
 The main novelty of the algorithm is in the elimination of the noise that is generated when estimating Jacobian-vector products using time-integration of perturbed initial conditions.
-We present numerical results that demonstrate the convergence properties of the numerical method, and use the method to show   the emergence of coarse equilibrium state
+We present numerical results that demonstrate the convergence properties of the numerical method, and use the method to show  the emergence of coarse equilibrium states.
 
@@
 
+\section{Convergence of the Newton Method}
+\input{newton_convergence.tex}
+
 \section{Application: Systemic Risk}
 
+\input{systemic_risk.tex}
 %\subsection{Including an interaction term}
-
-Until now, we considered a system of non-interacting particles.
-We will now extend our model to a mean field model by introducing a third parameter $\alpha$, which is the degree of interaction or cooperation in the system.
-A simple form of cooperative behaviour is the case where each agent tends to follow the state of the majority (or, each particle feels an attractive force towards the mean state of the system).
-To include this cooperation effect in our stochastic simulation, we add this mean reversion term to the SDE \eqref{SDE}:
-
-\begin{equation} 
-\label{SDE_meanfield}
-    \dd x = \mu V(x) \dd t + \sigma  \dd{W_t} + \alpha(\bar{x} -x) \dd t ,
-\end{equation}
-with  $\bar{x}(t) = \frac{1}{N} \sum_{i=1}^{N} x_i(t)$ denoting the empirical mean.
-
-An interesting application of this model to banks and insurance is the emergence of systemic risk.
-Banks will try to minimize their own individual risk by spreading the risk between each other.
-However, this may increase the risk that they may all fail: reducing individual risk on a micro-scale can increase systemic risk on a macro-scale.
-Garnier, Papanicolaou and Yang already made use of the dynamics in eq.
-\eqref{SDE_meanfield}  to show that interconnectedness between agents indeed affects the stability of the whole system, causing systemic risk  \cite{Garnier}.
-They defined $x_i$  as the state of risk of agent $i$.
-The bi-stable-state structure of the potential $V(x)$ ensures that each risk variable stays around $-1$ (defined as the normal state) or $+1$ (the failed state).
-A natural measure of systemic risk is then the transition probability of the empirical mean $\bar{x}$ from the normal state to the failed state.
-
-To establish the idea, let us repeat the numerical simulations with eq.
-\ref{SDE_meanfield}.
-The evolution of the system is now characterized by  by the initial conditions, the three parameters ($\mu$, $\sigma$, $\alpha$) and by the system size $N$.
-Figure \ref{fig:sysrisk}  illustrates the behaviour of the empirical mean $\bar{x}$.
-The simulations were performed with all agents initially in the normal state.
-Nevertheless, if randomness dominates the interaction, the agents can move immediately to the other potential well.
-The system then behaves like $N$ independent diffusions, and hence, by the symmetry of the potential, the mean state will be attracted to a single mixed state $\bar{x}=0$.
-Upon increasing the interaction parameter $\alpha$, however, we find two new macroscopic states, suggesting the precence of a pitchfork bifurcation at the macroscopic level.
-These solutions are no stable steady states, but rather coarse metastable states.
-Their lifetime is linked to the finite system size.
-
-\begin{center}
-\begin{figure}
-\includegraphics[width=0.5\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha1}
-\includegraphics[width=0.5\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha5}
-
-\caption{The empirical mean, simulated for different $\alpha$. \textit{Left}: the system has one single state $\bar{x}=0$. \textit{Right}: for $\alpha> \alpha_c$ two metastable equilibria emerge. 
- The red dashed lines are approximated  analytical solutions for the steady states. \label{fig:sysrisk}}
-\end{figure}
-\end{center}
-
- \begin{center}
-\begin{table}
-\caption{Parameter values}
-  \begin{tabular} { | c  c | c |}    \hline   
-     %FTCS-scheme  &   Euler-Maruyama-scheme  \\  
-   \textit{ {Discretization parameters}}&    &  SDE    \\ \hline
-    Discretization step  & $\Delta x $  & $10^{-2}$ \\ 
-        Number of discretization steps  & $n_x$ &  340 \\ 
-    Time step  &  $\Delta t$ &   $10^{-2}$ \\ 
-       Number of timesteps  & $n$  &  $10^{6}$ \\ \hline
-  \end{tabular}
-%\quad
-%     \begin{tabular}  { | c | c | c |  }  \hline
-%     \multicolumn{3}{|c|} {\textit{ System parameters}   }    \\
-%    \hline  
-%    Diffusion coefficient  & $D $ & $0.5$ \\ 
-%   Drift coefficient &  $\mu$ & $ 1$ \\ \hline
-%  \end{tabular}
-\end{table}
-\end{center}
-
-\subsection{Bifurcation study}
-
-\subsubsection{Nota's}
-Tijdens elke continueringsstap, wordt een nieuw punt gecreëerd dat initieel de dichtheid \U krijgt van het vorig punt en met een parameter (D) die horizontaal word geupdate.
-Dan wordt de norm van het residu berekend $\norm{ \U(t) - \U(t+\Delta t)} $, waarin de dichtheid op de volgende tijdsstap verkregen wordt door een simulatie van het systeem met de nieuwe parameter.
-De grootte van het residu is dus afhankelijk van de stapgrootte in de bifurcatieparamter.
-De newtontolerantie is de norm van het residu.
-Als de newton-tolerantie te groot gekozen wordt (in vergelijking met de stapgrootte van het residu), dan wordt er geen Newton-iteratie uitgevoerd.
-Als de Newton-tolerantie evenwel te klein gekozen wordt, dan bereikt de Newton-solver mogelijk geen convergentie ten gevolge van de ruis.
-De Newton-tolerantie moet dus weloverwogen gekozen worden.
-
-\subsection{Results for the pde}
-Ook hier zien we dat de keuze van de newtontolerantie belangrijk is (te groot = kans op dezelfde toestanden , te klein= failed iterations).
-Dit kan echter opgevangen worden door de oplossing van dx nauwkeuriger te maken door de tolerantie van de GMRES te verfijnen (het probleem was dat er dx=0 oplossingen werden teruggegeven)
-
-\begin{figure}
-%\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
-\caption{  Bifurcation diagram of the steady states calculated with the Newton-Krylov-Solver for the PDE.
-$\Delta t = 10^{-4}, \Delta T = 10^{-2}, \Delta x = 10^{-2}, \Delta D = 0.01, \epsilon_{GMRES}=10^{-5},  \delta_{Newton} = 10^{-7}, \epsilon=10^{-5}$
-}
-\end{figure}
-
-\subsection{Results for the sde}
-We chose $\Delta \sigma = 0.05$ as continuation step seize.
-Performing a bifurcation with smaller steps is pointless, because the difference in densities will become too small in comparison with the noise on the residual.
-
-\begin{figure}
-%\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
-\caption{  Bifurcation diagram of the steady states calculated with the Newton-Krylov-Solver for the SDE.
-$\Delta t = 10^{-3}, \Delta T = 10^{-1}, \Delta x = 10^{-2}, \epsilon_{GMRES}=10^{-5},  \delta_{Newton} = 10^{-7}, \epsilon=10^{-5}$
-}
-\end{figure}
-
-%\begin{center}
-%\begin{table}
-%\caption{Parameter values}
-%  \begin{tabular} { | c  | c |}    \hline   
-%     %FTCS-scheme  &   Euler-Maruyama-scheme  \\  
-%   \textit{ {Discretization parameters}}&    &  SDE    \\ \hline
-%       $\Delta t $  & $10^{-4}$ \\ 
-%       $\Delta T $ &  $10^{-2}$ \\ 
-%      $ \Delta x &  $10^{-2}$ \\ 
-%       $ \Delta D$  $  &  $ 0.01 $ \\ \hline
-%  \end{tabular}
-%\end{table}
-%\end{center}
-
-\subsubsection{Calculating fixed points}
-The steady  states are computed with the Newton-Krylov solver described in section \ref{sec:Newton-Krylov} and compared with the approximated analytical solutions, calculated by Garnier et.
-al for small $h$ \cite{Garnier}.
-
-\begin{center}
-\begin{figure}
-\includegraphics[width=1\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/steady_states.pdf}
-\caption{\label{fig:bif_anal}}
-\end{figure}
-\end{center}
-
-\subsubsection{Continuation}
-We use a pseudo-arclength contunaution method with secant prediction steps.
-In fig. \ref{todo} we show a bifurcation diagram of the densities.
-
-\begin{center}
-\begin{figure}
-\includegraphics[width=0.8\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/bif_alpha}
-\caption{\label{fig:bif_anal}}
-\end{figure}
-\end{center}
-
-%\begin{equation}
-%\xi = sqrt(1-3*C)*(1+ 6*mu/(sigma**2)*C**2*(1-2*C)/(1-3*C))
-%\end{equation}  
-%
-%
-
-By adding the mean reversion term, the Fokker-Plack equation \eqref{fokkerplanck}  describing the evolution of the density, now becomes
-
-\begin{equation}
-\label{fp_mean_field}
-\pa{\rho(x,t)}{t} = - \mu \pa{( V(x) \rho(x,t))}{x} - \alpha \pa{}{x} \left[ \left(\int x \rho(x,t) \mathrm{d}x  -x \right) \rho(x,t) \right] + \frac{\sigma^2}{2}  \ppa{\rho(x,t)}{x} .
-\end{equation}
-
-Explicit solutions of eq. \label{fp_mean_field} are not available in general, but we can find equilibrium solutions. Assuming
-
-\section{Convergence of the Newton Method}
-\input{newton_convergence.tex}
 

### 2016-07-04 Documentation/newton_convergence.tex
last bifurcation results

@@
 In fact, for low particle numbers (and a high variance on the left and right side of the linear equation we need to solve), we observe more failed Newton-iterations if the GMRES-tolerance is chosen too small.
+
 A question of interest is then `what is a good value for the GMRES-tolerance?'.
@@
 
-\caption{ The non-linear solver converges after 1 or 2 Newton iterations, up to a tolerance which depends on the number of particles $N$ (\textit{left}).
+\caption{ The non-linear solver converges after 1 Newton iteration, up to a tolerance which depends on the number of particles $N$ (\textit{left}).
 This best achieved tolerance converges to zero with  $\mathcal{O}(\frac{1}{\sqrt{N}})$ (\textit{right}).
@@
 \centering
-\includegraphics[width=0.5\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Newton_sde_res(k)_Dt_e-2_tol1e-4}
-\includegraphics[width=0.49\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Tolerance_on_NK-solution_converges_N-1_tol_1e-4}
+\includegraphics[width=0.32\linewidth]{../Problems/WeightedParticles/checkSystem/plots/Variance_3th_newton_it_Dt1e-1}
+\includegraphics[width=0.32\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Newton_sde_res(k)_Dt_e-2_tol1e-4}
+\includegraphics[width=0.32\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/Tolerance_on_NK-solution_converges_N-1_tol_1e-4}
 
-\caption{ The best achieved tolerance for $N=10^8$ is limited because of the size of the GMRES-tolerance ($\epsilon_{\texttt{GMRES}}=10^{-4}$)
-}
+\caption{If the GMRES-tolerance is chosen too small, failed Newton-iterations blew up the variance (in simulations with few particles), as shown in the left plot.
+However if the GMRES-tolerance is too big, it limits the accuracy of the Newton-Krylov-solver in simulations with many particles.
+For instance, the best achieved tolerance for $N=10^8$ is limited because of the size of the GMRES-tolerance ($\epsilon_{\texttt{GMRES}}=10^{-4}$), as shown in the two plots on the right.}
 \label{fig:Newton_sde_res(k)_tol1e-4}
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.7\linewidth]{../Problems/WeightedParticles/checkSystem/Newton/plots/adapt_GMREStol_every_Nstep}
+\caption{}
+\label{fig:adapt_GMREStol_every_Nstep}
+\end{figure}
+
 %

### 2016-07-04 Documentation/notes.tex
last bifurcation results

\documentclass[]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[utf8]{inputenc} 
\usepackage{subcaption}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{amsbsy}
\usepackage[left=4cm]{geometry}
 %http://tex.stackexchange.com/questions/595/how-can-i-get-bold-math-symbols
%opening
\title{Variance reduction in coarse bifurcation analysis of stochastic models}
\author{Pieter Van Nuffel}

\newcommand{\R}{\ensuremath{\mathbb{R}}} % commando zonder argumenten
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\E}{\ensuremath{\mathbb{E}}}
\newcommand{\norm}[1]{\left\|#1\right\|} % commando's met argumenten
\newcommand{\pa}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\ppa}[2]{\frac{\partial^2 #1}{\partial #2^2}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\U}{\ensuremath{\boldsymbol{\rho}}}
\newcommand{\cts}{\ensuremath{\boldsymbol{\Phi}^N_T}} %Coarse time step
\newcommand{\V}{\ensuremath{\mathbf{v}}} 
\newcommand{\jv}{\ensuremath{\mathbf{\hat{Jv}}}}
\newcommand{\jvpde}{\ensuremath{\mathbf{Jv}_{FP}}}

\theoremstyle{definition}
\newtheorem{Theorem}{Theorem}

\begin{document}
\subsubsection{Nota's}
Tijdens elke continueringsstap, wordt een nieuw punt gecreëerd dat initieel de dichtheid \U krijgt van het vorig punt en met een parameter (D) die horizontaal word geupdate. Dan wordt de norm van het residu berekend $\norm{ \U(t) - \U(t+\Delta t)} $, waarin de dichtheid op de volgende tijdsstap verkregen wordt door een simulatie van het systeem met de nieuwe parameter. De grootte van het residu is dus afhankelijk van de stapgrootte in de bifurcatieparamter. De newtontolerantie is de norm van het residu. Als de newton-tolerantie te groot gekozen wordt (in vergelijking met de stapgrootte van het residu), dan wordt er geen Newton-iteratie uitgevoerd. Als de Newton-tolerantie evenwel te klein gekozen wordt, dan bereikt de Newton-solver mogelijk geen convergentie ten gevolge van de ruis. De Newton-tolerantie moet dus weloverwogen gekozen worden.

\subsection{Results for the pde}
Ook hier zien we dat de keuze van de newtontolerantie belangrijk is (te groot = kans op dezelfde toestanden , te klein= failed iterations).
Dit kan echter opgevangen worden door de oplossing van dx nauwkeuriger te maken door de tolerantie van de GMRES te verfijnen (het probleem was dat er dx=0 oplossingen werden teruggegeven)

\begin{figure}
%\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
\caption{  Bifurcation diagram of the steady states calculated with the Newton-Krylov-Solver for the PDE. $\Delta t = 10^{-4}, \Delta T = 10^{-2}, \Delta x = 10^{-2}, \Delta D = 0.01, \epsilon_{GMRES}=10^{-5},  \delta_{Newton} = 10^{-7}, \epsilon=10^{-5}$
}
\end{figure}

\subsection{Results for the sde}
We chose $\Delta \sigma = 0.05$ as continuation step seize. Performing a bifurcation with smaller steps is pointless, because the difference in densities will become too small in comparison with the noise on the residual.

\begin{figure}
%\includegraphics{../Problems/WeightedParticles/checkSystem/plots/bifurcation_pde(D)}
\caption{  Bifurcation diagram of the steady states calculated with the Newton-Krylov-Solver for the SDE. $\Delta t = 10^{-3}, \Delta T = 10^{-1}, \Delta x = 10^{-2}, \epsilon_{GMRES}=10^{-5},  \delta_{Newton} = 10^{-7}, \epsilon=10^{-5}$
}
\end{figure}

%\begin{center}
%\begin{table}
%\caption{Parameter values}
%  \begin{tabular} { | c  | c |}    \hline   
%     %FTCS-scheme  &   Euler-Maruyama-scheme  \\  
%   \textit{ {Discretization parameters}}&    &  SDE    \\ \hline
%       $\Delta t $  & $10^{-4}$ \\ 
%       $\Delta T $ &  $10^{-2}$ \\ 
%      $ \Delta x &  $10^{-2}$ \\ 
%       $ \Delta D$  $  &  $ 0.01 $ \\ \hline
%  \end{tabular}
%\end{table}
%\end{center}

\end{document}

### 2016-07-04 Documentation/presentation_12-07.tex
last bifurcation results

\documentclass{beamer}
%\usepackage[dutch]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[makeroom]{cancel}
\usepackage{color}
\usepackage{color}
\useoutertheme{default}
\newcommand<>{\xxcancel}[1]{\alt#2{\cancel{#1}\vphantom{#1}}{#1}} %has apparently no effect
%\renewcommand{\CancelColor}{blue}

\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\N}{\ensuremath{\mathbb{N}}}
\newcommand{\E}{\ensuremath{\mathbb{E}}}
\newcommand{\norm}[1]{\left\|#1\right\|} % commando's met argumenten
\newcommand{\pa}[2]{\frac{\partial #1}{\partial #2}}
\newcommand{\ppa}[2]{\frac{\partial^2 #1}{\partial #2^2}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\U}{\ensuremath{\boldsymbol{\rho}}}
\newcommand{\cts}{\ensuremath{\boldsymbol{\Phi}_T}} %Coarse time step
\newcommand{\V}{\ensuremath{\mathbf{v}}}
\newcommand{\jv}{\ensuremath{\mathbf{\hat{Jv}}}}
\newcommand{\jvpde}{\ensuremath{\mathbf{Jv}_{FP}}}

\begin{document}

\title[Systemic Risk] % (optional, only for long titles)
{Variance Reduction in Equation-free Newton-Krylov-Methods}
%\subtitle{Evidence from India}
%\author[Author, Anders] % (optional, for multiple authors)
%{F.~Author\inst{1} \and S.~Anders\inst{2}}
%\institute[Universities Here and There] % (optional)
%{
%  \inst{1}%
%  Institute of Computer Science\\
%  University Here
%  \and
%  \inst{2}%
%  Institute of Theoretical Philosophy\\
%  University There
%}
\date{July 12, 2016} % (optional)
%{Conference on Presentation Techniques, 2004}
%\subject{Computer Science}

\begin{frame}{\titlepage}
\includegraphics[width=0.63\paperwidth,height=0.42\paperheight]{/home/pieter/riskmodel/domino-effect2}
%\includegraphics[width=0.6\paperwidth,height=0.53\paperheight]{/home/pieter/riskmodel/domino-effect3}
\end{frame}

\tableofcontents

\section{Model Problem}

\begin{frame}
\frametitle{Fokker-Planck equation}

\begin{equation*}
\label{fokkerplanck}
\pa{\rho(x,t)}{t} + \pa{(\mu(x) \rho(x,t))}{x} = \frac{\sigma^2}{2}  \ppa{\rho(x,t)}{x} 
\end{equation*}
%\begin{center}
%%\includegraphics[width=6.5cm]{../Problems/WeightedParticles/checkSystem/plots/rho_pde}
%\end{center}

Simulate an ensemble of $N$ particles evolving according to the corresponding SDE
\begin{equation*}
    \dd X = \mu f(X) \dd t + \sigma  \dd{W_t}.
\end{equation*}

\end{frame}

\begin{frame}

\frametitle{Coarse time stepper}

\begin{equation*}
\U(t + n \Delta t) = \cts(\U) = (\mathcal{R} \circ \mathcal{E}(n \Delta t, \omega) \circ  \mathcal{L}(\mathbf{\omega})  ) (\U(t))
\end{equation*}

\begin{itemize} [<+->]
\item 
\begin{block} 
{Lifting $\mathcal{L}: \U \rightarrow \mathbf{x}  $}
\end{block}
\item
 \begin{block} {Evolution  $\mathcal{E}$: \\ Simulation of the SDE for $N$ particles over $n$ timesteps. }
\begin{equation*}
   \mathbf{X}^{n + 1} = \mathbf{X}^{n} + \mu (\mathbf{X}^n) \Delta t +  \sqrt{\sigma \Delta t}\cdot \boldsymbol{\xi}^n 
\end{equation*}
with \begin{equation*}\xi^n(\omega)  \sim \mathcal{N} (0,1)
\end{equation*}
\end{block}

\item 
\begin{block} {Restriction $\mathcal{R}: \mathbf{x} \rightarrow \U $}
\begin{equation*}
\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(X_i) = \rho_j  
\end{equation*}
\end{block}

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Evaluating the bias on Jacobian-vector products}

\begin{equation*} 
 \mathbf{Bias}(\jv, \jvpde ) = \mathbb{\hat{E}}[\jv]  - \jvpde 
\end{equation*}

\pause

with 
\begin{equation*}
\jvpde \approx \frac{ \U^{\varepsilon}(t + n \Delta t)  - \U (t +  n\Delta t)}{\varepsilon}
\end{equation*}
 \pause

and $\U$ calculated by explicitly solving

\begin{equation*}
\label{fokkerplanck}
\pa{\rho(x,t)}{t} + \pa{(\mu(x) \rho(x,t))}{x} = D  \ppa{\rho(x,t)}{x} 
\end{equation*}

 using

\begin{equation*} 
\rho_i^{n+1} = \rho_i^n + \Delta t \left( \frac{D}{{\Delta x}^2} \left( \rho_{i+1}^{n} - 2 \rho_i^n + \rho_{i-1}^n \right)  - \frac{a(x)}{\Delta x} (\rho_i^n - \rho_{i-1}^n) \right)
\end{equation*}
for the value of $\rho$ at position $x=i \Delta x$ and time $t=(n+1) \Delta t$

\end{frame}

\begin{frame}
\frametitle{Evaluating the bias on Jacobian-vector products}

\begin{equation*} 
 \mathbf{Bias}(\jv, \jvpde ) = \mathbb{\hat{E}}[\jv]  - \jvpde 
\end{equation*}

 \begin{figure}
 \includegraphics[width=9.5cm]{../Problems/WeightedParticles/checkSystem/plots/Bias_Jv_N_M.pdf}

 \end{figure}

\end{frame}

\begin{frame}
\frametitle{Evaluating the bias on Jacobian-vector products}

\begin{equation*} 
 \mathbf{Bias}(\jv, \jvpde ) = \mathbb{\hat{E}}[\jv]  - \jvpde 
\end{equation*}

 \begin{figure}
 \includegraphics[width=9.5cm]{../Problems/WeightedParticles/checkSystem/plots/Bias_Jv_M_N.pdf}

 \end{figure}

\end{frame}

\section{Algorithm}

\begin{frame} {Newton-Krylov solver}

Computing steady states $\U^*$ by solving the non-linear system

\begin{equation*}
  {F}(\U^*) = \U^* - \cts(\U^*) =0 \label{non-linear_system}
\end{equation*}

\pause
%where $\mathbf{F}$ is nonlinear function from $\R^N$ to  $\R^N$. 
%To find the steady state $\U_*$, we apply Newton's method to eq. \ref{non-linear_system} . Starting from an initial state $\U^0$, 
\begin{itemize}
\item Starting from an initial state $\U^0$, we iterate
\begin{eqnarray*}
\begin{cases}
& \texttt{Solve         }  J(\U^k) \boldsymbol{\delta_k}  =  - {F}(\U^k) \label{linear_system}        \\          
& \texttt{Set     } \U^{k+1} = \U^k+ \boldsymbol{\delta_k} 
\end{cases}
\end{eqnarray*}
until convergence

\pause

%\item Each Newton iteration $n$ involves evaluating the Jacobian of the timestepper $J(\cts(\U))$
\item  No explicit formula for $J(\cts)$ $\Rightarrow$ using iterative method (GMRES) that only requires Jacobian-vector products

\pause
\item  These are estimated by a finite difference approximation

\begin{equation*}
\label{Jv_approx}
J(\cts) \cdot \mathbf{v} \approx \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} 
%&\approx & \frac{\cts (\U, \boldsymbol{\omega_1} )  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cts (\U, \boldsymbol{\omega_2}) }{\varepsilon} \nonumber
\end{equation*}

%\pause
%\item Problem: $ \texttt{Var}(\jv)  \sim \mathcal{O}(1/(\varepsilon^2  N)) $\\$ \Rightarrow $     Numerical noise for $ \varepsilon \ll 1$    
%%\end{frame}
%\end{itemize}

\pause
\begin{block}{Problem}
 $ \texttt{Var}(\jv)  \sim \mathcal{O}(1/(\varepsilon^2  N)) \Rightarrow $     Numerical noise for $ \varepsilon \ll 1$ 
 \end{block}  
%\end{frame}
\end{itemize}
\end{frame}
%
%\begin{frame} \frametitle{Variance reduction of Jacobian-vector products}
%
%\begin{equation*}
%\label{Jv_approx}
%J(\cts) \cdot \mathbf{v} \approx \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} 
%%&\approx & \frac{\cts (\U, \boldsymbol{\omega_1} )  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cts (\U, \boldsymbol{\omega_2}) }{\varepsilon} \nonumber
%\end{equation*}
%Problem: $ \texttt{Var}(\jv)  \sim \mathcal{O}(1/(\varepsilon^2  N)) \Rightarrow $ 
%    Numerical noise for $ \varepsilon \ll 1$    
%\end{frame}
%

\begin{frame} \frametitle{Variance reduction of Jacobian-vector products}

\begin{eqnarray}
\mathbf{Jv} = D(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \nonumber \\
&\approx & \frac{\cts (\U, \boldsymbol{\omega_1} )  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cts (\U, \boldsymbol{\omega_2}) }{\varepsilon} \nonumber
\end{eqnarray}

\pause 
\begin{block}{Solution: \\ Perturbations on the density $\rightarrow$  perturbations in the weights }
\begin{eqnarray*}
\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(x_i) &=& \rho_j  \\
\frac{1}{N} \sum_{i=1}^{N}  w^i_{\varepsilon} \cdot \chi_{\Delta_j} (x^i) &=& \rho_j + \varepsilon v_j .
\end{eqnarray*}
For the perturbated density: compute the weight per bin as $ w^j_{\varepsilon} = 1+ \frac{\varepsilon v_j}{\rho_j}$ and assign this value to each particle in $\Delta_j$.
\end{block}

\end{frame}

\begin{frame} \frametitle{Variance reduction of Jacobian-vector products}

\begin{eqnarray}
\mathbf{Jv} = D(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \nonumber \\
&\approx &  \frac{ \cancel{\cts (\U, \boldsymbol{\omega_1} ) }  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \cancel{ \cts (\U, \boldsymbol{\omega_1}) } }{\varepsilon} \nonumber
\end{eqnarray}

\begin{block}{Solution: \\ Perturbations on the density $\rightarrow$  perturbations in the weights }
\begin{eqnarray*}
\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(x_i) &=& \rho_j  \\
\frac{1}{N} \sum_{i=1}^{N}  w^i_{\varepsilon} \cdot \chi_{\Delta_j} (x^i) &=& \rho_j + \varepsilon v_j .
\end{eqnarray*}
For the perturbated density: compute the weight per bin as $ w^j_{\varepsilon} = 1+ \frac{\varepsilon v_j}{\rho_j}$ and assign this value to each particle in $\Delta_j$.
\end{block}

\end{frame}

\begin{frame} \frametitle{Variance reduction of Jacobian-vector products}

\begin{eqnarray}
\mathbf{Jv} = D(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \nonumber \\
&\approx &  \frac{ \xxcancel{\cts (\U, \boldsymbol{\omega_1} ) }  + \varepsilon D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \xxcancel{ \cts (\U, \boldsymbol{\omega_1}) } }{\varepsilon} \nonumber
\end{eqnarray}

\begin{block}{Solution: \\ Perturbations on the density $\rightarrow$  perturbations in the weights }
\begin{eqnarray*}
\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(x_i) &=& \rho_j  \\
\frac{1}{N} \sum_{i=1}^{N}  w^i_{\varepsilon} \cdot \chi_{\Delta_j} (x^i) &=& \rho_j + \varepsilon v_j .
\end{eqnarray*}
For the perturbated density: compute the weight per bin as $ w^j_{\varepsilon} = 1+ \frac{\varepsilon v_j}{\rho_j}$ and assign this value to each particle in $\Delta_j$.
\end{block}
\end{frame}

\begin{frame} \frametitle{Variance reduction of Jacobian-vector products}

\begin{eqnarray}
\mathbf{Jv} = D(\cts) \cdot \mathbf{v} &\approx& \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} \nonumber \\
&\approx &  \frac{ \xxcancel{\cts (\U, \boldsymbol{\omega_1} ) }  + \xxcancel{\varepsilon} D(\cts) (  \U, \boldsymbol{\omega_1})  \cdot \mathbf{v}  - \xxcancel{ \cts (\U, \boldsymbol{\omega_1}) } }{\xxcancel{\varepsilon}} \nonumber
\end{eqnarray}

\begin{block}{Solution: \\ Perturbations on the density $\rightarrow$  perturbations in the weights }
\begin{eqnarray*}
\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(x_i) &=& \rho_j  \\
\frac{1}{N} \sum_{i=1}^{N}  w^i_{\varepsilon} \cdot \chi_{\Delta_j} (x^i) &=& \rho_j + \varepsilon v_j .
\end{eqnarray*}
For the perturbated density: compute the weight per bin as $ w^j_{\varepsilon} = 1+ \frac{\varepsilon v_j}{\rho_j}$ and assign this value to each particle in $\Delta_j$.
\end{block}
\end{frame}

\begin{frame}
\frametitle{Variance reduction of Jacobian-vector products}

\begin{equation*} 
 \texttt{Var}(\jv) =  \hat{\mathbb{E}} \left[   \left(  \jv - \hat{\mathbb{E}}[\jv]   \right)^2 \right] \sim \mathcal{O}(1/ N)
\end{equation*}

\begin{figure}
\includegraphics[width=8cm]{../Problems/Particles/checkSystem/plots/Var_N_eps_nw}
\end{figure}

\end{frame}

\section{Systemic risk}

\begin{frame} {Mean Field Model}

\begin{block}{Interaction between components}
\begin{itemize} 
\item Adding \emph{mean field interaction} to model:  each particle feels an attractive force towards the mean state (each agent tends to follow the state of the majority) 
\begin{equation*} 
\label{SDE_meanfield}
    \dd X_j = \mu f(X_j) \dd t + \sigma  \dd{W_j} + \alpha(\bar{X} -X_j) \dd t 
\end{equation*}
\item Interconnectedness between agents can affect the stability of the whole system %, causing systemic risk 
\end{itemize}
\end{block}

\pause

\begin{block}{Application: Systemic Risk in Banking Systems}
\begin{itemize}
\item Banks cooperate. By spreading the risk of
credit shocks, they try to minimize their own risk.
\item  However, this increases the  risk that they may all fail 
\end{itemize}
\end{block}

\end{frame}

\begin{frame} {Mathematical Model for Systemic Risk}

\begin{equation*} 
\label{SDE_meanfield}
    \dd X_j = \mu f(X_j) \dd t + \sigma  \dd{W_j} + \alpha(\bar{X} -X_j) \dd t 
\end{equation*}

\begin{itemize}
\item $\mu$ The intrinsic stability of each component
\item $\sigma$ The strength of external random perturbations to the system
\item $\alpha$ The degree of interconnectedness between agents
\end{itemize}

\begin{figure}
\centering
\includegraphics[width=0.8\linewidth]{./Risk_potential_cartoon}
\label{fig:Risk_potential_cartoon}
\end{figure}

%
%\end{itemize}

\end{frame}

\begin{frame} {Metastable Coarse States}

% \item Strength of mean field interaction $\alpha$ = bifurcation parameter
\begin{minipage}{1\textwidth}
\includegraphics[width=0.53\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha1}
\includegraphics[width=0.53\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha5}
\end{minipage}
\pause
%Upon increasing interaction strength $\alpha$, two metastable coarse states arise %$\rightarrow$ two new macroscopic states
\begin{figure}
\centering
\includegraphics[width=0.52\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/bif_alpha}
\label{fig:bif_alpha}
\end{figure}

\end{frame}

\begin{frame}
\frametitle{Analytic Solution for Equilibrium Distribution}

\begin{equation*}
\label{fp_mean_field}
\pa{\rho}{t} = - \mu \pa{( f(x) \rho)}{x} - \alpha \pa{}{x} \left[ \left(\int x \rho \mathrm{d}x  -x \right) \rho \right]   + \frac{\sigma^2}{2}  \ppa{\rho }{x} .
\end{equation*}

\pause

Assuming that \(\xi = \lim_{t \to \infty} \int x \rho(x,t) \mathrm{d}x \), an equilibrium solution satisfies
\begin{equation*}
-\mu \pa{( f(x) \rho_{\xi}) }{x}   - \alpha \pa{}{x} \left [( \xi -x) \rho_{\xi}  \right] +\frac{\sigma^2}{2} \ppa{\rho_{\xi}}{x}=0
\end{equation*}

\pause

The non-zero solutions $\pm\xi$ are

\begin{equation*}
\xi = \pm \sqrt{1-3\frac{\sigma^2}{2\alpha}}\left(1+
\ \mu \frac{6}{\sigma^2}\left(\frac{\sigma^2}{2\alpha}\right)^2\frac{1-2\frac{\sigma^2}{2\alpha} }{1-3\frac{\sigma^2}{2\alpha}}\right) + \mathcal{O}(\mu^2)
\end{equation*}

\end{frame}

\begin{frame} {Calculate fixed points by applying variance reduced Newton-Krylov-solver}

\begin{equation*}
  {F}(\U^*) = \U^* - \cts(\U^*) =0 \label{non-linear_system}
\end{equation*}
\begin{eqnarray*}
\begin{cases}
& \texttt{Solve         }  J(\U^k) \boldsymbol{\delta_k}  =  - {F}(\U^k) \label{linear_system}        \\          
& \texttt{Set     } \U^{k+1} = \U^k+ \boldsymbol{\delta_k} 
\end{cases}
\end{eqnarray*}

\pause

\begin{itemize}
\item How to choose the Newton tolerance?
\item Which time window to choose for the coarse time stepper \cts? 
\end{itemize}
\begin{figure}
\centering
\includegraphics[width=0.52\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/steady_states_upd}
\label{fig:steady_states}
\end{figure}

\end{frame}

\begin{frame} {Estimating Stopping Criterion for the Non-linear Solver}

\begin{itemize}

\item The accuracy on the Newton-Krylov solution is inevitably limited by the noise on the stochastic coarse-time-stepper

\item When the Newton-Krylov solution is
converged, it stays oscillating around the true solution with a standard deviation depending
on the number of particles.

\end{itemize}

\begin{figure}
\includegraphics[width=0.57\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/res_norm_gmres_e-5_alpha5_mu0p05}
\includegraphics[width=0.56\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Tolerance_on_NK-solution_converges_N-1}
\end{figure}
%The needed number of Newton iterations and the tolerance after convergence depend on the number of particles $N$ (\textit{left}). This best achieved tolerance converges to zero with  $\mathcal{O}(\frac{1}{\sqrt{N}})$

\end{frame}

\begin{frame} {Estimating $\Delta T$}
%Calculate steady states $\U^*$ by applying variance reduced Newton-Krylov-solver

\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states(DT)_it0}
\label{fig:Newton_states(DT)_it0}
\end{figure}
\end{frame}
\begin{frame} {Estimating $\Delta T$}
\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states(DT)_it1}
\label{fig:Newton_states(DT)_it1}
\end{figure}
\end{frame}
\begin{frame} {Estimating $\Delta T$}
\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states(DT)_it2}
\label{fig:Newton_states(DT)_it2}
\end{figure}
\end{frame}
\begin{frame}  {Estimating $\Delta T$}
\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states(DT)_it3}

\label{fig:Newton_states(DT)_it3}
\end{figure}
\end{frame}
\begin{frame}  {Estimating $\Delta T$}
\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states(DT)_it4}
\label{fig:Newton_states(DT)_it4}
\end{figure}
\end{frame}
\begin{frame}  {Estimating $\Delta T$}
\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states(DT)_it5}
\label{fig:Newton_states(DT)_it5}
\end{figure}
\end{frame}
\begin{frame}  {Estimating $\Delta T$}
\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states(DT)_it6}
\label{fig:Newton_states(DT)_it6}
\end{figure}
\end{frame}
\begin{frame}  {Estimating $\Delta T$}
\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states(DT)_it7}
\label{fig:Newton_states(DT)_it7}
\end{figure}
\end{frame}
\begin{frame}  {Estimating $\Delta T$}
\begin{figure}
\centering
\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states(DT)_it8}
\label{fig:Newton_states(DT)_it8}
\end{figure}

\end{frame}

\begin{frame}  {Estimating $\Delta T$}
\begin{figure}
\centering
\includegraphics[width=0.89\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states_abs_bias_N1e6}

\end{figure}

\end{frame}

\begin{frame}   {Estimating $\Delta T$}
\begin{figure}
\centering
\includegraphics[width=0.82\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states_abs_bias_N1e4_5_6_7}
\end{figure}
\end{frame}

\begin{frame} {Efficiency compared with direct simulation}
\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/bias_compare_Newton_direct}
\label{fig:bias_compare_Newton_direct}
\end{figure}

\end{frame}

%\begin{frame}{Bifurcation diagram}

%
%\begin{figure} % {Bifurcation diagram}
%
%%Increment the continuation parameter and solve a new problem using the previous solution as an initial guess
%
%\centering
%\includegraphics[width=0.9\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/bif_alpha}
%\label{fig:bifurcation}
%\end{figure}
%\end{frame}

\begin{frame}{Bifurcation diagram}

\begin{figure} % {Bifurcation diagram}

%Increment the continuation parameter and solve a new problem using the previous solution as an initial guess

\centering
\includegraphics[width=0.9\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/bifurcation_0}
\label{fig:bifurcation}
\end{figure}

\end{frame}

\begin{frame}{Bifurcation diagram}

\begin{figure} % {Bifurcation diagram}

%Increment the continuation parameter and solve a new problem using the previous solution as an initial guess

\centering
\includegraphics[width=0.9\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/bifurcation_1}
\label{fig:bifurcation}
\end{figure}

\end{frame}

\begin{frame}{Bifurcation diagram}

\begin{figure} % {Bifurcation diagram}

%Increment the continuation parameter and solve a new problem using the previous solution as an initial guess

\centering
\includegraphics[width=0.9\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/bifurcation_fin}
\label{fig:bifurcation}
\end{figure}

\end{frame}

\section{Research Plan}

\section{Other}

\begin{frame} {Courses}

\begin{block}{Contribution to Education}
\begin{itemize}
\item Exercises Analysis 1
\item Exercises Analysis 2
\end{itemize}
\end{block}

\begin{block}{Doctoral Training Programme}
\begin{itemize}
\item Seminar scientific integrity 
\item Teacher assistant training 
\item Science Communication and Outreach 
\end{itemize}
\end{block}

\begin{block}{Other courses}
\begin{itemize}
\item Functional Analysis
\end{itemize}
\end{block}
\end{frame}

%\cite{Garnier}
%\nocite{avitabile14}

%\begin{posterbox}[name=noisereducedjv, column=1, row=3, below=systemic, above=bottom]{References} %ofJacobian-vector-products}

%\renewcommand\refname{\vskip -0.4cm}
%\bibliographystyle{abbrv}
%\bibliography{biblio.bib}

\end{document}

### 2016-07-04 Documentation/systemic_risk.tex
last bifurcation results

@@
-\documentclass[]{article}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage[utf8]{inputenc} 
-\usepackage{subcaption}
-\usepackage{caption}
-\usepackage{hyperref}
-\usepackage{amsbsy}
-\usepackage[left=4.5cm]{geometry}
-%\usepackage{cancel}
- %http://tex.stackexchange.com/questions/595/how-can-i-get-bold-math-symbols
-%opening
-\title{}
-\author{}
-
-%\newcommand{\R}{\ensuremath{\mathbb{R}}} % commando zonder argumenten
-\newcommand{\C}{\ensuremath{\mathbb{C}}}
-\newcommand{\N}{\ensuremath{\mathbb{N}}}
-\newcommand{\E}{\ensuremath{\mathbb{E}}}
-\newcommand{\norm}[1]{\left\|#1\right\|} % commando's met argumenten
-\newcommand{\pa}[2]{\frac{\partial #1}{\partial #2}}
-\newcommand{\ppa}[2]{\frac{\partial^2 #1}{\partial #2^2}}
-\newcommand{\dd}{\ensuremath{\mathrm{d}}}
-\newcommand{\U}{\ensuremath{\boldsymbol{\rho}}}
-\newcommand{\cts}{\ensuremath{\boldsymbol{\Phi}_T}} %Coarse time step
-\newcommand{\V}{\ensuremath{\mathbf{v}}}
-\newcommand{\jv}{\ensuremath{\mathbf{\hat{Jv}}}}
-\newcommand{\jvpde}{\ensuremath{\mathbf{Jv}_{FP}}}
-%\newcommand{\expect}[1]{\ensuremath { \mathbb{E} \left[ #1  \right] }
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-
 %
@@
 %strategies can be applied to other ABMs. Therefore, we provide a detailed account of
-%the algorithmic steps involved in defining an accurate equation-free Newton-Krylov
+%the algorithmic steps involved in defining anf accurate equation-free Newton-Krylov
 %method and testing its convergence properties
 
-\end{abstract}
+%\section{Systemic risk}
 
-\section{Systemic risk}
+Until now, we considered a system of non-interacting particles.
+We will now extend our model to a mean field model by introducing a third parameter $\alpha$, which is the degree of interaction or cooperation in the system.
+A simple form of cooperative behaviour is the case where each agent tends to follow the state of the majority (or, each particle feels an attractive force towards the mean state of the system).
+To include this cooperation effect in our stochastic simulation, we add this mean reversion term to the SDE \eqref{SDE}:
 
 \begin{equation} 
-\label{fokkerplanck}
-\pa{\rho(x,t)}{t} = - \mu \pa{( U(x) \rho(x,t))}{x} - \alpha \pa{}{x} \left[ \left(\int x \rho(x,t) \mathrm{d}x  -x \right) \rho(x,t) \right] + D  \ppa{\rho(x,t)}{x} 
+\label{SDE_meanfield}
+    \dd x = \mu V(x) \dd t + \sigma  \dd{W_t} + \alpha(\bar{x} -x) \dd t ,
+\end{equation}
+with  $\bar{x}(t) = \frac{1}{N} \sum_{i=1}^{N} x_i(t)$ denoting the empirical mean.
+
+An interesting application of this model to banks and insurance is the emergence of systemic risk.
+Banks will try to minimize their own individual risk by spreading the risk between each other.
+However, this may increase the risk that they may all fail: reducing individual risk on a micro-scale can increase systemic risk on a macro-scale.
+Garnier, Papanicolaou and Yang already made use of the dynamics in eq.
+\eqref{SDE_meanfield}  to show that interconnectedness between agents indeed affects the stability of the whole system, causing systemic risk  \cite{Garnier}.
+They defined $x_i$  as the state of risk of agent $i$.
+The bi-stable-state structure of the potential $V(x)$ ensures that each risk variable stays around $-1$ (defined as the normal state) or $+1$ (the failed state).
+A natural measure of systemic risk is then the transition probability of the empirical mean $\bar{x}$ from the normal state to the failed state.
+
+To establish the idea, let us repeat the numerical simulations with eq.
+\ref{SDE_meanfield}.
+The evolution of the system is now characterized by  by the initial conditions, the three parameters ($\mu$, $\sigma$, $\alpha$) and by the system size $N$.
+Figure \ref{fig:sysrisk}  illustrates the behaviour of the empirical mean $\bar{x}$.
+The simulations were performed with all agents initially in the normal state.
+Nevertheless, if randomness dominates the interaction, the agents can move immediately to the other potential well.
+The system then behaves like $N$ independent diffusions, and hence, by the symmetry of the potential, the mean state will be attracted to a single mixed state $\bar{x}=0$.
+Upon increasing the interaction parameter $\alpha$, however, we find two new macroscopic states, suggesting the presence of a pitchfork bifurcation at the macroscopic level.
+These solutions are no stable steady states, but rather coarse metastable states.
+Their lifetime is linked to the finite system size.
+
+\begin{center}
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha1}
+\includegraphics[width=0.5\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha5}
+
+\caption{The empirical mean, simulated for different $\alpha$. \textit{Left}: the system has one single state $\bar{x}=0$. \textit{Right}: for $\alpha> \alpha_c$ two metastable equilibria emerge. 
+ The red dashed lines are approximated  analytical solutions for the steady states. \label{fig:sysrisk}}
+\end{figure}
+\end{center}
+
+ \begin{center}
+\begin{table}
+\caption{Parameter values}
+  \begin{tabular} { | c  c | c |}    \hline   
+     %FTCS-scheme  &   Euler-Maruyama-scheme  \\  
+   \textit{ {Discretization parameters}}&    &  SDE    \\ \hline
+    Discretization step  & $\Delta x $  & $10^{-2}$ \\ 
+        Number of discretization steps  & $n_x$ &  340 \\ 
+    Time step  &  $\Delta t$ &   $10^{-2}$ \\ 
+       Number of timesteps  & $n$  &  $10^{6}$ \\ \hline
+  \end{tabular}
+%\quad
+%     \begin{tabular}  { | c | c | c |  }  \hline
+%     \multicolumn{3}{|c|} {\textit{ System parameters}   }    \\
+%    \hline  
+%    Diffusion coefficient  & $D $ & $0.5$ \\ 
+%   Drift coefficient &  $\mu$ & $ 1$ \\ \hline
+%  \end{tabular}
+\end{table}
+\end{center}
+
+\subsubsection{Calculating fixed points}
+The steady  states are computed with the Newton-Krylov solver described in section \ref{sec:Newton-Krylov} and compared with the approximated analytical solutions, calculated by Garnier et.
+al for small $h$ \cite{Garnier}.
+
+\begin{center}
+\begin{figure}
+\includegraphics[width=1\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/steady_states.pdf}
+\caption{\label{fig:bif_anal}}
+\end{figure}
+\end{center}
+
+\subsubsection{Continuation}
+We use a pseudo-arclength contunaution method with secant prediction steps.
+In fig. \ref{todo} we show a bifurcation diagram of the densities.
+
+\begin{center}
+\begin{figure}
+\includegraphics[width=0.8\textwidth]{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/bif_alpha}
+\caption{\label{fig:bif_anal}}
+\end{figure}
+\end{center}
+
+By adding the mean reversion term, the Fokker-Planck equation \eqref{fokkerplanck}  describing the evolution of the density, now becomes
+
+\begin{equation}
+\label{fp_mean_field}
+\pa{\rho(x,t)}{t} = - \mu \pa{( f(x) \rho(x,t))}{x} - \alpha \pa{}{x} \left[ \left(\int x \rho(x,t) \mathrm{d}x  -x \right) \rho(x,t) \right] + \frac{\sigma^2}{2}  \ppa{\rho(x,t)}{x} .
 \end{equation}
 
+Explicit solutions of eq.
+\label{fp_mean_field} are not available in general, but we can find equilibrium solutions.
+Assuming that \(\xi = \lim_{t \to \infty} \int x \rho(x,t) \mathrm{d}x \), an equilibrium solution satisfies
+\begin{equation}
+\mu \pa{( f(x) \rho_{\xi}) }{x}   - \alpha \pa{}{x} \left [( \xi -x) \rho_{\xi}  \right] +\frac{\sigma^2}{2} \ppa{\rho_{\xi}}{x}
+\end{equation}
+ -
+and has the form 
+\begin{equation}
+\rho_{\xi}(x) = \frac{1} {Z_{\xi}} \exp{ \left( -\frac{(y-\xi)^2}{2 \frac{\sigma^2}{2 \alpha}} - \mu \frac{2}{\sigma^2} V(x)  \right)}
+\end{equation}
+In addition, the non-zero solutions $\pm\xi$ are
+
+\begin{equation}
+\xi = \pm \sqrt{1-3\frac{\sigma^2}{2\theta}}\left(1+
+\ \mu \frac{6}{\sigma^2}\left(\frac{\sigma^2}{2\theta}\right)^2\frac{1-2\frac{\sigma^2}{2\theta} }{1-3\frac{\sigma^2}{2\theta}}\right) + \mathcal{O}(\mu^2)
+\end{equation}
+
+%\begin{equation}
+%\label{fokkerplanck}
+%\pa{\rho(x,t)}{t} = - \mu \pa{( U(x) \rho(x,t))}{x} - \alpha \pa{}{x} \left[ \left(\int x \rho(x,t) \mathrm{d}x  -x \right) \rho(x,t) \right] + D  \ppa{\rho(x,t)}{x} 
+%\end{equation}\mathrm{d}x
+
+%
+%\begin{figure}
+%\includegraphics{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha5}
+%
+%\end{figure}
+
 \begin{figure}
-\includegraphics{../../riskmodel/Problems/WeightedParticles/checkSystem/plots/xmean_alpha5}
-
+\centering
+\includegraphics[width=0.49\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states_abs_bias_N1e6}
+\includegraphics[width=0.49\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states_abs_bias_N1e4_5_6_7}
+\caption{The bias between the analytical solution $\xi$ and the mean $\bar{x}$ of the density calculated with the Newton-Krylov solver is plotted as a function of the number of Newton iterations $k$.
+This bias converges to zero with a rate depending on the timestep used in the coarse time stepper $\Delta T$, but not on the number of particles $N$.
+($\Delta x = 10^{-2}$, $\Delta t = 10^{-2}$, $\mu=0.05$, $\alpha=5$, $\sigma=1$)}
+\label{fig:Newton_states_bias_mean}
 \end{figure}
 
-\end{document}
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.49\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Newton_states_abs_bias_N1e4_5_6_7_log}
+%\includegraphics[width=0.49\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/bias_steady_states_f(Dt)}
+%\caption{Is the total number of timesteps used in the Newton-Krylov-method (\textit{left}) significantly lower than the number of timesteps used in direct simulation (\textit{right}) to reach the same accuracy?
+Not clear from these plots}
+%\label{fig:bias_steady_states_f(Dt)}
+%\end{figure}
+%
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/bias_compare_Newton_direct}
+\caption{The total number of timesteps used in the Newton-Krylov-method is lower than the number of timesteps  in direct simulation to calculate a fixed point with the same accuracy.
+($N=10^6$,$\Delta x = 10^{-2}$, $\Delta t = 10^{-2}$, $\mu=0.05$, $\alpha=5$, $\sigma=1$) }
+\label{fig:bias_compare_Newton_direct}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.49\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/res_norm_gmres_e-5_alpha5_mu0p05}
+\includegraphics[width=0.49\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/Tolerance_on_NK-solution_converges_N-1}
+\caption{The needed number of Newton iterations and the tolerance after convergence depend on the number of particles $N$ (\textit{left}).
+This best achieved tolerance converges to zero with  $\mathcal{O}(\frac{1}{\sqrt{N}})$ (\textit{right}).
+(Parameter values: $\Delta x = 10^{-2}$, $\Delta t = 10^{-2}$, $\Delta T = 10 $ $\mu=0.05$, $\alpha=5$, $\sigma=1$, $ \epsilon_{\texttt{GMRES}}=10^{-5}$)}
+\label{fig:Tolerance_on_NK-solution_converges_N-1}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.7\linewidth]{/home/pieter/riskmodel/Problems/WeightedParticles/checkSystem/plots/bifurcation}
+\caption{}
+\label{fig:bifurcation}
+\end{figure}

### 2016-07-04 Documentation/talk_21_06.tex
last bifurcation results

--- 2016-07-04 Documentation/presentation_12-07.tex
@@
 
+ % % %
+
 \title[Systemic Risk] % (optional, only for long titles)
-{Variance Reduction in Equation-free Newton-Krylov-Methods}
+{Applying Newton-Krylov Solver on a Model for Systemic Risk}
 %\subtitle{Evidence from India}
@@
 %}
-\date{July 12, 2016} % (optional)
+\date{\today} % (optional)
 %{Conference on Presentation Techniques, 2004}
@@
 
-\tableofcontents
-
-\section{Model Problem}
-
-\begin{frame}
-\frametitle{Fokker-Planck equation}
-
-\begin{equation*}
-\label{fokkerplanck}
-\pa{\rho(x,t)}{t} + \pa{(\mu(x) \rho(x,t))}{x} = \frac{\sigma^2}{2}  \ppa{\rho(x,t)}{x} 
-\end{equation*}
-%\begin{center}
-%%\includegraphics[width=6.5cm]{../Problems/WeightedParticles/checkSystem/plots/rho_pde}
-%\end{center}
-
-Simulate an ensemble of $N$ particles evolving according to the corresponding SDE
-\begin{equation*}
-    \dd X = \mu f(X) \dd t + \sigma  \dd{W_t}.
-\end{equation*}
-
-\end{frame}
-
-\begin{frame}
-
-\frametitle{Coarse time stepper}
-
-\begin{equation*}
-\U(t + n \Delta t) = \cts(\U) = (\mathcal{R} \circ \mathcal{E}(n \Delta t, \omega) \circ  \mathcal{L}(\mathbf{\omega})  ) (\U(t))
-\end{equation*}
-
-\begin{itemize} [<+->]
-\item 
-\begin{block} 
-{Lifting $\mathcal{L}: \U \rightarrow \mathbf{x}  $}
-\end{block}
-\item
- \begin{block} {Evolution  $\mathcal{E}$: \\ Simulation of the SDE for $N$ particles over $n$ timesteps. }
-\begin{equation*}
-   \mathbf{X}^{n + 1} = \mathbf{X}^{n} + \mu (\mathbf{X}^n) \Delta t +  \sqrt{\sigma \Delta t}\cdot \boldsymbol{\xi}^n 
-\end{equation*}
-with \begin{equation*}\xi^n(\omega)  \sim \mathcal{N} (0,1)
-\end{equation*}
-\end{block}
-
-\item 
-\begin{block} {Restriction $\mathcal{R}: \mathbf{x} \rightarrow \U $}
-\begin{equation*}
-\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(X_i) = \rho_j  
-\end{equation*}
-\end{block}
-
-\end{itemize}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Evaluating the bias on Jacobian-vector products}
-
-\begin{equation*} 
- \mathbf{Bias}(\jv, \jvpde ) = \mathbb{\hat{E}}[\jv]  - \jvpde 
-\end{equation*}
-
-\pause
-
-with 
-\begin{equation*}
-\jvpde \approx \frac{ \U^{\varepsilon}(t + n \Delta t)  - \U (t +  n\Delta t)}{\varepsilon}
-\end{equation*}
- \pause
-
-and $\U$ calculated by explicitly solving
-
-\begin{equation*}
-\label{fokkerplanck}
-\pa{\rho(x,t)}{t} + \pa{(\mu(x) \rho(x,t))}{x} = D  \ppa{\rho(x,t)}{x} 
-\end{equation*}
-
- using
-
-\begin{equation*} 
-\rho_i^{n+1} = \rho_i^n + \Delta t \left( \frac{D}{{\Delta x}^2} \left( \rho_{i+1}^{n} - 2 \rho_i^n + \rho_{i-1}^n \right)  - \frac{a(x)}{\Delta x} (\rho_i^n - \rho_{i-1}^n) \right)
-\end{equation*}
-for the value of $\rho$ at position $x=i \Delta x$ and time $t=(n+1) \Delta t$
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Evaluating the bias on Jacobian-vector products}
-
-\begin{equation*} 
- \mathbf{Bias}(\jv, \jvpde ) = \mathbb{\hat{E}}[\jv]  - \jvpde 
-\end{equation*}
-
- \begin{figure}
- \includegraphics[width=9.5cm]{../Problems/WeightedParticles/checkSystem/plots/Bias_Jv_N_M.pdf}
-
- \end{figure}
-
-\end{frame}
-
-\begin{frame}
-\frametitle{Evaluating the bias on Jacobian-vector products}
-
-\begin{equation*} 
- \mathbf{Bias}(\jv, \jvpde ) = \mathbb{\hat{E}}[\jv]  - \jvpde 
-\end{equation*}
-
- \begin{figure}
- \includegraphics[width=9.5cm]{../Problems/WeightedParticles/checkSystem/plots/Bias_Jv_M_N.pdf}
-
- \end{figure}
-
-\end{frame}
+%\tableofcontents
 
 \section{Algorithm}
+
+%\begin{frame}
+%
+%\frametitle{Coarse time stepper}
+%
+%\begin{equation*}
+%\U(t + n \Delta t) = \cts(\U) = (\mathcal{R} \circ \mathcal{E}(n \Delta t, \omega) \circ  \mathcal{L}(\mathbf{\omega})  ) (\U(t))
+%\end{equation*}
+%
+%
+%\begin{itemize} [<+->]
+%\item 
+%\begin{block} 
+%{Lifting $\mathcal{L}: \U \rightarrow \mathbf{x}  $}
+%\end{block}
+%\item
+% \begin{block} {Evolution  $\mathcal{E}$: \\ Simulation of the SDE for $N$ particles over $n$ timesteps. }
+%\begin{equation*}
+%   \mathbf{x}^{n + 1} = \mathbf{x}^{n} + a(\mathbf{x}^n) \Delta t +  \sqrt{2D \Delta t}\cdot \boldsymbol{\xi}^n 
+%\end{equation*}
+%with \begin{equation*}\xi^n(\omega)  \sim \mathcal{N} (0,1)
+%\end{equation*}
+%\end{block}
+%
+%\item 
+%\begin{block} {Restriction $\mathcal{R}: \mathbf{x} \rightarrow \U $}
+%\begin{equation*}
+%\frac{1}{N} \sum_{i=1}^{N}  w_i \cdot \chi_{\Delta_j}(x_i) = \rho_j  
+%\end{equation*}
+%\end{block}fjlk
+%
+%\end{itemize}
+%
+%\end{frame}
+%
+%
+%\begin{frame}
+%\frametitle{Evaluating Jacobian-vector products}
+%\begin{equation*}
+% \U_* - \cts(\U_*) =0
+%\end{equation*}
+%
+%\begin{equation*}
+%\mathbf{Jv} = D(\cts) \cdot \mathbf{v} \approx \frac{\cts (\U + \varepsilon \mathbf{v}, \boldsymbol{\omega_1} )  - \cts (\U, \boldsymbol{\omega_2})}{\varepsilon} 
+%\end{equation*}
+%
+%\pause
+%\begin{block}{Problem: numerical noise for $ \varepsilon \ll 1$ }
+%Repeating $\cts$ with two sets of random numbers $\boldsymbol{\omega_1},  \boldsymbol{\omega_2}$  will give different results
+%$\rightarrow  \mathcal{O}(1/(\varepsilon^2 N))$ variance. 
+%\end{block}
+%
+%\end{frame}
+%
 
@@
 
+\end{itemize}
 \begin{equation*}
@@
 \end{equation*}
-
-%\pause
-%\item Problem: $ \texttt{Var}(\jv)  \sim \mathcal{O}(1/(\varepsilon^2  N)) $\\$ \Rightarrow $     Numerical noise for $ \varepsilon \ll 1$    
-%%\end{frame}
-%\end{itemize}
-
-\pause
-\begin{block}{Problem}
- $ \texttt{Var}(\jv)  \sim \mathcal{O}(1/(\varepsilon^2  N)) \Rightarrow $     Numerical noise for $ \varepsilon \ll 1$ 
- \end{block}  
+%Problem: $ \texttt{Var}(\jv)  \sim \mathcal{O}(1/(\varepsilon^2  N)) \Rightarrow $ 
+%    Numerical noise for $ \varepsilon \ll 1$    
 %\end{frame}
-\end{itemize}
 \end{frame}
@@
 
-\section{Research Plan}
-
-\section{Other}
-
-\begin{frame} {Courses}
-
-\begin{block}{Contribution to Education}
-\begin{itemize}
-\item Exercises Analysis 1
-\item Exercises Analysis 2
-\end{itemize}
-\end{block}
-
-\begin{block}{Doctoral Training Programme}
-\begin{itemize}
-\item Seminar scientific integrity 
-\item Teacher assistant training 
-\item Science Communication and Outreach 
-\end{itemize}
-\end{block}
-
-\begin{block}{Other courses}
-\begin{itemize}
-\item Functional Analysis
-\end{itemize}
-\end{block}
-\end{frame}
-
 %\cite{Garnier}

### 2016-07-04 Continuer/Continuer.py

@@ -62,6 +62,7 @@ class Continuer(object):
         ploty['field'] = 'u'
         ploty['func'] = 'norm'
         param['ploty']=ploty
+        param['add_failures'] = False
 
         param['growth_factor']=1.2
         
@@ -169,6 +170,7 @@ class Continuer(object):
         lambd_step = scipy.zeros_like(self.points[-1].lambd)
         cont_step = self.param['cont_step']
         bifpar_index = self.param['plotx']['func']
+        add_failed_points= self.param['add_failures']
      #   bifpar_index = self.param['bifpar_index']
         lambd_step[bifpar_index] = cont_step       #Insert the index of the prefered bifurcation parameter
         growth = self.param['growth_factor']
@@ -194,6 +196,8 @@ class Continuer(object):
                            
             if not new_success :
                 fail += 1
+                if add_failed_points = True:
+                    self.addPoint(new_point)
 
             else :  # if new_success
 #                for i in range ( len(self.points)):

### 2016-07-04 Problems/WeightedParticles/checkSystem/check_newton_sde.py

@@ -100,7 +100,7 @@ if __name__=="__main__":
    # fp_sde = particles.Particles(lifting,restriction,rho,grid, lambd, param=param)       
 
  #   Nlist = scipy.array([1000, 2000, 4000, 8000, 16000, 32000, 64000, 128000])
-    Nlist = scipy.array([ 1e5]) 
+    Nlist = scipy.array([ 1e7]) 
   #  Nlist = scipy.array([1e5]) #, 800, 1600, 3200])
     eps_list_exponents=[5]
     eps_list= [1e-5]
@@ -127,11 +127,11 @@ if __name__=="__main__":
 #    residual_directsim = scipy.zeros(len(Dtlist)) 
     save_flag= False
 
-#    
-            
-    residual1 = scipy.zeros(nN)
-    residual2 = scipy.zeros(nN)
-    residual3 = scipy.zeros(nN)
+#            
+#    residual1 = scipy.zeros(len(Dtlist))
+#    residual2 = scipy.zeros(len(Dtlist))
+    residual = scipy.zeros(11)
+
     residual_directsim = scipy.zeros(nN) 
     newton_res_norms = scipy.zeros(nN)
     
@@ -183,8 +183,8 @@ if __name__=="__main__":
                 
                             #LINEAR SOLVER
                 gmres_param = GMRES.GMRESLinearSolver.getDefaultParameters()
-                gmres_param['tol']=1e-7   #zet terup 1e-4
-                gmres_param['print']='short'
+                gmres_param['tol']=1e-1   #zet terup 1e-4
+                gmres_param['print']='none'
                 gmres_param['builtin']=False
                 
                 linsolv = GMRES.GMRESLinearSolver(gmres_param)
@@ -197,7 +197,7 @@ if __name__=="__main__":
                 newt_param['rel_tol']=1e-5
                 newt_param['abs_tol']=1e-10
                 newt_param['print']='short'
-                newt_param['max_iter']=10
+                newt_param['max_iter']=8
                 newt_param['damping']=1
                 nsolv = NewtonSolver.NewtonSolver(linsolv,newt_param)
             #    nsolv2 = NewtonSolver.NewtonSolver(linsolv2,newt_param)
@@ -236,15 +236,14 @@ if __name__=="__main__":
              #   rho_sq[n] = rho_sq[n] +   norm(rho_fix)**2
[…]

### 2016-07-04 Problems/WeightedParticles/checkSystem/post_processing_Newton.py

@@ -12,21 +12,35 @@ import matplotlib.pylab as plt
 
 from scipy.linalg import norm
 from scipy import sqrt
+import numpy as np
 
 import matplotlib.pylab as plt
 
 
 if __name__=="__main__":
+    
+    
+    
+    sigma=1.0
+    mu=1.0
+    
+    dx=1e-2
+    xL = -1.7
+    xR = 1.7
+    grid = scipy.arange(xL+dx/2.,xR,dx)
+    
+
+    
+    norm_c = sum( np.exp( (-grid**4 + grid**2)*2*mu/(sigma**2)))*dx
+    rho_ss = np.exp( (-grid**4 + grid**2 )*2*mu/(sigma**2)) /norm_c
+
 
 #    residual_directsim = np.loadtxt('Newton/residual_sim_N1e6.out')
 #    residual1 = np.loadtxt('Newton/residual1_N1e6.out') 
 #    residual2 = np.loadtxt('Newton/residual2_N1e6.out') 
 #    residual3 = np.loadtxt('Newton/residual3_N1e6.out')  
 #    
-#    residual_directsim = np.loadtxt('Newton/residual_sim(N)_Dte-1_tol1e-3.out')
-#    residual1 = np.loadtxt('Newton/residual1(N)__Dte-1_tol1e-3.out') 
-#    residual2 = np.loadtxt('Newton/residual2(N)__Dte-1_tol1e-3.out') 
-#    residual3 = np.loadtxt('Newton/residual3(N)_Dte-1_tol1e-3.out')  
+
 #    
 
 #    resnorm_t100_N1000 = np.loadtxt('Newton/resnorm_Dte-1_N1000')
@@ -240,7 +254,7 @@ if __name__=="__main__":
     plt.yscale('log')
     plt.title('Tol = 1e-5')
     plt.gca().add_patch(triangle_o1var)
-    plt.ylabel(r'$||  \rho^* - \Phi^N_T(\rho^*)||$'  , fontsize = 15)
+    plt.ylabel(r'$||  \hat{\rho^*} - \Phi^N_T(\hat{\rho^*)}||$'  , fontsize = 15)
     plt.xlabel(r'$1/N$', fontsize = 14)
     plt.annotate(order,  xy=(1.1e-6, 1.5e-3), xytext=(1.1e-6, 1.5e-3), fontsize=11, color='grey')
[…]

### 2016-07-04 Problems/WeightedParticles/checkSystem/post_processing_Newton_variance.py

@@ -48,7 +48,7 @@ if __name__=="__main__":
         variance[n] = rho_sq_tol2[n] - norm(E_rho_tol2[n])**2
         variance4[n] = rho_sq_tol4[n] - norm(E_rho_tol4[n])**2
         variance3[n] =  rho_sq_tol3[n] - norm(E_rho_tol3[n])**2 
-        variance1[n] =  rho_sq_tol1[n] - norm(E_rho_tol1[n])**2 
+      #  variance1[n] =  rho_sq_tol1[n] - norm(E_rho_tol1[n])**2 
         variance5[n] =  rho_sq_tol5[n] - norm(E_rho_tol5[n])**2 
     
     log_flag = 'log'
@@ -60,10 +60,14 @@ if __name__=="__main__":
     ax = plt.subplot(111)
     box = ax.get_position()
     ax.set_position([box.x0, box.y0, box.width * 0.95, box.height])   
-    plot_var = plt.plot(Nlist_inv, variance/bins, label=r'$\epsilon_{GMRES}=10^{-2}$', linewidth=2  )
-    plot_var = plt.plot(Nlist_inv, variance3/bins,  label=r'$\epsilon_{GMRES}=10^{-3}$',linewidth=2  )
-    plot_var = plt.plot(Nlist_inv, variance4/bins,  label=r'$\epsilon_{GMRES}=10^{-4}$',linewidth=2  )
-    plot_var = plt.plot(Nlist_inv, variance5/bins,  label=r'$\epsilon_{GMRES}=10^{-5}$',linewidth=2  )
+    
+            
+    points_o1var = [[1e-5, 1e-2], [1e-5, 1e-1], [1e-4, 1e-1]] 
+    triangle_o1var = plt.Polygon(points_o1var, fill=None  ,edgecolor='grey') 
+    plot_var = plt.plot(Nlist_inv, variance/bins, label=r'$\varepsilon_{GMRES}=10^{-2}$', linewidth=2  )
+    plot_var = plt.plot(Nlist_inv, variance3/bins,  label=r'$\varepsilon_{GMRES}=10^{-3}$',linewidth=2  )
+    plot_var = plt.plot(Nlist_inv, variance4/bins,  label=r'$\varepsilon_{GMRES}=10^{-4}$',linewidth=2  )
[…]

### 2016-07-04 Problems/WeightedParticles/checkSystem/post_processing_bif_obs.py

@@ -26,7 +26,7 @@ save_flag= True
 N_points = len(y_mean)
 
 #sigma_list= scipy.arange(2, 0.9, -0.1)   #voor RL
-sigma_list= scipy.arange(1, 2.1, 0.1)  #voor LR
+sigma_list= scipy.arange(1.0, 2.1, 0.1)  #voor LR
         
 for i in range (N_points):
     if (i % 2 == 0): 

### 2016-07-04 Solver/NewtonSolver.py

@@ -92,6 +92,8 @@ class NewtonSolver(Solver.Solver):
         while (iter < max_iter and (stop_norm > abs_tol)):
             iter += 1
             self.nb_newt += 1
+            self.linsolv.param['tol']= self.linsolv.param['tol']/10
+            print "GMRES tolerance = " ,  self.linsolv.param['tol']
             dx, status = self.linsolv.solve(-res)        # res = rho - rho.Dt
             if print_it == 'long':  
                 plt.plot(res)                  
